### 2012-09-07 bidir_pathtrace_skeleton.tex

\documentclass[final,envcountsame]{llncs}

%\usepackage{graphicx} %Pouvoir insérer des images
\usepackage{epsfig}
\usepackage{amsmath} %Permet d'utiliser une jolie panoplie de symboles mathématiques
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsfonts}

\def\mydef#1{{\em #1}}
\def\myem#1{{\em #1}}
\def\myvec#1{\vec{#1}}

\def\Nset{\mathbb{N}}
\def\Zset{\Z}
\def\Rset{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\def\allfaces#1{\mathbb{F}^{#1}}
\def\subfaces#1#2{\allfaces{#1}_{#2}}
\def\subcomplex#1#2{#1 \preceq #2}
\def\complex#1#2{\subcomplex{#2}{\allfaces{#1}}}
\def\cell#1{\hat{#1}}
\def\cellstrict#1{\hat{#1}^*}
\def\cont#1{\check{#1}}
\def\contstrict#1{\check{#1}^*}
\def\closure#1{#1^-}
\def\princ#1{#1^+}
\def\detach#1#2{#2 \oslash #1}
\def\attach#1#2{\mbox{\it{}Attach\/}(\cell{#1},#2)}
\def\mydim#1{\dim(#1)}
\def\type#1{\mbox{\it{}Dir\/}(#1)}
\def\orient#1{\mbox{\it{}Orient\/}(#1)}
\def\border#1{\mbox{\it{}CBorder\/}(#1)}
\def\suchthat{\;|\;}
\def\inter{\cap}
\def\union{\cup}
\def\dinterval#1#2{\{#1,\ldots,#2\}}
\newcommand{\Compl}[1]{\overline{#1}}
\def\finproof{\square}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\newcommand{\Card}[1]{|#1|}

\def\figpath#1{../../Images/EPS/#1}

\begin{document}

\mainmatter

\title{A 3d curvilinear skeletonization algorithm with application to bidirectional path tracing}
\author{John Chaussard \and Venceslas Biri \and Michel Couprie}
\institute{Universit\'{e} Paris Est, LABINFO-IGM, A2SI-ESIEE \\ 2, boulevard Blaise Pascal, Cit\'{e} DESCARTES \\ BP 99  93162 Noisy le Grand CEDEX, France \\ \email{chaussaj@esiee.fr, bertrang@esiee.fr, coupriem@esiee.fr}}

\maketitle

\begin{abstract}
What we propose is really great...
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                        %
%                   INTRODUCTION                         %
%                                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec::intro}
Blah blah blah...

\section{The cubical complex framework}

\subsection{Basic definitions}

In the 3d cubical complex framework, objects are no more made of voxel, but of cubes, squares, lines and vertices $\mathbb{R}$. 
Let $\Z$ be the set of integers, we consider the family of sets $\mathbb{F}^1_0$ and $\mathbb{F}^1_1$, such that $\mathbb{F}^1_0 = \{\{a\} \suchthat a \in \Z\}$ and $\mathbb{F}^1_1 = \{ \{a, a+1\} \suchthat a \in \Z\}$. Any subset $f$ of $\Z^n$ such that $f$ is the cartesian product of $m$ elements of $\mathbb{F}^1_1$ and $(n-m)$ elements of $\mathbb{F}^1_0$ is called a face or an \myem{$m$-face} of $\Z^n$, $m$ is the dimension of $f$, we write $dim(f)=m$. A $0$-face is called a \myem{vertex}, a $1$-face is an \myem{edge}, a $2$-face is a \myem{square}, and a $3$-face is a \myem{cube}.

We denote by $\allfaces{n}$ the set composed of all faces in $\Z^n$. Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.

Let $f \in \allfaces{n}$. We set $\cell{f} = \{g \in \allfaces{n}\vert g \subseteq f\}$, and $\cellstrict{f} = \cell{f} \setminus \{f\}$. Any element of $\cell{f}$ is \myem{a face of $f$}, and any element of $\cellstrict{f}$ is \myem{a proper face of $f$}. We call \myem{star of $f$} the set $\cont{f} = \{g \in \allfaces{n} \vert f \subseteq g\}$, and we write $\contstrict{f} = \cont{f} \setminus \{f\}$: any element of $\cont{f}$ is a \myem{coface of $f$}. It is plain that $g\in\cell{f}$ iff $f\in\cont{g}$.

A set $X$ of faces in $\allfaces{n}$ is a \myem{cell}, or \myem{$m$-cell}, if there exists an $m$-face $f \in X$ such that $X=\cell{f}$. The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$. The set $\Compl{X}$ is $\allfaces{n} \setminus X$.

\begin{definition}
A finite set $X$ of faces in $\allfaces{n}$ is a \myem{cubical complex} if $X=\closure{X}$, and we write $\complex{n}{X}$.

Any subset $Y$ of $X$ which is also a complex is a \myem{subcomplex of $X$}, and we write $\subcomplex{Y}{X}$.
\end{definition}

Informally, in 3d, a set of face is a cubical complex if, for each cube of the complex, the six squares (sides) of the cube also belong to the complex, for each square of the complex, the four edges (sides) of the square also belong to the complex, and for each edge of the complex, the two vertices (extremities) of the edge also belong to the complex.

A face $f \in X$ is \myem{a facet of $X$} if $f$ is not a proper face of any face of $X$. We denote by $\princ{X}$ the set composed of all facets of $X$ . A complex $X$ is \myem{pure} if all its facets have the same dimension. The \myem{dimension of $X$} is $\mydim{X} = \max\{\mydim{f} \suchthat f \in X\}$. If $\mydim{X} = d$, then we say that $X$ is a $d$-complex. The notions of purity and dimensions can be trivially extended to sets of faces.

In $\allfaces{n}$, a complex $X$ is \myem{thin} if $\mydim{X}<n$. In $\allfaces{3}$, a complex is thin if it contains no cube.

\subsection{From binary images to cubical complex}
\label{sec::binary_to_cubic}

Traditionally, a binary image is defined as a finite subset of $\Zset^n$ (with $n=2$ or $n=3$). Given $S \subseteq \Zset^n$, the object voxels are the elements of $S$. This kind of image is the most common one in the field of image processing so, in order to work in cubical complex framework, we need to find a way to transpose a binary image to cubical complex framework.

Informally, to do so, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$ (to a pixel we associate a square, to a voxel we associate a cube). More precisely, let $x=(x_1,...,x_n) \in S$, we define the n-face $\Phi(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$. We can extend the map $\Phi$ to sets: $\Phi(S)=\{\Phi(x) | x \in S\}$. Given a binary image $S$, we associate to it the cubical complex $\closure{\Phi(S)}$ (see Fig.~\ref{fig::voxel_to_complex}).

In the following, most of the objects we consider were indeed binary images which were then transposed into cubical complex framework: this is why most of the two-dimensional complexes we show are pure 2-complexes, and most of the three-dimensional complexes we show are pure 3-complexes.

\subsection{Thinning: the collapse operation}
The collapse operation is the basic operation for performing homotopic thinning of a complex. It consists of removing two distinct elements $(f,g)$ from a complex $X$ under the condition that $g$ is contained in $f$ and is not contained in any other element of $X$. This operation may be repeated several times.

\begin{definition}
Let $\complex{n}{X}$, and let $f,g$ be two faces of $X$. The face $g$ is \myem{free for $X$}, and the pair $(f,g)$ is \myem{a free pair for $X$} if $f$ is the only face of $X$ such that $g$ is a proper face of $f$. 
\end{definition}

In other terms, $(f,g)$ is a free pair for $X$ whenever $\contstrict{g} \inter X = \{f\}$ ($g$ is included only in $f$). It can be easily seen that if $(f,g)$ is a free pair for a complex $X$ and $\dim(f)=m$, then $f$ is a facet and $\dim(g) = m-1$.

\begin{definition}
Let $\complex{n}{X}$, and let $(f,g)$ be a free pair for $X$. The complex $X \setminus \{f,g\}$ is an \myem{elementary collapse of $X$}.

Let $\complex{n}{Y}$, the complex $X$ \myem{collapses onto} $Y$ if there exists a sequence of complexes $(X_0,...,X_\ell)$ of $\allfaces{n}$ such that $X=X_0$, $Y=X_\ell$ and for all $i \in \dinterval{1}{\ell}, X_{i}$ is an elementary collapse of $X_{i-1}$. We also say, in this case, that $Y$ \myem{is a collapse of} $X$.
\end{definition}

We insist here on the fact that, if $Y$ is a collapse of a complex $X$, then $Y$ is a complex.

Let us now introduce some elements that will serve later for proving the thinness of our skeletons.
Let $f_0,f_\ell$ be two $n$-faces of $\allfaces{n}$ (with $\ell$ being even). An \myem{$(n-1)$-path from $f_0$ to $f_\ell$} is a sequence $\pi =
(f_0,...,f_\ell)$ of faces of $\allfaces{n}$ such that for all $i \in \dinterval{0}{\ell}$, either $i$ is even and $f_i$ is an $n$-face, or $i$ is odd and $f_i$ is an $(n-1)$-face with $\contstrict{f_i}=\{f_{i-1}, f_{i+1}\}$ (such path always exists).

\begin{proposition}
\label{prop:thinness}
Let $\complex{n}{X}$ be an $n$-complex, with $n > 0$. Then $X$ has at least one free $(n-1)$-face.
\end{proposition}

\begin{proof}
Since $X$ is an $n$-complex (hence $X$ is finite) there exists an $n$-face $a$ in $X$ and an $n$-face $b$ in $\Compl{X}$. Obviously, there exists an $(n-1)$-path from $a$ to $b$. Let $h$ be the first $n$-face of $\pi$ that is not in~$X$, let $k$ be the last $n$-face of $\pi$ before $h$ (thus $k$ is in $X$), and let $e = k \inter h$ be the $(n-1)$-face of $\pi$ between $k$ and $h$. Since $k$ and $h$ are the only two $n$-faces of $\allfaces{n}$ that contain $e$, we see that the pair $(k,e)$ is free for $X$. $\finproof$
\end{proof}

In conclusion, in $\allfaces{3}$, as long as a complex still contains $3$-faces, it has a free $2$-face and more collapse operations can be performed. Therefore, it is possible to perform collapse on a complex until no more $3$-faces (volumes) can be found (until it is thin).

As in the DT framework, it is sometimes necessary to perform collapse in the cubical complex framework while preserving some faces safe from deletion: these faces are the so-called inhibitor set. When using an inhibitor set during collapse, the guarantee of having a thin result does no more hold. As illustrated on figure ffjkz, the inhibitor set can be thin and a result of the constrained thinning can still not be thin.
However, under some conditions, it is possible to use an inhibitor set while collapsing, and still have the guarantee of getting a thin result. In order to prove this, let us state the following:

\begin{lemma}
\label{lem::remove_small_pairs}
Let $\complex{n}{X}$, let $(f_1,g_1)$ be a free pair of $X$, and $(f_2,g_2)$ be a free pair of $(X \backslash \{f_1,g_1\})$. If $\dim(f_2) > \dim(f_1)$, then $(f_2,g_2)$ is free for $X$. 
\end{lemma}

\begin{proof} If $(f_2,g_2)$ is a free pair of $(X \backslash \{f_1,g_1\})$, then $g_2$ is included in only one face of $(X \backslash \{f_1,g_1\})$, which is $f_2$. As $(\dim(g_2) = \dim(f_2) -1)$ and $(\dim(f_1) \leq (\dim(f_2)-1))$, then $g_2 \nsubseteq f_1$ and $g_2 \nsubseteq g_1$. Therefore, $g_2$ is included in only one face of $X$, which is $f_2$: therefore, $(f_2,g_2)$ is free for $X$.$\finproof$
\end{proof}

This lemma implies that, when one has a sequence of removal of free pairs of faces from a complex, one can only keep the free pairs of highest dimension and still have a sequence of removal of free pairs from the complex.

The following proposition explains which conditions on the inhibitor set (denoted $W$) need to be matched in order to guarantee the thinness of the result of a thinning constrained by $W$.

\begin{proposition}
\label{prop::thin_with_inhibit}
Let $X$ be an $n$-complex with $n>0$, let $S$ be a collapse of $X$ such that $\dim(S) \leq (n-1)$ and let $\subcomplex{W}{S}$. Let $Y$ be a collapse of $X$ such that $W \subseteq Y$ and such that there are no free pairs in $Y$ included in $Y \backslash W$. Then, the dimension of $Y$ is inferior or equal to $(n-1)$.
\end{proposition}

\begin{proof}
In the following, we show that if there exists an $n$-face in $Y$, then it belongs to a free pair for $Y$ that is included in $Y \backslash W$, a contradiction with the hypothesis of the proposition.

Let $C = ((a_1, b_1),...,(a_k, b_k))$ be a sequence of removal of free faces which allows to obtain $S$ from $X$: for all $i \in [1;k], (a_i, b_i)$ is free for $X \backslash \{a_1, b_1, ..., a_{i-1}, b_{i-1}\}$ and $S = X \backslash \{a_1, b_1, ..., a_k, b_k\}$. 
Let $C'$ be the sequence $C$ restrained only to free pairs containing an $n$-face: $C' = ((f_1, g_1),...,(f_h, g_h))$. A consequence of lemma \ref{lem::remove_small_pairs} is that, for all $i \in [1;h], (f_i, g_i)$ is free for $X \backslash \{f_1, g_1, ..., f_{i-1}, g_{i-1}\}$ and $(X \backslash \{f_1, g_1, ..., f_k, g_k\})$ is a collapse of $X$.

Any $n$-face $c \in Y$ is such that $c \in X$ and $c \notin S$, and therefore there exists $j \in [1;h]$ such that $c = f_j$. Without loss of generality, let $j$ be the smallest integer such that $f_j \in Y$: for all $k \in [1;j-1]$, $f_k \notin Y$ and $g_k \notin Y$. As previously said, $(f_j, g_j)$ is free for $(X \backslash \{f_1, g_1, ..., f_{j-1}, g_{j-1}\})$: $g_j$ is included in only one face of $(X \backslash \{f_1, g_1, ..., f_{j-1}, g_{j-1}\})$, and this face is $f_j$.

As $\{f_1, g_1, ..., f_{j-1}, g_{j-1}\} \cap Y = \emptyset$ and that $f_j \in Y$, then $g_j$ is included in only one face of $Y$, and that face is $f_j$. Consequently, the pair $(f_j, g_j)$ is free for $Y$. Moreover, the pair $(f_j, g_j)$ belongs to the sequence $C$, therefore $f_j \notin S$ and $g_j \notin S$. As $W \subseteq S$, the pair $(f_j, g_j)$ is included in $Y \backslash W$.$\finproof$
\end{proof}

\section{A parallel directional thinning based on cubical complex}

\subsection{Removing free pairs in parallel}
In the cubical complex framework, parallel removal of simple pairs can be easily achieved when following simple rules that we will give now. First, we need to define the \myem{direction} and the \myem{orientation} of a free face.

Let  $f \in \allfaces{n}$, the \myem{center of $f$} is the center of mass of the points in $f$, that is, $c_f= \frac{1}{\Card{f}}\sum_{a \in f}{a}$. The center of $f$ is an element of $[\Zhalf]^n$, where $\Zhalf$ denotes the set of half integers.
Let $\complex{n}{X}$, let $(f,g)$ be a free pair for $X$, and let $c_f$ and $c_g$ be the respective centers of the faces $f$ and $g$. We denote by $V(f,g)$ the vector $(c_f -c_g)$ of $[\Zhalf]^n$.

We define a surjective function $\type{}: \allfaces{n} \times \allfaces{n} \rightarrow \dinterval{0}{n-1}$ such that, for all free pairs $(f,g)$ and $(i,j)$ for $X$, $\type{f,g} = \type{i,j}$ if and only if $V(f,g)$ and $V(i,j)$ are collinear (we don't bother defining $\type{}$ for non free pairs as it won't be useful in this case). The number $\type{f,g}$ is called the \myem{direction} of the free pair $(f,g)$. Let $(f,g)$ be a free pair, the vector $V(f,g)$ has only one non-null coordinate: the pair $(f,g)$ has a \myem{positive orientation}, and we write $\orient{f,g}=1$, if the non-null coordinate of $V(f,g)$ is positive; otherwise $(f,g)$ has a \myem{negative orientation}, and we write $\orient{f,g}=0$.
On Fig.~\ref{fig:parallel_collapse}, the free pair $(a,c)$ and the free pair $(d,e)$ have different directions; the free pairs $(a,b)$ and $(d,e)$ have the same direction, but opposite orientations.

Now, we give a property of collapse which brings a necessary and sufficient condition for removing two free pairs of faces in parallel from a complex, while preserving topology (see Fig.~\ref{fig:parallel_collapse}c).

\begin{proposition}
\label{prop:collpar2}
Let $\complex{n}{X}$, and let $(f,g)$ and $(k,\ell)$ be two distinct free pairs for $X$. The complex $X$ collapses onto $X \setminus \{f,g,k,\ell\}$ if and only if $f \neq k$. 
\end{proposition}

\begin{proof}
If $f=k$, then it is plain that $(k,\ell)$ is not a free pair for $Y=X \setminus \{f,g\}$ as $k=f \notin Y$. Also, $(f,g)$ is not free for $X \setminus \{k,\ell\}$. If $f\neq k$, then we have $g \neq \ell$, $\contstrict{g} \inter X = \{f\}$ ($g$ is free for $X$) and $\contstrict{\ell} \inter X = \{k\}$ ($\ell$ is free for $X$). Thus, we have $\contstrict{\ell} \inter Y = \{k\}$ as $\ell\neq g$ and $k\neq f$. Therefore, $(k,\ell)$ is a free pair for $Y$. $\finproof$
\end{proof}

From Prop.~\ref{prop:collpar2}, the following corollary is immediate.

\begin{corollary}
\label{cor:parallel_removal}
Let $\complex{n}{X}$, and let $(f_1,g_1) \ldots (f_m,g_m)$ be $m$ distinct free pairs for $X$ such that, for all $a,b \in
\dinterval{1}{m}$ (with $a \neq b$), $f_a \neq f_b$. The complex $X$ collapses onto $X \setminus \{f_1,g_1 \ldots f_m,g_m\}$.
\end{corollary}

Considering two distinct free pairs $(f,g)$ and $(i,j)$ for $\complex{n}{X}$ such that $\type{f,g}=\type{i,j}$ and $\orient{f,g} = \orient{i,j}$, we have $f \neq i$.  From this observation and Cor.~\ref{cor:parallel_removal}, we deduce the following property.

\begin{corollary}
\label{cor:parallel_removal_v2}
Let $\complex{n}{X}$, and let $(f_1,g_1) \ldots (f_m,g_m)$ be $m$ distinct free pairs for $X$ having all the same direction and the same orientation. The complex $X$ collapses onto $X \setminus \{f_1,g_1 \ldots f_m,g_m\}$. 
\end{corollary}

\subsection{A directional parallel thinning algorithm}

We say that a $d$-face of $X$ is a \myem{border face} if it contains a free $(d-1)$-face. Define $\border{X}$ as the set of all border faces of $X$. We are now ready to introduce a directional parallel thinning algorithm (Alg.~\ref{algo::pardircollapse_v0}).

\begin{algorithm}[h!]
\label{algo::pardircollapse_v0}
\SetKwFor{ForAll}{for all}{do}{end}
\caption[ParDirCollapse]{$ParDirCollapse(X,W,\ell)$}
\KwData{A cubical complex $\complex{n}{X}$, a subcomplex
  $\subcomplex{W}{X}$ which represents faces of $X$ which should not
  be removed, and $\ell \in \Nset$, the number of layers of free faces
  which should be removed from $X$} 
\KwResult{A cubical complex}
\lnl{pc:01}\While{there exists free faces in $X \setminus W$ and $\ell > 0$}{
\lnl{pc:02}  $L=\border{X}^-$\;
\lnl{pc:03}  \For{$t=0 \rightarrow n-1$}{
\lnl{pc:04}    \For{$s=0 \rightarrow 1$}{
\lnl{pc:05}      \For{$d=n \rightarrow 1$}{
\lnl{pc:06}        $E=\{(f,g)$ free for $X \suchthat g \notin W,$\\
                      $\;$~~~~~~~$\type{f,g}=t,$ $\orient{f,g}=s,$ $\mydim{f}=d\}$\;
\lnl{pc:07}        $G=\{(f,g) \in E \suchthat f \in L$ and $g \in L\}$\;
\lnl{pc:08}        $X=X \setminus G$\;
                 }
               }
             }
\lnl{pc:09}  $l = l-1$\;
	   }
\lnl{pc:10} \Return $X$\;
\end{algorithm}

Intuitively, we want the algorithm to remove free faces \quotes{layer by layer}: we don't want, after a single execution ($\ell =1$), to have unequal thinning of the input complex. Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex: this is why we introduce, on line \ref{pc:02} the set $L$, and that we remove only faces located in $L$ on line \ref{pc:08}. The sets $E$ (line \ref{pc:06}) and $G$ (line \ref{pc:07}) allows to remove whole sets of free faces in parallel from $X$, thanks to the direction and orientation of faces previously defined. A detailed view of each step of the algorithm is shown on Fig.~\ref{fig:step_thinning}.

Different definitions of the orientation and direction can be given, corresponding to different order of removal of free faces in the complex. These changes lead to different results, but arbitrary choices on the order of removal of free pairs must be made in order to obtain, at the end, a thin skeleton (no more $n$-faces when working in the $n$-dimension). Once orientation and direction have been defined, the results of the algorithm are uniquely defined.

Algorithm~\ref{algo::pardircollapse_v0} may be easily implemented to run in linear time complexity (proportionally to the number of faces of the complex). Indeed, checking if a face is free or not may be easily done in constant time. Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the faces of $f$ and the cofaces of $g$ in order to find new free faces, as other faces' status won't change (the implemented algorithm contains these optimizations).

\section{Aspect preservation during thinning: a parameter-free method}
\label{sec::aspect_no_param}
Algorithm \ref{algo::pardircollapse_v0} does not necessarily preserves "geometrical information" of the original object in the resulting skeleton. For example, if the input was a filled human shape (one connected component, no tunnel nor cavity) and that $W=\emptyset$ and $l=+\inf$, then the result of Alg.~\ref{algo::pardircollapse_v0} would be a simple vertex. However, when one wants to preserve "geometrical information" from the original shape in the skeleton, then one expects to obtain a "stickman" from a human shape. However, it is important to not retain "too much information" during thinning in order to avoid noisy branches in the skeleton.

In the following, we will see a new method in the cubical complex, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory visual properties. This will be achieved by adding, in the set $W$ of Alg.~\ref{algo::pardircollapse_v0} some faces of the original object.

\subsection{The lifespan of a face}
In the following, we define new notions in the cubical complex. The first one we present is the \myem{death date} of a face.

\begin{definition}
Let $f \in \complex{n}{X}$, the \myem{death date of $f$ in $X$}, denoted by $\death{f}{X}$, is the smallest integer $d$ such that $f \notin ParDirCollapse(X, \emptyset, d)$. 
\end{definition}

The death date of a face indicates how many layers of free faces should be removed from a complex $X$, using alg.~\ref{algo::pardircollapse_v0}, before removing completely the face from $X$. We now define the \myem{birth date} of a face:

\begin{definition}
Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the minimum between the smallest integer $b$ such that $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, and $\death{f}{X}$. 
\end{definition}

The birth date indicates how many layers of free faces must be removed from $X$ with Alg.\ref{algo::pardircollapse_v0} before transforming $f$ into a facet of $X$ (we consider a face "lives" when it is a facet).
Finally, we can define the \myem{lifespan} of a face :

\begin{definition}
Let $f \in \complex{n}{X}$, the lifespan of $f$ in $X$ is the integer

\begin{math}
\lifespan{f}{X} = \left\{
\begin{array}{ll}
+\infty & \text{if $\death{f}{X}=+\infty$}\\
\death{f}{X} - \birth{f}{X} & \text{otherwise}
\end{array}
\right.
\end{math}
\end{definition}

\begin{figure}[ht]
\begin{center}
\begin{tabular}{ccc}
	\myincludegraphics{0.3\textwidth}{EPS/mario_life1}{.pdf} &
	\myincludegraphics{0.3\textwidth}{EPS/mario_life2}{.pdf} &
	\myincludegraphics{0.3\textwidth}{EPS/mario_life3}{.pdf} \\
	a) A complex $X$ & b) $ParDirCollapse(X,\emptyset,1)$ & c) $ParDirCollapse(X,\emptyset,2)$\\
	\myincludegraphics{0.3\textwidth}{EPS/mario_life4}{.pdf} &
	\myincludegraphics{0.3\textwidth}{EPS/mario_life5}{.pdf} &
	\myincludegraphics{0.3\textwidth}{EPS/mario_life6}{.pdf} \\
	d) $ParDirCollapse(X,\emptyset,3)$ & e) $ParDirCollapse(X,\emptyset,4)$ & f) $ParDirCollapse(X,\emptyset,5)$\\
\end{tabular}
\caption[Computing the lifespan of a face]{\label{fig:lifespan_exemple} \textbf{Computing the lifespan of a face - }Let $e$ be the edge of the complex $X$, pointed out by a red arrow on \textbf{a}. Two iterations of $ParDirCollapse$ algorithm are necessary before $e$ becomes a facet, as depicted on \textbf{c}. Therefore, $\birth{e}{X}=2$. Five iterations of $ParDirCollapse$ algorithm are necessary for removing $e$ from the complex, as depicted in \textbf{f}. Therefore, $\death{e}{X}=5$. We have $\lifespan{e}{X}=3$.}
\end{center}
\end{figure}

The three parameters previously defined are dependant on the order of direction and orientation chosen for algorithm $ParDirCollapse$.

The lifespan of a face $f$ of $X$ indicates how many "rounds" this face "survives" as a facet in $X$, when removing free faces with algorithm \ref{algo::pardircollapse_v0}. The lifespan, the death and the birth, as defined here, are dependant of Alg.~\ref{algo::pardircollapse_v0}, used for performing the thinning. It is of course possible to use another algorithm for performing the thinning, leading to other values of birth, death and lifespan. It is recommended, in order to have "comparable" values, to compute these three parameters with the same thinning technique.

Algorithm \ref{algo::lifespan} computes the lifespan of all faces of a complex. The algorithm is not linear in time, however, a linear implementation of such algorithm exists and is presented in the appendix. On Fig.~\ref{fig:lifespan_exemple}, we show a sequence of collapse allowing to compute the lifespan of an edge.
%Todo lien avec annexe

\begin{algorithm}[h!]
\label{algo::lifespan}
\caption[Lifespan]{$Lifespan(X)$}
\SetKwFor{ForAll}{for all}{do}{end}
\KwData{A cubical complex $\complex{n}{X}$}
\KwResult{The map giving the lifespan of all faces of $X$}
\ForAll{$f \in X$}{
  $Death(f)=+\infty$\;
  \If{$f \in \princ{X}$}{
    $Birth(f)=0$\;
  }
  \Else{
    $Birth(f)=+\infty$\;
  }
}
$Y=X; l=0$\;
\While{there exists free faces for $Y$}{
  $Y=ParDirCollapse(Y,\emptyset,1)$\;
  $l=l+1$\;
  \ForAll{$f \in \princ{Y}$ such that $Birth(f)=+\infty$}{
  	$Birth(f)=l$\;
  }
  \ForAll{$f \notin Y$ such that $Death(f)=+\infty$ and $f \in X$}{
  	$Death(f)=l$\;
  }
}
\ForAll{$f \in X$}{
  $Birth(f)=\min(Birth(f), Death(f))$\;
  \If{$Death(f)=+\infty$}{
    $Lifespan(f)=+\infty$\;
  }
  \Else{
    $Lifespan(f)=Death(f) - Birth(f)$\;
  }
}
\Return $Lifespan$\;
\end{algorithm}

The lifespan is a good indicator of how important a face can be in an object. Typically, higher the lifespan is, and more representative of an object's visual feature the face is. The lifespan, also called \myem{saliency}, was used in \cite{Liu2009} (under the name "medial persistence") in order to propose a thinning algorithm in cubical complexes based on two parameters.

\subsection{Distance map and opening function}
\label{subsec::distmap}
In addition to the lifespan of a face, the proposed homotopic thinning method will use information on distance between faces in order to decide if a face should be kept safe from deletion. We define hereafter the various notions needed for this, based on distance in the DT framework.

Two points $x,y \in \Zset^n$ are \myem{$1$-neighbours} if the Euclidean distance between $x$ and $y$ is equal or inferior to $1$ (also called direct neighbours). A \myem{$1$-path from $x$ to $y$} is a sequence $\mathcal{C}=(z_0,...,z_k)$ of points of $\Zset^n$ such that $z_0=x$, $z_k=y$, and for all $j \in [1;k]$, $z_j$ and $z_{j-1}$ are $1$-neighbours. The length of $\mathcal{C}$ is $k$.

We set $d_1(x,y)$ as the length of the shortest $1$-path from $x$ to $y$. Let $S \subset \Zset^n$, we set $d_1(x,S) = \displaystyle \min_{y \in S} d_1(x,y)$. The \myem{$1$-ball of radius $r$ centered on $x$} is the set $\OneBall{r}(x)=\{y \in \Zset^n | d_1(x,y) < r\}$. Remark that $d_1$ is indeed the so-called $4$-distance in the 2d DT framework, and the $6$-distance in the 3d DT framework.
Given $X \subset \Zset^n$, the \myem{maximal $1$-ball of $X$ centered on $x$} is the set $\MaxOneBall{X}(x) = \OneBall{d_1(x,\Compl{X})}(x)$.

We set, for all $x \in X$, $\dd_1(x,\Compl{X}) = \displaystyle \max_{x \in \MaxOneBall{X}(y)} d_1(y,\Compl{X})$: this value indicates the radius of the largest maximal $1$-ball contained in $X$ and containing $x$. If $x \in \Compl{X}$, we set $\dd_1(x,\Compl{X}) = 0$. The map $\dd_1$ is known as the opening function (based on the $1$-distance): it allows to compute efficiently results of openings by balls of various radius, and gives information on the local thickness of an object on each of its points. We show some examples of the opening function on Fig.~\ref{fig::globalopening} and \ref{fig::detailopening}.

\end{document}

### 2012-09-07 bidir_pathtrace_skeleton.tex
Le tex principal devrait compiler

@@
 
+\def\axisloc#1{\mathcal{LOC}_{#1}}
+\def\axislocbis#1#2{\mathcal{LOC}_{#1}^{>#2}}
+\def\axislc#1{\mathcal{LC}_{#1}}
+\newcommand{\centerness}{Decenter}
+
+\def\OneBall#1{\mathbb{B}^{1}_{#1}}
+\def\MaxOneBall#1{\hat{\OneBall{#1}}}
+
+\def\dd{\omega}
+\def\DD{\Omega}
+
 \def\mydef#1{{\em #1}}
@@
 \def\myvec#1{\vec{#1}}
+\def\quotes#1{``#1''}
+
+\def\birth#1#2{Birth_{#2}(#1)}
+\def\death#1#2{Death_{#2}(#1)}
+\def\lifespan#1#2{Lifespan_{#2}(#1)}
 
@@
 
-\begin{figure}[ht]
-\begin{center}
-\begin{tabular}{ccc}
-	\myincludegraphics{0.3\textwidth}{EPS/mario_life1}{.pdf} &
-	\myincludegraphics{0.3\textwidth}{EPS/mario_life2}{.pdf} &
-	\myincludegraphics{0.3\textwidth}{EPS/mario_life3}{.pdf} \\
-	a) A complex $X$ & b) $ParDirCollapse(X,\emptyset,1)$ & c) $ParDirCollapse(X,\emptyset,2)$\\
-	\myincludegraphics{0.3\textwidth}{EPS/mario_life4}{.pdf} &
-	\myincludegraphics{0.3\textwidth}{EPS/mario_life5}{.pdf} &
-	\myincludegraphics{0.3\textwidth}{EPS/mario_life6}{.pdf} \\
-	d) $ParDirCollapse(X,\emptyset,3)$ & e) $ParDirCollapse(X,\emptyset,4)$ & f) $ParDirCollapse(X,\emptyset,5)$\\
-\end{tabular}
-\caption[Computing the lifespan of a face]{\label{fig:lifespan_exemple} \textbf{Computing the lifespan of a face - }Let $e$ be the edge of the complex $X$, pointed out by a red arrow on \textbf{a}.
-Two iterations of $ParDirCollapse$ algorithm are necessary before $e$ becomes a facet, as depicted on \textbf{c}.
-Therefore, $\birth{e}{X}=2$.
-Five iterations of $ParDirCollapse$ algorithm are necessary for removing $e$ from the complex, as depicted in \textbf{f}.
-Therefore, $\death{e}{X}=5$.
-We have $\lifespan{e}{X}=3$.}
-\end{center}
-\end{figure}
-
 The three parameters previously defined are dependant on the order of direction and orientation chosen for algorithm $ParDirCollapse$.
@@
 
+Given $\complex{n}{X}$, the value of $\dd_1(x,\Compl{X})$ of every $x \in X$ can be computed by performing successive dilations of values of the map $d_1$.
+The algorithm presented in Alg.\ref{algo::naive_opening} is naive, and a more efficient implementation (linear in time depending on the image's size) is discussed in the appendix (see Sec.~\ref{sec::better_open}, p.~\pageref{sec::better_open}).
+
+In order to extend $d_1$ and $\dd_1$ to the cubical complex framework, let us introduce the map $\Phi^{-1}$, inverse of the bijective map $\Phi$ defined in Sec.~\ref{sec::binary_to_cubic}.
+It is used to project any $n$-face of $\allfaces{n}$ into $\Zset^n$.
+We indifferently use $\Phi^{-1}$ as a map from $\subfaces{n}{n}$ to $\Zset^n$, and as a map from $\mathcal{P}(\subfaces{n}{n})$ to $\mathcal{P}(\Zset^n)$.
+
+Given $\complex{n}{X}$, we set $X^v = \Phi^{-1}(X \inter \subfaces{n}{n})$: the set $X^v$ is a subset of $\Zset^n$.
+We define the map $\tilde{D_1}(X): \allfaces{n} \rightarrow \Nset$ as an extension of $d_1$ to the cubical complex framework: for all $f \in \allfaces{n}$,
+
+\bigskip
+\begin{math}
+\tilde{D_1}(X)(f) = \left\{
+\begin{array}{ll}
+d_1(\Phi^{-1}(f),\Compl{X^v}) & \text{if $f$ is an $n$-face}\\
+\displaystyle \max_{g \in \contstrict{f}} \tilde{D_1}(X)(g) & \text{else}
+\end{array}
+\right.
+\end{math}
+
+The same way, we define $\tilde{\DD_1}(X): \allfaces{n} \rightarrow \Nset$ as an extension of $\dd_1$ to the cubical complex framework.
+
+\subsection{Parameter-free thinning based on the lifespan, opening function and decenterness}
+Thanks to these notions, we can now define sets of faces that will help preserve the visual aspect of an object during thinning.
+First, let us define the decenterness of a complex as the map $\centerness(X) = \tilde{\DD_1}(X) - \tilde{D_1}(X)$.
+For each face of a complex $X$, the decenterness value of this face gives an information on how well a face is centered inside a visual feature of the object: the lower this value is, the better centered the face is.
+On Fig.~\ref{fig::globalopening}, we give an example of the decenterness map of a shape, where the highest values are represented by the darkest colours.
+
+Faces relevant of the visual aspect of a complex must have a high lifespan ("survive" long to the homotopic thinning process) and a low decenterness (is centered in the object).
+In a $3$-dimensional complex (resp.
+a $2$-dimensional complex), squares (resp.
+lines) whose lifespan is higher than the decenterness will be chosen as relevant of the surfacic (resp.
+curvilinear) parts of the complex.
+
+\begin{definition}
+Given $\complex{n}{X}$, the \myem{$k$-LC axis} (stands for "Lifespan Centerness") of $X$ is the set
+
+\begin{center}
+$\axislc{k}(X) = \closure{\{f \in X$ $|$ $\dim(f)=k$ and $\lifespan{f}{X} > \centerness(X)(f)\}}$.
+\end{center}
+
+\end{definition}
+
+The LC-axis fails in selecting a good set of curves relevant of the curvilinear parts of a three-dimensional complex.
+Indeed, given a $3$-complex $X$, the set $\axislc{1}(X)$ contains generally too many curves.
+This happens because the algorithm $ParDirCollapse$ "takes more time" to eliminate lines than to eliminate squares in a $3$-complex.
+Consequently, lines of a $3$-complex tend to have a high lifespan even though they are not representative of any curvilinear part of the complex.
+The thicker is the input object, and the more important is the phenomenon.
+
+Lines relevant of the curvilinear parts of a $3$-complex have a high lifespan and a low decenterness, especially in the thick parts of the complex.
+Lines whose lifespan is higher than the decenterness added to the local thickness of the complex will be relevant of the curvilinear parts of the complex.
+
+\begin{definition}
+Given $\complex{n}{X}$, the \myem{$k$-LOC axis} (stands for "Lifespan Opening Centerness") of $X$ is the set
+
+\begin{center}
+$\axisloc{k}(X) = \closure{\{f \in X$ $|$ $\dim(f)=k$ and $\lifespan{f}{X} > \tilde{\DD_1}(X)(f) + \centerness(X)(f)\}}$.
+\end{center}
+\end{definition}
+
+The various sets previously defined represent faces which should be kept safe from deletion during homotopic thinning of a complex in order to obtain a pruned skeleton containing visual features from the original object.
+The set $\axislc{1}$ of a $2$-complex represents curvilinear parts to keep in this complex, the set $\axisloc{1}$ of a $3$-complex represents curvilinear parts to keep in this complex, and the set $\axislc{2}$ of a $3$-complex represents surfacic parts to keep in this complex.
+
+\begin{lemma}
+\label{lem::axis_subset_facet}
+Given $\complex{n}{X}$, for every $k<n$, $\axislc{k}(X) \subseteq CollapseFacet(X, \emptyset)$ and $\axisloc{k}(X) \subseteq CollapseFacet(X, \emptyset)$.
+\end{lemma}
+
+Lemma \ref{lem::axis_subset_facet} is straightforward once observed that $CollapseFacet(X, \emptyset)$ is indeed the set of $k$-faces of $X$ ($k<n$) which have a strictly positive lifespan in $X$. 
+Based on these sets, we propose three algorithms: one for computing a 1d skeleton from a bi-dimensional complex, one for computing a 2d skeleton (which can contain 1d parts) from a three-dimensional complex, and one for computing a curvilinear skeletons from a three-dimensional complex.
+This last algorithm should produce one-dimensional results (a set of lines) however, the output can be a two-dimensional complex.
+
+\begin{algorithm}[ht]
+\label{algo::3d_input_curv_skel}
+\SetKwFor{ForAll}{for all}{do}{end}
+\caption[CurvilinearSkeleton]{$CurvilinearSkeleton(X)$}
+\KwData{A cubical complex $\complex{3}{X}$}
+\KwResult{A cubical complex $\complex{3}{Y}$ such that $\dim{Y} \leq 2$}
+$W=\axisloc{1}(X)$\;
+$ParDirCollapse(X,W,+\infty)$\;
+\Return $X$\;
+\end{algorithm}
+
+Algorithm \ref{algo::3d_input_curv_skel} produces as output a 2d or 1d complex, result of the homotopic thinning of a three-dimensional complex.
+The skeleton produced by this algorithm should be a set of lines describing the shape of the input, however, it may contain squares representative of the topology of the initial object (if the input complex contains a cavity, the skeleton will contain surfacic parts).
+
+Despite the use of constraint sets (sets of faces which should not be removed during thinning) in the three algorithms, they produce thin results.
+-> On peut simplement modifier algo pour y incorporer directement les elements a conserver pdt la skel...
+
 \end{document}

### 2012-09-07 pathtracint_application_skeleton.tex
Add the tex file that describe the skeleton applied to path tracing

\documentclass[final,envcountsame]{llncs}

%\usepackage{graphicx} %Pouvoir insérer des images
\usepackage{epsfig}
\usepackage{amsmath} %Permet d'utiliser une jolie panoplie de symboles mathématiques
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{amsfonts}

\def\mydef#1{{\em #1}}
\def\myem#1{{\em #1}}
\def\myvec#1{\vec{#1}}

\def\Nset{\mathbb{N}}
\def\Zset{\Z}
\def\Rset{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\def\allfaces#1{\mathbb{F}^{#1}}
\def\subfaces#1#2{\allfaces{#1}_{#2}}
\def\subcomplex#1#2{#1 \preceq #2}
\def\complex#1#2{\subcomplex{#2}{\allfaces{#1}}}
\def\cell#1{\hat{#1}}
\def\cellstrict#1{\hat{#1}^*}
\def\cont#1{\check{#1}}
\def\contstrict#1{\check{#1}^*}
\def\closure#1{#1^-}
\def\princ#1{#1^+}
\def\detach#1#2{#2 \oslash #1}
\def\attach#1#2{\mbox{\it{}Attach\/}(\cell{#1},#2)}
\def\mydim#1{\dim(#1)}
\def\type#1{\mbox{\it{}Dir\/}(#1)}
\def\orient#1{\mbox{\it{}Orient\/}(#1)}
\def\border#1{\mbox{\it{}CBorder\/}(#1)}
\def\suchthat{\;|\;}
\def\inter{\cap}
\def\union{\cup}
\def\dinterval#1#2{\{#1,\ldots,#2\}}
\newcommand{\Compl}[1]{\overline{#1}}
\def\finproof{\square}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\newcommand{\Card}[1]{|#1|}

\def\figpath#1{../../Images/EPS/#1}

\begin{document}

\section{Application to path tracing}

\subsection{The path tracing algorithm}

Path tracing is a global illumination algorithm that is able to render photo-realistic images of a scene viewed by a camera. The idea behind the algorithm is very simple: for each pixel of the image, throw rays of light through the pixel that will bounce on the scene to accumulate lighting exchanges. It is based on physical laws of radiometry.

\subsubsection{Basic definitions of radiometry}

Let $S$ be a set of surfaces of $\Rset^3$ that don't overlap ($\forall x \in \Rset^3$ there exists at most one surface of $S$ that contains $x$) and $S_L \subset S$ a set of surface lights. $S$ is called the scene.

The basic radiometric quantity is the \myem{radiance} $L(x \rightarrow \Theta)$ that express the quantity of light energy produce by the surface point $x \in \Rset^3$ towards direction $\Theta \in \Omega_x$ per unit of time per unit area per unit solid angle (expressed in $W.m^{-2}.sr^{-1}$).
This is the quantity that the eye actually "sees" and that must be computed.

% image radiance

We can define another quantity related to radiance called \myem{incoming radiance} $L(x \leftarrow \Phi)$ that represents the radiance that reach $x$ from direction $\Phi$.

\begin{property}

The radiance remains constant along straight lines in the void.

\end{property}

That property allows us to write the incoming radiance in terms of radiance.

\begin{equation*}
L(x \leftarrow \Phi) = L(r(x, \Phi) \rightarrow -\Phi)
\end{equation*}

$r(x, \Phi)$ is the surface point seen by $x$ in the direction $\Phi$. It can be defined formally by:

\begin{align*}
r(x, \Phi) &= x + t_{min}\Phi \\
t_{min} &= min \lbrace t \in (0, +\infty]~ |~ x + t\Phi \in S \rbrace
\end{align*}

If $t_{min} = \infty$ then $r(x, \Phi)$ is undefined and $L(x \leftarrow \Phi) = 0$. We call $r$ the \myem{raytrace} function. In practice it is implemented by throwing a ray in the scene and looking for the first intersection.

A third quantity related to radiance is \myem{emitted radiance} $L_e(x \rightarrow \Theta)$. It is not null if $x$ is a point of a light source ($x \in S_L$). In practice the emitted radiance is given with the set $S_L$ as an input of the algorithm (in can be constant across each of the lights or given by a texture).

The radiance is the solution of an equation called the Rendering Equation:

\begin{equation}
L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi) d\omega_\Phi
\end{equation}

It's actually this equation that a path tracer try to resolve. Let $O$ be the origine of the camera and $I$ the image we want to compute (a rectangle embedded in $\Rset^3$). Then the path tracing algorithm compute an approximation of the radiances $L(O \leftarrow \vec{OP})$, $\forall P \in I$.

\subsubsection{Solving the rendering equation}

Parler rapidement de l'échantillonage et donner l'estimateur de monte carlo

\subsubsection{The algorithm}

Donner l'algorithme en pseudo-code

\subsection{Skeleton based importance sampling}

Décrire la stratégie d'échantillonage basée sur le squelette

\subsection{Multiple importance sampling}

Décrire comment combiner les stratégies brdf + skeleton

\subsection{Coarse irradiance estimation using curvilinear skeleton}

Décrire les heuristiques

\end{document}

### 2012-09-08 bidir_pathtrace_skeleton.tex
Quelques corrections sur l'article, vire quelques parties inutiles.

@@
 
-In the 3d cubical complex framework, objects are no more made of voxel, but of cubes, squares, lines and vertices $\mathbb{R}$. 
+In the 3d voxel framework, objects are mae of voxels. In the 3d cubical complex framework, objects made of voxel, but of cubes, squares, lines and vertices. 
 Let $\Z$ be the set of integers, we consider the family of sets $\mathbb{F}^1_0$ and $\mathbb{F}^1_1$, such that $\mathbb{F}^1_0 = \{\{a\} \suchthat a \in \Z\}$ and $\mathbb{F}^1_1 = \{ \{a, a+1\} \suchthat a \in \Z\}$.
@@
 
-Informally, in 3d, a set of face is a cubical complex if, for each cube of the complex, the six squares (sides) of the cube also belong to the complex, for each square of the complex, the four edges (sides) of the square also belong to the complex, and for each edge of the complex, the two vertices (extremities) of the edge also belong to the complex.
+Informally, in 3d, a set of faces $X$ is each side (square) of a cube belonging to $X$ also belong to $X$, each edge of a square belonging to $X$ also belong to $X$, and each vertex of an edge belonging to $X$ also belong to $X$.
 
@@
 If $\mydim{X} = d$, then we say that $X$ is a $d$-complex.
-The notions of purity and dimensions can be trivially extended to sets of faces.
-
-In $\allfaces{n}$, a complex $X$ is \myem{thin} if $\mydim{X}<n$. In $\allfaces{3}$, a complex is thin if it contains no cube.
+In $\allfaces{n}$, a complex $X$ is \myem{thin} if $\mydim{X}<n$; in $\allfaces{3}$, a complex is thin if it contains no cube.
 
@@
 
-Traditionally, a binary image is defined as a finite subset of $\Zset^n$ (with $n=2$ or $n=3$).
-Given $S \subseteq \Zset^n$, the object voxels are the elements of $S$.
-This kind of image is the most common one in the field of image processing so, in order to work in cubical complex framework, we need to find a way to transpose a binary image to cubical complex framework.
+Traditionally, a voxel image is defined as a finite subset of $\Zset^n$.
+This kind of image is the most common one in the field of image processing and we give now a simple way to transpose a binary image to the cubical complex framework.
 
@@
 We can extend the map $\Phi$ to sets: $\Phi(S)=\{\Phi(x) | x \in S\}$.
-Given a binary image $S$, we associate to it the cubical complex $\closure{\Phi(S)}$ (see Fig.~\ref{fig::voxel_to_complex}).
-
-In the following, most of the objects we consider were indeed binary images which were then transposed into cubical complex framework: this is why most of the two-dimensional complexes we show are pure 2-complexes, and most of the three-dimensional complexes we show are pure 3-complexes.
+Given a binary image $S$, we associate to it the cubical complex $\closure{\Phi(S)}$.
 
@@
 
-In other terms, $(f,g)$ is a free pair for $X$ whenever $\contstrict{g} \inter X = \{f\}$ ($g$ is included only in $f$).
+In other terms, $(f,g)$ is a free pair for $X$ whenever $\contstrict{g} \inter X = \{f\}$.
 It can be easily seen that if $(f,g)$ is a free pair for a complex $X$ and $\dim(f)=m$, then $f$ is a facet and $\dim(g) = m-1$.
@@
 
-We insist here on the fact that, if $Y$ is a collapse of a complex $X$, then $Y$ is a complex.
-
-Let us now introduce some elements that will serve later for proving the thinness of our skeletons.
+If $Y$ is a collapse of a complex $X$, then $Y$ is a complex. 
+We now introduce some elements that will serve for proving the thinness of our skeletons.
 Let $f_0,f_\ell$ be two $n$-faces of $\allfaces{n}$ (with $\ell$ being even). An \myem{$(n-1)$-path from $f_0$ to $f_\ell$} is a sequence $\pi =
@@
 Therefore, it is possible to perform collapse on a complex until no more $3$-faces (volumes) can be found (until it is thin).
-
-As in the DT framework, it is sometimes necessary to perform collapse in the cubical complex framework while preserving some faces safe from deletion: these faces are the so-called inhibitor set.
-When using an inhibitor set during collapse, the guarantee of having a thin result does no more hold.
-As illustrated on figure ffjkz, the inhibitor set can be thin and a result of the constrained thinning can still not be thin.
-However, under some conditions, it is possible to use an inhibitor set while collapsing, and still have the guarantee of getting a thin result. In order to prove this, let us state the following:
-
-\begin{lemma}
-\label{lem::remove_small_pairs}
-Let $\complex{n}{X}$, let $(f_1,g_1)$ be a free pair of $X$, and $(f_2,g_2)$ be a free pair of $(X \backslash \{f_1,g_1\})$. If $\dim(f_2) > \dim(f_1)$, then $(f_2,g_2)$ is free for $X$. 
-\end{lemma}
-
-\begin{proof} If $(f_2,g_2)$ is a free pair of $(X \backslash \{f_1,g_1\})$, then $g_2$ is included in only one face of $(X \backslash \{f_1,g_1\})$, which is $f_2$.
-As $(\dim(g_2) = \dim(f_2) -1)$ and $(\dim(f_1) \leq (\dim(f_2)-1))$, then $g_2 \nsubseteq f_1$ and $g_2 \nsubseteq g_1$.
-Therefore, $g_2$ is included in only one face of $X$, which is $f_2$: therefore, $(f_2,g_2)$ is free for $X$.$\finproof$
-\end{proof}
-
-This lemma implies that, when one has a sequence of removal of free pairs of faces from a complex, one can only keep the free pairs of highest dimension and still have a sequence of removal of free pairs from the complex.
-
-The following proposition explains which conditions on the inhibitor set (denoted $W$) need to be matched in order to guarantee the thinness of the result of a thinning constrained by $W$.
-
-\begin{proposition}
-\label{prop::thin_with_inhibit}
-Let $X$ be an $n$-complex with $n>0$, let $S$ be a collapse of $X$ such that $\dim(S) \leq (n-1)$ and let $\subcomplex{W}{S}$.
-Let $Y$ be a collapse of $X$ such that $W \subseteq Y$ and such that there are no free pairs in $Y$ included in $Y \backslash W$.
-Then, the dimension of $Y$ is inferior or equal to $(n-1)$.
-\end{proposition}
-
-\begin{proof}
-In the following, we show that if there exists an $n$-face in $Y$, then it belongs to a free pair for $Y$ that is included in $Y \backslash W$, a contradiction with the hypothesis of the proposition.
-
-Let $C = ((a_1, b_1),...,(a_k, b_k))$ be a sequence of removal of free faces which allows to obtain $S$ from $X$: for all $i \in [1;k], (a_i, b_i)$ is free for $X \backslash \{a_1, b_1, ..., a_{i-1}, b_{i-1}\}$ and $S = X \backslash \{a_1, b_1, ..., a_k, b_k\}$.
-Let $C'$ be the sequence $C$ restrained only to free pairs containing an $n$-face: $C' = ((f_1, g_1),...,(f_h, g_h))$.
-A consequence of lemma \ref{lem::remove_small_pairs} is that, for all $i \in [1;h], (f_i, g_i)$ is free for $X \backslash \{f_1, g_1, ..., f_{i-1}, g_{i-1}\}$ and $(X \backslash \{f_1, g_1, ..., f_k, g_k\})$ is a collapse of $X$.
-
-Any $n$-face $c \in Y$ is such that $c \in X$ and $c \notin S$, and therefore there exists $j \in [1;h]$ such that $c = f_j$.
-Without loss of generality, let $j$ be the smallest integer such that $f_j \in Y$: for all $k \in [1;j-1]$, $f_k \notin Y$ and $g_k \notin Y$.
-As previously said, $(f_j, g_j)$ is free for $(X \backslash \{f_1, g_1, ..., f_{j-1}, g_{j-1}\})$: $g_j$ is included in only one face of $(X \backslash \{f_1, g_1, ..., f_{j-1}, g_{j-1}\})$, and this face is $f_j$.
-
-As $\{f_1, g_1, ..., f_{j-1}, g_{j-1}\} \cap Y = \emptyset$ and that $f_j \in Y$, then $g_j$ is included in only one face of $Y$, and that face is $f_j$.
-Consequently, the pair $(f_j, g_j)$ is free for $Y$.
-Moreover, the pair $(f_j, g_j)$ belongs to the sequence $C$, therefore $f_j \notin S$ and $g_j \notin S$.
-As $W \subseteq S$, the pair $(f_j, g_j)$ is included in $Y \backslash W$.$\finproof$
-\end{proof}
 

### 2012-09-08 bidir_pathtrace_skeleton.tex
J'ai fait un include du fichier de Laurent dans le principal > il faut compiler le fichier principal.
Quelques corrections (encore d'autres à venir)

@@
 \def\orient#1{\mbox{\it{}Orient\/}(#1)}
-\def\border#1{\mbox{\it{}CBorder\/}(#1)}
+\def\border#1{\mbox{\it{}Border\/}(#1)}
 \def\suchthat{\;|\;}
@@
 \subsection{Removing free pairs in parallel}
-In the cubical complex framework, parallel removal of simple pairs can be easily achieved when following simple rules that we will give now.
+In the cubical complex framework, parallel removal of simple pairs can be easily achieved when following simple rules that we give now.
 First, we need to define the \myem{direction} and the \myem{orientation} of a free face.
@@
 Let $(f,g)$ be a free pair, the vector $V(f,g)$ has only one non-null coordinate: the pair $(f,g)$ has a \myem{positive orientation}, and we write $\orient{f,g}=1$, if the non-null coordinate of $V(f,g)$ is positive; otherwise $(f,g)$ has a \myem{negative orientation}, and we write $\orient{f,g}=0$.
-On Fig.~\ref{fig:parallel_collapse}, the free pair $(a,c)$ and the free pair $(d,e)$ have different directions; the free pairs $(a,b)$ and $(d,e)$ have the same direction, but opposite orientations.
+%On Fig.~\ref{fig:parallel_collapse}, the free pair $(a,c)$ and the free pair $(d,e)$ have different directions; the free pairs $(a,b)$ and $(d,e)$ have the same direction, but opposite orientations.
 
-Now, we give a property of collapse which brings a necessary and sufficient condition for removing two free pairs of faces in parallel from a complex, while preserving topology (see Fig.~\ref{fig:parallel_collapse}c).
+Now, we give a property of collapse which brings a necessary and sufficient condition for removing two free pairs of faces in parallel from a complex, while preserving topology.
 
@@
 
+Intuitively, we want the algorithm to remove free faces of a complex \quotes{layer by layer}: we don't want to have unequal thinning of the input complex.
+Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex.
 We say that a $d$-face of $X$ is a \myem{border face} if it contains a free $(d-1)$-face.
-Define $\border{X}$ as the set of all border faces of $X$.
-We are now ready to introduce a directional parallel thinning algorithm (Alg.~\ref{algo::pardircollapse_v0}).
+We define $\border{X}$ as the set of all border faces of $X$.
+We now  introduce a directional parallel thinning algorithm (Alg.~\ref{algo::pardircollapse_v0}).
+
+On a single execution of the main loop of Alg.~\ref{algo::pardircollapse_v0}, only faces located on the border of the complex are removed (l.~\ref{pc:07}).
+Thanks to corollary \ref{cor:parallel_removal_v2}, we also remove faces with same direction and orientation in parallel (l.~\ref{pc::08}).
 
@@
 
-Intuitively, we want the algorithm to remove free faces \quotes{layer by layer}: we don't want, after a single execution ($\ell =1$), to have unequal thinning of the input complex.
-Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex: this is why we introduce, on line \ref{pc:02} the set $L$, and that we remove only faces located in $L$ on line \ref{pc:08}.
-The sets $E$ (line \ref{pc:06}) and $G$ (line \ref{pc:07}) allows to remove whole sets of free faces in parallel from $X$, thanks to the direction and orientation of faces previously defined.
-A detailed view of each step of the algorithm is shown on Fig.~\ref{fig:step_thinning}.
-
-Different definitions of the orientation and direction can be given, corresponding to different order of removal of free faces in the complex.
-These changes lead to different results, but arbitrary choices on the order of removal of free pairs must be made in order to obtain, at the end, a thin skeleton (no more $n$-faces when working in the $n$-dimension).
-Once orientation and direction have been defined, the results of the algorithm are uniquely defined.
-
+Different definitions of the orientation and direction can be given, corresponding to different order of removal of free faces in the complex and leading to different results.
 Algorithm~\ref{algo::pardircollapse_v0} may be easily implemented to run in linear time complexity (proportionally to the number of faces of the complex).
 Indeed, checking if a face is free or not may be easily done in constant time.
-Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the faces of $f$ and the cofaces of $g$ in order to find new free faces, as other faces' status won't change (the implemented algorithm contains these optimizations).
+Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the elements of $\cellstrict{f} \union \contstrict{g}$ in order to find new free faces, as other faces' status won't change.
+
+The following property is a direct consequence of of Prop.~\ref{prop:thinness} and the fact that Alg.~\ref{algo::pardircollapse_v0} stops when no more free faces can be found in the complex :
+
+\begin{proposition}
+\label{prop::alg_v0_thin}
+Let $\complex{n}{X}$, the result of $ParDirCollapse(X,\emptyset,+\infty)$ is a thin complex.
+\end{proposition}
 
@@
 
+\input{pathtracint_application_skeleton}
+
 \end{document}

### 2012-09-08 pathtracint_application_skeleton.tex
J'ai fait un include du fichier de Laurent dans le principal > il faut compiler le fichier principal.
Ajout de l'explication de la rendering equation et début de la partie sur la résolution de l'equation

@@
-\documentclass[final,envcountsame]{llncs}
-
-%\usepackage{graphicx} %Pouvoir insérer des images
-\usepackage{epsfig}
-\usepackage{amsmath} %Permet d'utiliser une jolie panoplie de symboles mathématiques
-\usepackage{amssymb}
-%\usepackage{amsthm}
-\usepackage{stmaryrd}
-\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
-\usepackage{amsfonts}
-
-\def\mydef#1{{\em #1}}
-\def\myem#1{{\em #1}}
-\def\myvec#1{\vec{#1}}
-
-\def\Nset{\mathbb{N}}
-\def\Zset{\Z}
-\def\Rset{\mathbb{R}}
-\def\Z{\mathbb{Z}}
-\def\Zhalf{\frac{\mathbb{Z}}{2}}
-\def\allfaces#1{\mathbb{F}^{#1}}
-\def\subfaces#1#2{\allfaces{#1}_{#2}}
-\def\subcomplex#1#2{#1 \preceq #2}
-\def\complex#1#2{\subcomplex{#2}{\allfaces{#1}}}
-\def\cell#1{\hat{#1}}
-\def\cellstrict#1{\hat{#1}^*}
-\def\cont#1{\check{#1}}
-\def\contstrict#1{\check{#1}^*}
-\def\closure#1{#1^-}
-\def\princ#1{#1^+}
-\def\detach#1#2{#2 \oslash #1}
-\def\attach#1#2{\mbox{\it{}Attach\/}(\cell{#1},#2)}
-\def\mydim#1{\dim(#1)}
-\def\type#1{\mbox{\it{}Dir\/}(#1)}
-\def\orient#1{\mbox{\it{}Orient\/}(#1)}
-\def\border#1{\mbox{\it{}CBorder\/}(#1)}
-\def\suchthat{\;|\;}
-\def\inter{\cap}
-\def\union{\cup}
-\def\dinterval#1#2{\{#1,\ldots,#2\}}
-\newcommand{\Compl}[1]{\overline{#1}}
-\def\finproof{\square}
-\def\Zhalf{\frac{\mathbb{Z}}{2}}
-\newcommand{\Card}[1]{|#1|}
-
-\def\figpath#1{../../Images/EPS/#1}
-
-\begin{document}
-
 \section{Application to path tracing}
@@
 
-\begin{equation}
-L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi) d\omega_\Phi
-\end{equation}
+\begin{align*}
+L(x \rightarrow \Theta) &= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi) d\omega_\Phi \\
+&= L_e(x \rightarrow \Theta) + L_r(x \rightarrow \Theta)
+\end{align*}
 
-It's actually this equation that a path tracer try to resolve.
-Let $O$ be the origine of the camera and $I$ the image we want to compute (a rectangle embedded in $\Rset^3$).
+This equation express an intuitive idea: the radiance produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.
+
+The reflected radiance is computed by taking all the possible directions of the hemisphere around $x$ and integrate the incoming radiance along those directions scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$.
+
+The cosine is used to attenuate the radiance that arrive by a "direction rasante".
+
+The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
+It has to respect three properties to be physically coherent:
+
+\begin{itemize}
+\item Helmotz reciprocity: $f_r(x, \Theta \leftrightarrow \Phi) = f_r(x, \Phi \leftrightarrow \Theta)$
+\item Positivity: $f_r(x, \Theta \leftrightarrow \Phi) \geq 0$
+\item Conservation of energy: $\int_{\Omega_x} f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi) d\omega_\Phi \leq 1$
+\end{itemize}
+
+It's actually the rendering equation that a path tracer try to resolve.
+Let $O$ be the origin of the camera and $I$ the image we want to compute (a rectangle embedded in $\Rset^3$).
 Then the path tracing algorithm compute an approximation of the radiances $L(O \leftarrow \vec{OP})$, $\forall P \in I$.
@@
 
-Parler rapidement de l'échantillonage et donner l'estimateur de monte carlo
+The rendering equation is a recursive integral equation. The integrand of the equation contains the radiance function that we must compute:
+
+\begin{align*}
+L(x \rightarrow \Theta) &= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi) d\omega_\Phi \\
+&= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi) d\omega_\Phi
+\end{align*}
+
+In practive it is impossible to compute anatically the integral, it has to be estimated.
+The most commonly used method is the Monte-Carlo integration that provide the following estimator for the integral:
+
+\begin{equation*}
+\langle L_r(x \rightarrow \Theta) \rangle = \frac{1}{N} \sum_{i = 1}^{N} \frac{f_r(x, \Theta \leftrightarrow \Phi_i) L(r(x, \Phi_i) \rightarrow -\Phi_i) cos(N_x, \Phi_i)}{p(\Phi_i)}
+\end{equation*}
+
+$p$ is a probability density function that is used to sample $N$ samples $(\Phi_i)_{i = 1...N}$ of the hemisphere $\Omega_x$.
+
+This estimator is unbiased, it means that the esperance $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$.
+
+% parler de la variance
 
@@
 Décrire comment combiner les stratégies brdf + skeleton
-
-\subsection{Coarse irradiance estimation using curvilinear skeleton}
-
-Décrire les heuristiques
-
-\end{document}

### 2012-09-08 pathtracint_application_skeleton.tex
Mise à jour de la partie sur la résolution de la RE

@@
 
-$p$ is a probability density function that is used to sample $N$ samples $(\Phi_i)_{i = 1...N}$ of the hemisphere $\Omega_x$.
+$p$ is a probability density function (pdf) that is used to sample $N$ samples $(\Phi_i)_{i = 1...N}$ of the hemisphere $\Omega_x$.
 
@@
 
-% parler de la variance
+The variance of the estimator express its quality. The lower the variance is, the better is the estimator. We have:
+
+\begin{equation*}
+\sigma^2[\langle L_r(x \rightarrow \Theta) \rangle] = \frac{1}{N} \int_{\Omega_x} (\frac{f(\Phi)}{p(\Phi)} - L_r(x \rightarrow \Theta))^2 d\omega_\Phi
+\end{equation*}
+
+Then if we increase the number of samples we decrease the variance and the estimator gives better results.
+Another way to decrease the variance is to choose a good pdf.
+Actually the best is to choose a pdf that match the shape of the function to integrate (ie.
+give high density to samples that have high values for the function and low density to samples that have low values).
+
+This str	ategy of choosing a good pdf is called \myem{importance sampling} and is over-used in global illumination to improve the convergence of the algorithms.
 

### 2012-09-08 bidir_pathtrace_skeleton.tex
Quelques corrections supplémentaires

@@
 \def\OneBall#1{\mathbb{B}^{1}_{#1}}
-\def\MaxOneBall#1{\hat{\OneBall{#1}}}
+\def\MaxOneBall#1{\mathbb{M}\OneBall{#1}}
 
@@
 \label{sec::aspect_no_param}
-Algorithm \ref{algo::pardircollapse_v0} does not necessarily preserves "geometrical information" of the original object in the resulting skeleton.
+Algorithm \ref{algo::pardircollapse_v0} does not necessarily preserves "geometrical information" from the original object in the resulting skeleton.
 For example, if the input was a filled human shape (one connected component, no tunnel nor cavity) and that $W=\emptyset$ and $l=+\inf$, then the result of Alg.~\ref{algo::pardircollapse_v0} would be a simple vertex.
@@
 In the following, we will see a new method in the cubical complex, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory visual properties.
-This will be achieved by adding, in the set $W$ of Alg.~\ref{algo::pardircollapse_v0} some faces of the original object.
+This will be achieved by adding, in the set $W$ of Alg.~\ref{algo::pardircollapse_v0}, some faces from the original object during the thinning process.
 
 \subsection{The lifespan of a face}
-In the following, we define new notions in the cubical complex. The first one we present is the \myem{death date} of a face.
+In the following, we define more notions in the cubical complex. The first one we present is the \myem{death date} of a face.
 
@@
 
-The death date of a face indicates how many layers of free faces should be removed from a complex $X$, using alg.~\ref{algo::pardircollapse_v0}, before removing completely the face from $X$.
+The death date of a face indicates how many layers of free faces should be removed from a complex $X$, using Alg.~\ref{algo::pardircollapse_v0}, before removing completely the face from $X$.
 We now define the \myem{birth date} of a face:
@@
 \begin{definition}
-Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the minimum between the smallest integer $b$ such that $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, and $\death{f}{X}$.
+Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or either $f \notin ParDirCollapse(X, \emptyset, b)$.
 \end{definition}
@@
 
-The lifespan of a face $f$ of $X$ indicates how many "rounds" this face "survives" as a facet in $X$, when removing free faces with algorithm \ref{algo::pardircollapse_v0}.
-The lifespan, the death and the birth, as defined here, are dependant of Alg.~\ref{algo::pardircollapse_v0}, used for performing the thinning.
-It is of course possible to use another algorithm for performing the thinning, leading to other values of birth, death and lifespan.
-It is recommended, in order to have "comparable" values, to compute these three parameters with the same thinning technique.
-
-Algorithm \ref{algo::lifespan} computes the lifespan of all faces of a complex.
-The algorithm is not linear in time, however, a linear implementation of such algorithm exists and is presented in the appendix.
-On Fig.~\ref{fig:lifespan_exemple}, we show a sequence of collapse allowing to compute the lifespan of an edge.
-%Todo lien avec annexe
-
-\begin{algorithm}[h!]
-\label{algo::lifespan}
-\caption[Lifespan]{$Lifespan(X)$}
-\SetKwFor{ForAll}{for all}{do}{end}
-\KwData{A cubical complex $\complex{n}{X}$}
-\KwResult{The map giving the lifespan of all faces of $X$}
-\ForAll{$f \in X$}{
-  $Death(f)=+\infty$\;
-  \If{$f \in \princ{X}$}{
-    $Birth(f)=0$\;
-  }
-  \Else{
-    $Birth(f)=+\infty$\;
-  }
-}
-$Y=X; l=0$\;
-\While{there exists free faces for $Y$}{
-  $Y=ParDirCollapse(Y,\emptyset,1)$\;
-  $l=l+1$\;
-  \ForAll{$f \in \princ{Y}$ such that $Birth(f)=+\infty$}{
-  	$Birth(f)=l$\;
-  }
-  \ForAll{$f \notin Y$ such that $Death(f)=+\infty$ and $f \in X$}{
-  	$Death(f)=l$\;
-  }
-}
-\ForAll{$f \in X$}{
-  $Birth(f)=\min(Birth(f), Death(f))$\;
-  \If{$Death(f)=+\infty$}{
-    $Lifespan(f)=+\infty$\;
-  }
-  \Else{
-    $Lifespan(f)=Death(f) - Birth(f)$\;
-  }
-}
-\Return $Lifespan$\;
-\end{algorithm}
+The lifespan of a face $f$ of $X$ indicates how many "rounds" this face "survives" as a facet in $X$, when removing free faces with Alg.~\ref{algo::pardircollapse_v0}.
 
@@
 Typically, higher the lifespan is, and more representative of an object's visual feature the face is.
-The lifespan, also called \myem{saliency}, was used in \cite{Liu2009} (under the name "medial persistence") in order to propose a thinning algorithm in cubical complexes based on two parameters.
+The lifespan, sometimes called \myem{saliency}, was used in \cite{Liu2009} (with the name "medial persistence") in order to propose a thinning algorithm in cubical complexes based on two parameters.
 
@@
 \label{subsec::distmap}
-In addition to the lifespan of a face, the proposed homotopic thinning method will use information on distance between faces in order to decide if a face should be kept safe from deletion.
-We define hereafter the various notions needed for this, based on distance in the DT framework.
+In addition to the lifespan of a face, the proposed homotopic thinning method uses information on distance between faces in order to decide if a face should be kept safe from deletion.
+We define hereafter the various notions needed for this, based on distance in the voxel framework.
 
@@
 The length of $\mathcal{C}$ is $k$.
+Remark that $d_1$ is indeed the $6$-distance in the 3d voxel framework.
 
@@
 The \myem{$1$-ball of radius $r$ centered on $x$} is the set $\OneBall{r}(x)=\{y \in \Zset^n | d_1(x,y) < r\}$.
-Remark that $d_1$ is indeed the so-called $4$-distance in the 2d DT framework, and the $6$-distance in the 3d DT framework.
 Given $X \subset \Zset^n$, the \myem{maximal $1$-ball of $X$ centered on $x$} is the set $\MaxOneBall{X}(x) = \OneBall{d_1(x,\Compl{X})}(x)$.
@@
 If $x \in \Compl{X}$, we set $\dd_1(x,\Compl{X}) = 0$.
-The map $\dd_1$ is known as the opening function (based on the $1$-distance): it allows to compute efficiently results of openings by balls of various radius, and gives information on the local thickness of an object on each of its points.
-We show some examples of the opening function on Fig.~\ref{fig::globalopening} and \ref{fig::detailopening}.
+The map $\dd_1$ is known as the opening function (based on the $1$-distance): it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.
 
-Given $\complex{n}{X}$, the value of $\dd_1(x,\Compl{X})$ of every $x \in X$ can be computed by performing successive dilations of values of the map $d_1$.
-The algorithm presented in Alg.\ref{algo::naive_opening} is naive, and a more efficient implementation (linear in time depending on the image's size) is discussed in the appendix (see Sec.~\ref{sec::better_open}, p.~\pageref{sec::better_open}).
+Given $X \subset \Zset^n$, the value of $\dd_1(x,\Compl{X})$ of every $x \in X$ can be naively computed by performing successive morphological dilations of values of the map $d_1$. 
+%TODO Parler du fait que l'on a une version dans la thèse, et une version plus amélioré dans thèse et public avenir
 
-In order to extend $d_1$ and $\dd_1$ to the cubical complex framework, let us introduce the map $\Phi^{-1}$, inverse of the bijective map $\Phi$ defined in Sec.~\ref{sec::binary_to_cubic}.
+In order to extend $d_1$ and $\dd_1$ to the cubical complex framework, let us introduce the map $\Phi^{-1}$, inverse of the bijective map $\Phi : \subfaces{n}{n} \rightarrow \Zset^{n}$ defined in Sec.~\ref{sec::binary_to_cubic}.
 It is used to project any $n$-face of $\allfaces{n}$ into $\Zset^n$.
@@
 
-Given $\complex{n}{X}$, we set $X^v = \Phi^{-1}(X \inter \subfaces{n}{n})$: the set $X^v$ is a subset of $\Zset^n$.
-We define the map $\tilde{D_1}(X): \allfaces{n} \rightarrow \Nset$ as an extension of $d_1$ to the cubical complex framework: for all $f \in \allfaces{n}$,
+Given $Y \subset \allfaces{n}$, we define the map $\tilde{D_1}(X): \allfaces{n} \rightarrow \Nset$ as an extension of $d_1$ to the cubical complex framework: for all $f \in \allfaces{n}$,
 
@@
 \begin{math}
-\tilde{D_1}(X)(f) = \left\{
+\tilde{D_1}(Y)(f) = \left\{
 \begin{array}{ll}
-d_1(\Phi^{-1}(f),\Compl{X^v}) & \text{if $f$ is an $n$-face}\\
-\displaystyle \max_{g \in \contstrict{f}} \tilde{D_1}(X)(g) & \text{else}
+d_1(\Phi^{-1}(f),\Phi^{-1}(Y \inter \subfaces{n}{n})) & \text{if $f$ is an $n$-face}\\
+\displaystyle \max_{g \in \contstrict{f} \inter \subfaces{n}{n}} \tilde{D_1}(Y)(g) & \text{else}
 \end{array}
@@
 
-The same way, we define $\tilde{\DD_1}(X): \allfaces{n} \rightarrow \Nset$ as an extension of $\dd_1$ to the cubical complex framework.
+Informally, if $f$ is a 3-face, then $\tilde{D_1}(Y)(f)$ is the length of the shortest 1-path between the voxel \quotes{corresponding} to $f$ and the set of voxels corresponding to $Y$.
+The same way, we define $\tilde{\DD_1}(Y): \allfaces{n} \rightarrow \Nset$ as an extension of $\dd_1$ to the cubical complex framework.
+
+Finally, we define the decenterness map:
+
+\begin{definition}
+The \myem{decenterness map}, denoted by $\centerness(Y)$, is equal to $\tilde{\DD_1}(Y) - \tilde{D_1}(Y)$.
+\end{definition}
 
 \subsection{Parameter-free thinning based on the lifespan, opening function and decenterness}
-Thanks to these notions, we can now define sets of faces that will help preserve the visual aspect of an object during thinning.
-First, let us define the decenterness of a complex as the map $\centerness(X) = \tilde{\DD_1}(X) - \tilde{D_1}(X)$.
-For each face of a complex $X$, the decenterness value of this face gives an information on how well a face is centered inside a visual feature of the object: the lower this value is, the better centered the face is.
-On Fig.~\ref{fig::globalopening}, we give an example of the decenterness map of a shape, where the highest values are represented by the darkest colours.
+As previously said, we will add edges to the set $W$ of Alg.~\ref{algo::pardircollapse_v0} in order to retain, in the resulting curvilinear skeleton, important edges from the original object.
+If an edge has a high decenterness value, then it is probably located too close to the object's border and does not represent an interesting visual feature to preserve.
+%TODO ref vers image de decenterness
+On the other hand, if an edge has a high lifespan, then it means it was not removed quickly, after becoming a cell, by the thinning algorithm and might represent some precious visual information on the original object.
+An idea would be to keep, during thinning, all edges whose lifespan is superior to the decenterness value.
+Unfortunately, this strategy produces skeletons with many spurious branches in surfacic areas of the original object.
 
-Faces relevant of the visual aspect of a complex must have a high lifespan ("survive" long to the homotopic thinning process) and a low decenterness (is centered in the object).
-In a $3$-dimensional complex (resp.
-a $2$-dimensional complex), squares (resp.
-lines) whose lifespan is higher than the decenterness will be chosen as relevant of the surfacic (resp.
-curvilinear) parts of the complex.
-
-\begin{definition}
-Given $\complex{n}{X}$, the \myem{$k$-LC axis} (stands for "Lifespan Centerness") of $X$ is the set
-
-\begin{center}
-$\axislc{k}(X) = \closure{\{f \in X$ $|$ $\dim(f)=k$ and $\lifespan{f}{X} > \centerness(X)(f)\}}$.
-\end{center}
-
-\end{definition}
-
-The LC-axis fails in selecting a good set of curves relevant of the curvilinear parts of a three-dimensional complex.
-Indeed, given a $3$-complex $X$, the set $\axislc{1}(X)$ contains generally too many curves.
-This happens because the algorithm $ParDirCollapse$ "takes more time" to eliminate lines than to eliminate squares in a $3$-complex.
-Consequently, lines of a $3$-complex tend to have a high lifespan even though they are not representative of any curvilinear part of the complex.
-The thicker is the input object, and the more important is the phenomenon.
-
-Lines relevant of the curvilinear parts of a $3$-complex have a high lifespan and a low decenterness, especially in the thick parts of the complex.
-Lines whose lifespan is higher than the decenterness added to the local thickness of the complex will be relevant of the curvilinear parts of the complex.
-
-\begin{definition}
-Given $\complex{n}{X}$, the \myem{$k$-LOC axis} (stands for "Lifespan Opening Centerness") of $X$ is the set
-
-\begin{center}
-$\axisloc{k}(X) = \closure{\{f \in X$ $|$ $\dim(f)=k$ and $\lifespan{f}{X} > \tilde{\DD_1}(X)(f) + \centerness(X)(f)\}}$.
-\end{center}
-\end{definition}
-
-The various sets previously defined represent faces which should be kept safe from deletion during homotopic thinning of a complex in order to obtain a pruned skeleton containing visual features from the original object.
-The set $\axislc{1}$ of a $2$-complex represents curvilinear parts to keep in this complex, the set $\axisloc{1}$ of a $3$-complex represents curvilinear parts to keep in this complex, and the set $\axislc{2}$ of a $3$-complex represents surfacic parts to keep in this complex.
-
-\begin{lemma}
-\label{lem::axis_subset_facet}
-Given $\complex{n}{X}$, for every $k<n$, $\axislc{k}(X) \subseteq CollapseFacet(X, \emptyset)$ and $\axisloc{k}(X) \subseteq CollapseFacet(X, \emptyset)$.
-\end{lemma}
-
-Lemma \ref{lem::axis_subset_facet} is straightforward once observed that $CollapseFacet(X, \emptyset)$ is indeed the set of $k$-faces of $X$ ($k<n$) which have a strictly positive lifespan in $X$. 
-Based on these sets, we propose three algorithms: one for computing a 1d skeleton from a bi-dimensional complex, one for computing a 2d skeleton (which can contain 1d parts) from a three-dimensional complex, and one for computing a curvilinear skeletons from a three-dimensional complex.
-This last algorithm should produce one-dimensional results (a set of lines) however, the output can be a two-dimensional complex.
+We can identify surfacic areas of a complex as zones where squares have a high lifespan.
+Therefore, in order to avoid spurious branches in surfacic areas, we need to make it harder for edges to appear in these zones.
+It can be achieved by deciding that an edge will be kept safe from deletion by the thinning algorithm if its lifespan is superior to the decenterness value plus the lifespan of squares \quotes{around} this edge.
+This leads us to
 
@@
 
-Algorithm \ref{algo::3d_input_curv_skel} produces as output a 2d or 1d complex, result of the homotopic thinning of a three-dimensional complex.
-The skeleton produced by this algorithm should be a set of lines describing the shape of the input, however, it may contain squares representative of the topology of the initial object (if the input complex contains a cavity, the skeleton will contain surfacic parts).
-
-Despite the use of constraint sets (sets of faces which should not be removed during thinning) in the three algorithms, they produce thin results.
--> On peut simplement modifier algo pour y incorporer directement les elements a conserver pdt la skel...
-
 \input{pathtracint_application_skeleton}

### 2012-09-09 pathtracint_application_skeleton.tex
Mise à jour de la partie applications

@@
 
-The cosine is used to attenuate the radiance that arrive by a "direction rasante".
+The cosine is used to attenuate the radiance that arrive by a "direction rasante". $N_x$ is the surface's normal at point $x$.
 
 The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
-It has to respect three properties to be physically coherent:
+Intuitively it represents the material properties of the surface at point $x$.
+A mirror doesn't reflect light the same manners a wall do, the BRDF encodes that.
+The BRDF has to respect three properties to be physically coherent:
 
@@
 
+In the path tracing algorithm we cannot use more than one sample.
+Actually the evaluation of the integrand implies the evaluation of $L$ and thus a new application of the Monte-Carlo estimator.
+So if we use more than one sample the complexity of the algorithm becomes exponential (O($d^N$) with $d$ the length of a path and $N$ the number of samples).
+
 \subsubsection{The algorithm}
 
-Donner l'algorithme en pseudo-code
+We are now able to construct the algorithm for computing radiance:
+
+\begin{algorithm}[h!]
+\label{algo::incoming_radiance}
+\caption[incoming_radiance]{Incoming radiance $L(x \leftarrow \Phi)$}
+\KwData{An incoming direction $\Phi$, a point $x$} 
+\KwResult{$L(x \leftarrow \Phi)$}
+\Return $L(r(x, \Phi) \rightarrow -\Phi)$\;
+\end{algorithm}
+
+\begin{algorithm}[h!]
+\label{algo::radiance}
+\caption[radiance]{Radiance $L(x \rightarrow \Theta)$}
+\KwData{An outcoming direction $\Theta$, a point $x$} 
+\KwResult{$L(x \rightarrow \Theta)$}
+$L \gets L_e(x \rightarrow \Theta)$\;
+\If{continue the recursion} {
+	$\Phi \gets sample(\Omega_x, p)$ \\
+	$L \gets L + \frac{f_r(x, \Phi, \Theta) L(x \leftarrow \Phi) cos(N_x, \Phi)}{p(\Phi)}$\;
+}
+\Return L\;
+\end{algorithm}
+
+This algorithm is called for each pixel of the final image to compute. To improve the quality of the result we can throw multiple rays accross each pixel and computing the mean of the results.
+
+In the radiance algorithm the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$. We will discuss what pdf we use in the next section.
+
+The stopping condition of the recursion is not specified here.
+In our implementation we simply fix a maximum depth for the path but it's a bad solution.
+Actually it biased the method because some light paths have no chance to be evaluated.
+A better solution is to use a method called russian roulette.
 
@@
 
-Décrire la stratégie d'échantillonage basée sur le squelette
+We have presented the path tracing algorithm without describe the pdf $p$ used to sample the hemisphere around a point.
+The choice of the pdf is very important because the convergence speed of the algorithm depends on it.
+Here is the integrand of the integral estimated with Monte-Carlo:
 
-\subsection{Multiple importance sampling}
+\begin{equation*}
+f_r(x, \Phi, \Theta) L(x \leftarrow \Phi) cos(N_x, \Phi)
+\end{equation*}
 
-Décrire comment combiner les stratégies brdf + skeleton
+To improve the quality we have to choose a pdf that match the shape of this function.
+There is a cosine factor in the function.
+It takes higher values when $\Phi$ is close to $N_x$.
+That means we have to give better chances to be samples to directions that are close to the normal.
+But it's not enough: when the environment is not totally diffuse, the brdf take higher values for some pair of directions.
+A perfect mirror, for exemple, reflect light only in a single direction: the symetric of $\Theta$ by the normal $N_x$.
+In that case it's totally useless to sample $\Omega_x$, we juste have to take the correct direction without taking into account the cosine.
+
+The most common strategy used to sample $\Omega_x$ is to use the BRDF because we know it: it is given as an input of the algorithm for exemple with textures.
+
+$L$ is rarelly used because we cannot evaluate it in constant time.
+Remember that it represents the distribution of light in the scene.
+It takes high values when $\Phi$ carry a lot of energy.
+Out method is based an a topological skeleton of the void (because light travels in the void) to sample $\Omega_x$ according to $L$.
+
+\subsubsection{Computation of the skeleton}
+
+The skeleton is computed in two steps:
+
+\begin{itemize}
+\item Voxelization of the void
+\item Skeletonization in the cubical complex framework
+\end{itemize}
+
+The voxelization is done with a software called \of Binvox \fg.
+It voxelizes the surfaces of the scene and we invert it to get the voxelization of the void.
+The voxelization is then converted to a cubical complex and we apply our algorithm to skeletonize it.
+We get a curvilinear skeleton that has the same topology than the void.
+
+The skeleton is represented by a graph of 3D positions with edges encoding the topology.
+
+\subsubsection{Construction of the importance points}
+
+Given the skeleton we compute a set of points called importance points. This points will be used to sample $\Omega_x$ in the path tracing algorithm.
+
+Let $L$ be a light of the scene ($L \in S_L$) and $n_L$ the shortest visible node by $L$.
+We first compute a tree of shortest paths along the skeleton with $n_L$ as root.
+For that we use the Dijkstra algorithm with a distance based on visibility from $n_L$: We put $d(n, m) = 1$ if $m$ is visible from $n_L$ and $d(n, m) = 10$ if not.
+It results that the paths that are enlighted by $L$ will have a shorter distance.
+
+Let $n$ be a node of the skeleton and $V_n$ the set of visible nodes from $n$ along the shortest path toward $n_L$.
+The importance point $imp_n$ associated to $n$ is computed by taking the barycenter of $V_n$.
+
+\subsubsection{Sampling according to $L$}
+
+\subsubsection{Problems}
+
+\subsection{Skeleton based multiple importance sampling}
+
+\subsubsection{Multiple importance sampling}
+
+\subsubsection{Application to path tracing}

### 2012-09-09 pathtracint_application_skeleton.tex
Fin du draft sur la partie application

@@
 
-The voxelization is done with a software called \of Binvox \fg.
+The voxelization is done with a software called Binvox.
 It voxelizes the surfaces of the scene and we invert it to get the voxelization of the void.
@@
 
+Given a point $x$ and a direction $\Theta$ we want to compute $L(x \rightarrow \Theta)$ and then sample the hemisphere $\Omega_x$.
+We search for the nearest skeleton node $n$ and its importance point $imp_n$.
+We sample the hemisphere with a power-cosine pdf center on $\vec{ximp_n}$:
+
+\begin{equation*}
+p_{skel}(\Phi) = \frac{s + 1}{2\pi} cos^s \alpha
+\end{equation*}
+
+with $\alpha$ the angle between $\vec{ximp_n}$ and $\Phi$ and $s$ a parameter called skeleton strength.  The more $s$ is high, the more we sample close $\vec{ximp_n}$.
+
+\subsubsection{Results}
+
 \subsubsection{Problems}
+
+A few problems arise from this sampling strategy:
+
+\begin{itemize}
+\item $\vec{ximp_n}$ may point backward the surface (ie. $cos(\vec{ximp_n}, N_x) < 0$)
+\item The importance point $imp_n$ may be not visible from $x$
+\end{itemize}
+
+In those case we don't use the skeleton and switch to a basic BRDF sampling.
+
+Another problem from the method is the apparition of speckles.
+Speckles are anormally white pixels on the computed image that tend to vanish when we augment the number of rays thrown by pixel.
+It comes from the Monte-Carlo estimator when we sample a direction $\Phi$ with a low pdf $p(\Phi)$ but a high value $f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi)$: the ratio becomes very high and speckles appears.
+In the next section we discuss a method to attenuate the speckles.
 
@@
 
+Our strategy is good, especially in dark areas.
+It reduces noise by taking into account the distribution of energy in the scene.
+The problem is that we take into account only one importance direction for sampling, we totally ignore the BRDF and the cosine factor (unless we fall in a bad case like the two describe above).
+We implemented a solution called multiple importance sampling (MIS) that is able to combine multiple sampling strategies into one.
+
 \subsubsection{Multiple importance sampling}
 
-\subsubsection{Application to path tracing}
+Let $f(x) = f_1(x)f_2(x)...f_n(x)$ a function that we want to integrate over a domain $D$:
+
+\begin{equation*}
+F = \int_{D} f_1(x)f_2(x)...f_n(x) dx
+\end{equation*}
+
+Monte-Carlo suggests us to use the estimator (with one sample $x$ in that case, like in path tracing):
+
+\begin{equation*}
+\langle F \rangle = \frac{f_1(x)f_2(x)...f_n(x)}{p(x)}
+\end{equation*}
+
+But how to choose efficiently $p$ ?
+Suppose we have $n$ sampling strategy $p_1, p_2, ..., p_n$, with $p_i$ efficient for the function $f_i$.
+Let $c_1, ..., c_n$ be $n$ scalar from $[0,1]$ such that $\sum_{i = 1}{n} c_i = 1$.
+$c_i$ is the probability to select the strategy $p_i$ for sampling.
+Then the combined pdf of a sample $x$ is:
+
+\begin{equation*}
+p(x) = \sum_{i = 1}^{n} c_i p_i(x)
+\end{equation*}
+
+We use this pdf for Monte-Carlo estimation.
+First we choose the strategy to use, based on the probabilities $(c_i)_{i = 1...n}$, then we sample with $p_i$ and then we estimate the integral by $\frac{f_1(x)f_2(x)...f_n(x)}{p(x)}$.
+This method is called multiple importance sampling.
+
+\subsubsection{MIS for skeleton based path tracing}
+
+In our case we have the following function to integrate:
+
+\begin{equation*}
+f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi)
+\end{equation*}
+
+We suppose that we have a pdf $p_{brdf}$ that efficiently sample $\Omega_x$ according to the product $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$ and our strategy $p_{skel}$ that efficiently sample according to $L(x \leftarrow \Phi)$.
+
+Let $c_{skel}$ be the probability to sample with the strategy $p_{skel}$ and $1 - c_{skel}$ the probability to sample with $p_{brdf}$ ($c_{skel}$ could be a parameter of the program).
+That gives us the following algorithm to sample:
+
+\begin{algorithm}[h!]
+\label{algo::MIS}
+\caption[MIS]{Multiple importance sampling}
+\KwData{An outcoming direction $\Theta$, a point $x$} 
+\KwResult{A incoming direction $\Phi$ and its pdf $p(\Phi)$}
+$r \gets$ a random number between 0 and 1\;
+\If{$r \leq c_{skel}$} {
+	\Return $\Phi \gets sample(\Omega_x, p_{skel})$
+} \Else {
+	\Return $\Phi \gets sample(\Omega_x, p_{brdf})$
+}
+
+\Return $\Phi, p(\Phi) \gets c_{skel} p_{skel}(\Phi) + (1 - c_{skel}) p_{brdf}(\Phi)$
+\end{algorithm}
+
+\subsubsection{Results}

### 2012-09-09 bidir_pathtrace_skeleton.tex
Nouvelles corrections

@@
 
-\def\dd{\omega}
-\def\DD{\Omega}
+\def\mydist1{D_1}
+\def\mydd1{\Omega_1}
+\def\mydecent{\mathcal{DC}_1}
+\def\dist1{d_1}
+\def\Dist1#1{\mydist1(#1)}
+\def\DD1#1{\mydd1(#1)}
+\def\decent#1{\mydecent(#1)}
+\def\cplDist1#1{\mydist1^{cc}(#1)}
+\def\cplDD1#1{\mydd1^{cc}(#1)}
+\def\cpldecent#1{\mydecent^{cc}(#1)}
 
@@
 \def\Zset{\Z}
+\def\Nset{\mathbb{N}}
 \def\Rset{\mathbb{R}}
@@
 If $\mydim{X} = d$, then we say that $X$ is a $d$-complex.
-In $\allfaces{n}$, a complex $X$ is \myem{thin} if $\mydim{X}<n$; in $\allfaces{3}$, a complex is thin if it contains no cube.
 
@@
 
-If $Y$ is a collapse of a complex $X$, then $Y$ is a complex. 
-We now introduce some elements that will serve for proving the thinness of our skeletons.
-Let $f_0,f_\ell$ be two $n$-faces of $\allfaces{n}$ (with $\ell$ being even). An \myem{$(n-1)$-path from $f_0$ to $f_\ell$} is a sequence $\pi =
-(f_0,...,f_\ell)$ of faces of $\allfaces{n}$ such that for all $i \in \dinterval{0}{\ell}$, either $i$ is even and $f_i$ is an $n$-face, or $i$ is odd and $f_i$ is an $(n-1)$-face with $\contstrict{f_i}=\{f_{i-1}, f_{i+1}\}$ (such path always exists).
-
-\begin{proposition}
-\label{prop:thinness}
-Let $\complex{n}{X}$ be an $n$-complex, with $n > 0$. Then $X$ has at least one free $(n-1)$-face.
-\end{proposition}
-
-\begin{proof}
-Since $X$ is an $n$-complex (hence $X$ is finite) there exists an $n$-face $a$ in $X$ and an $n$-face $b$ in $\Compl{X}$.
-Obviously, there exists an $(n-1)$-path from $a$ to $b$.
-Let $h$ be the first $n$-face of $\pi$ that is not in~$X$, let $k$ be the last $n$-face of $\pi$ before $h$ (thus $k$ is in $X$), and let $e = k \inter h$ be the $(n-1)$-face of $\pi$ between $k$ and $h$.
-Since $k$ and $h$ are the only two $n$-faces of $\allfaces{n}$ that contain $e$, we see that the pair $(k,e)$ is free for $X$.
-$\finproof$
-\end{proof}
-
-In conclusion, in $\allfaces{3}$, as long as a complex still contains $3$-faces, it has a free $2$-face and more collapse operations can be performed.
-Therefore, it is possible to perform collapse on a complex until no more $3$-faces (volumes) can be found (until it is thin).
-
 \section{A parallel directional thinning based on cubical complex}
@@
 
-We define a surjective function $\type{}: \allfaces{n} \times \allfaces{n} \rightarrow \dinterval{0}{n-1}$ such that, for all free pairs $(f,g)$ and $(i,j)$ for $X$, $\type{f,g} = \type{i,j}$ if and only if $V(f,g)$ and $V(i,j)$ are collinear (we don't bother defining $\type{}$ for non free pairs as it won't be useful in this case).
+We define a surjective function $\type{}: \allfaces{n} \times \allfaces{n} \rightarrow \dinterval{0}{n-1}$ such that, for all free pairs $(f,g)$ and $(i,j)$ for $X$, $\type{f,g} = \type{i,j}$ if and only if $V(f,g)$ and $V(i,j)$ are collinear.
 The number $\type{f,g}$ is called the \myem{direction} of the free pair $(f,g)$.
@@
 If $f\neq k$, then we have $g \neq \ell$, $\contstrict{g} \inter X = \{f\}$ ($g$ is free for $X$) and $\contstrict{\ell} \inter X = \{k\}$ ($\ell$ is free for $X$).
-Thus, we have $\contstrict{\ell} \inter Y = \{k\}$ as $\ell\neq g$ and $k\neq f$.
-Therefore, $(k,\ell)$ is a free pair for $Y$.
+Thus, we have $\contstrict{\ell} \inter Y = \{k\}$ as $\ell\neq g$ and $k\neq f$ and $(k,\ell)$ is a free pair for $Y$.
 $\finproof$
@@
 
-\subsection{Distance map and opening function}
+\subsection{Distance map, opening function and decenterness map}
 \label{subsec::distmap}
@@
 
-We set $d_1(x,y)$ as the length of the shortest $1$-path from $x$ to $y$.
-Let $S \subset \Zset^n$, we set $d_1(x,S) = \displaystyle \min_{y \in S} d_1(x,y)$.
-The \myem{$1$-ball of radius $r$ centered on $x$} is the set $\OneBall{r}(x)=\{y \in \Zset^n | d_1(x,y) < r\}$.
-Given $X \subset \Zset^n$, the \myem{maximal $1$-ball of $X$ centered on $x$} is the set $\MaxOneBall{X}(x) = \OneBall{d_1(x,\Compl{X})}(x)$.
+We set $\dist1(x,y)$ as the length of the shortest $1$-path from $x$ to $y$.
+Let $S \subset \Zset^n$, we set, for all $x \in Zset^n$, the map $\Dist1{S}: \Zset^n \rightarrow \Nset$ such that $\Dist1{S}(x) = \displaystyle \min_{y \in \Compl{S}} \dist1(x,y)$.
 
-We set, for all $x \in X$, $\dd_1(x,\Compl{X}) = \displaystyle \max_{x \in \MaxOneBall{X}(y)} d_1(y,\Compl{X})$: this value indicates the radius of the largest maximal $1$-ball contained in $X$ and containing $x$.
-If $x \in \Compl{X}$, we set $\dd_1(x,\Compl{X}) = 0$.
-The map $\dd_1$ is known as the opening function (based on the $1$-distance): it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.
+The \myem{maximal $1$-ball of $S$ centered on $x$} is the set $\MaxOneBall{S}(x)=\{y \in \Zset^n | \dist1(x,y) < \Dist1{S}(x)\}$.
+We set, for all $x \in S$, the map $\DD1{S}: \Zset^n \rightarrow \Nset$ such that $\DD1{S}(x) = \displaystyle \max_{x \in \MaxOneBall{S}(y)} \Dist1{S}(y)$: this value indicates the radius of the largest maximal $1$-ball contained in $S$ and containing $x$.
+If $x \in \Compl{S}$, we set $\DD1{S}(x) = 0$.
+The map $\DD1$ is known as the opening function (based on the $1$-distance): it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.
 
-Given $X \subset \Zset^n$, the value of $\dd_1(x,\Compl{X})$ of every $x \in X$ can be naively computed by performing successive morphological dilations of values of the map $d_1$. 
+Given $S \subset \Zset^n$, the value of $\DD1{S}(x)$ of every $x \in S$ can be naively computed by performing successive morphological dilations of values of the map $\Dist1{S}$. 
 %TODO Parler du fait que l'on a une version dans la thèse, et une version plus amélioré dans thèse et public avenir
 
-In order to extend $d_1$ and $\dd_1$ to the cubical complex framework, let us introduce the map $\Phi^{-1}$, inverse of the bijective map $\Phi : \subfaces{n}{n} \rightarrow \Zset^{n}$ defined in Sec.~\ref{sec::binary_to_cubic}.
+Finally, we define the \myem{decenterness map} :
+
+\begin{definition}
+\label{def::decenterness_map}
+Given $S \subset \Zset^n$, the \myem{decenterness map of $S$} is the map $\decent{S} = \DD1{S} - \Dist1{S}$.
+\end{definition}
+
+In order to extend all these previous maps defined in $\Zset^n$ to the cubical complex framework, let us introduce the map $\Phi^{-1}$, inverse of the bijective map $\Phi : \subfaces{n}{n} \rightarrow \Zset^{n}$ defined in Sec.~\ref{sec::binary_to_cubic}.
 It is used to project any $n$-face of $\allfaces{n}$ into $\Zset^n$.
@@
 
-Given $Y \subset \allfaces{n}$, we define the map $\tilde{D_1}(X): \allfaces{n} \rightarrow \Nset$ as an extension of $d_1$ to the cubical complex framework: for all $f \in \allfaces{n}$,
+Given $Y \subset \allfaces{n}$, we define the map $\cplDist1{Y}: \allfaces{n} \rightarrow \Nset$ as follows: for all $f \in \allfaces{n}$,
 
@@
 \begin{math}
-\tilde{D_1}(Y)(f) = \left\{
+\cplDist1{Y}(f) = \left\{
 \begin{array}{ll}
-d_1(\Phi^{-1}(f),\Phi^{-1}(Y \inter \subfaces{n}{n})) & \text{if $f$ is an $n$-face}\\
-\displaystyle \max_{g \in \contstrict{f} \inter \subfaces{n}{n}} \tilde{D_1}(Y)(g) & \text{else}
+\Dist1{\Phi^{-1}(Y \inter \subfaces{n}{n})}(Phi^{-1}(f)) & \text{if $f$ is an $n$-face}\\
+\displaystyle \max_{g \in \contstrict{f} \inter \subfaces{n}{n}} \cplDist1{Y}(g) & \text{else}
 \end{array}
@@
 
-Informally, if $f$ is a 3-face, then $\tilde{D_1}(Y)(f)$ is the length of the shortest 1-path between the voxel \quotes{corresponding} to $f$ and the set of voxels corresponding to $Y$.
-The same way, we define $\tilde{\DD_1}(Y): \allfaces{n} \rightarrow \Nset$ as an extension of $\dd_1$ to the cubical complex framework.
-
-Finally, we define the decenterness map:
-
-\begin{definition}
-The \myem{decenterness map}, denoted by $\centerness(Y)$, is equal to $\tilde{\DD_1}(Y) - \tilde{D_1}(Y)$.
-\end{definition}
+Informally, if $f$ is a 3-face, then $\cplDist1{Y}(f)$ is the length of the shortest 1-path between the voxel \quotes{corresponding} to $f$ and the set of voxels corresponding to $Y$.
+The same way, we define $\cplDD1{Y}$ and $\cpldecent{Y}$.
 
 \subsection{Parameter-free thinning based on the lifespan, opening function and decenterness}
-As previously said, we will add edges to the set $W$ of Alg.~\ref{algo::pardircollapse_v0} in order to retain, in the resulting curvilinear skeleton, important edges from the original object.
-If an edge has a high decenterness value, then it is probably located too close to the object's border and does not represent an interesting visual feature to preserve.
+As previously said, we add edges to the set $W$ of Alg.~\ref{algo::pardircollapse_v0} in order to retain, in the resulting curvilinear skeleton, important edges from the original object.
+Given a cubical complex $X$, if an edge of $X$ has a high decenterness value for $X$, then it is probably located too close to the border of $X$ and does not represent an interesting visual feature to preserve.
 %TODO ref vers image de decenterness
-On the other hand, if an edge has a high lifespan, then it means it was not removed quickly, after becoming a cell, by the thinning algorithm and might represent some precious visual information on the original object.
+On the other hand, if an edge has a high lifespan for $X$, then it means it was not removed quickly, after becoming a cell, by the thinning algorithm and might represent some precious visual information on the original object.
 An idea would be to keep, during thinning, all edges whose lifespan is superior to the decenterness value.
@@
 It can be achieved by deciding that an edge will be kept safe from deletion by the thinning algorithm if its lifespan is superior to the decenterness value plus the lifespan of squares \quotes{around} this edge.
-This leads us to
+This leads us to proposing Alg.~\ref{algo::3d_curvskel}.
 
 \begin{algorithm}[ht]
-\label{algo::3d_input_curv_skel}
-\SetKwFor{ForAll}{for all}{do}{end}
+\label{algo::3d_curvskel}
 \caption[CurvilinearSkeleton]{$CurvilinearSkeleton(X)$}
 \KwData{A cubical complex $\complex{3}{X}$}
-\KwResult{A cubical complex $\complex{3}{Y}$ such that $\dim{Y} \leq 2$}
-$W=\axisloc{1}(X)$\;
-$ParDirCollapse(X,W,+\infty)$\;
-\Return $X$\;
+\KwResult{A cubical complex $\complex{3}{Y}$}
+\lnl{curvskel:01} $W=\{f \in X | \lifespan{f}{X} > \cpldecent{X}(f) + \birth{f}{X} - \cplDist1{X}(f)$ and $\mydim{f}=1\}$\;
+\Return $ParDirCollapse(X,W,+\infty)$\;
 \end{algorithm}
+
+In order to understand what was realised on line \ref{curvskel:01} of Alg.~\ref{algo::3d_curvskel}, we might point out that the birth date of an edge corresponds to the highest death date of the squares containing this edge.
+Moreover, the map $\cplDist1{X}$ gives, for all 3-faces of $X$, their death date (as the thinning algorithm naturally follows this map to eliminate cubes from a 3-complex).
+Therefore, for an edge $f$ of $X$, $\cplDist1{X}(f)$ informs us on the highest death date of cubes containing $f$, also equal to the highest birth date of squares containing $f$.
+In conclusion, $\birth{f}{X} - \cplDist1{X}(f)$ is an approximation of the lifespan of the squares containing $f$.
 

### 2012-09-10 bidir_pathtrace_skeleton.tex
Nouvelles corrcections + bibliographie

@@
 
-In the 3d voxel framework, objects are mae of voxels. In the 3d cubical complex framework, objects made of voxel, but of cubes, squares, lines and vertices. 
+In the 3d voxel framework, objects are made of voxels. In the 3d cubical complex framework, objects are made of cubes, squares, lines and vertices. 
 Let $\Z$ be the set of integers, we consider the family of sets $\mathbb{F}^1_0$ and $\mathbb{F}^1_1$, such that $\mathbb{F}^1_0 = \{\{a\} \suchthat a \in \Z\}$ and $\mathbb{F}^1_1 = \{ \{a, a+1\} \suchthat a \in \Z\}$.
@@
 Any element of $\cell{f}$ is \myem{a face of $f$}, and any element of $\cellstrict{f}$ is \myem{a proper face of $f$}.
-We call \myem{star of $f$} the set $\cont{f} = \{g \in \allfaces{n} \vert f \subseteq g\}$, and we write $\contstrict{f} = \cont{f} \setminus \{f\}$: any element of $\cont{f}$ is a \myem{coface of $f$}.
-It is plain that $g\in\cell{f}$ iff $f\in\cont{g}$.
+We set $\cont{f} = \{g \in \allfaces{n} \vert f \subseteq g\}$, and $\contstrict{f} = \cont{f} \setminus \{f\}$.
+It is plain that $g \in \cell{f}$ iff $f \in \cont{g}$.
 
@@
 \begin{definition}
-A finite set $X$ of faces in $\allfaces{n}$ is a \myem{cubical complex} if $X=\closure{X}$, and we write $\complex{n}{X}$.
+A finite set $X$ of faces in $\allfaces{n}$ is a \myem{complex} if $X=\closure{X}$, and we write $\complex{n}{X}$.
 
@@
 
-Informally, in 3d, a set of faces $X$ is each side (square) of a cube belonging to $X$ also belong to $X$, each edge of a square belonging to $X$ also belong to $X$, and each vertex of an edge belonging to $X$ also belong to $X$.
+Informally, in 3d, a set of faces $X$ is a complex if each side (square) of a cube of $X$ also belong to $X$, each edge of a square of $X$ also belong to $X$, and each vertex of an edge of $X$ also belong to $X$.
 
@@
 
-Traditionally, a voxel image is defined as a finite subset of $\Zset^n$.
-This kind of image is the most common one in the field of image processing and we give now a simple way to transpose a binary image to the cubical complex framework.
+Traditionally, a binary image is a finite subset of $\Zset^n$ (called voxel image when $n=3$). We give now a simple way to transpose such image to the cubical complex framework.
 
-Informally, to do so, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$ (to a pixel we associate a square, to a voxel we associate a cube).
+To do so, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$.
 More precisely, let $x=(x_1,...,x_n) \in S$, we define the n-face $\Phi(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$.
 We can extend the map $\Phi$ to sets: $\Phi(S)=\{\Phi(x) | x \in S\}$.
-Given a binary image $S$, we associate to it the cubical complex $\closure{\Phi(S)}$.
+Given a set $S \subset \Zset^n$, we associate to it the cubical complex $\closure{\Phi(S)}$.
 
 \subsection{Thinning: the collapse operation}
-The collapse operation is the basic operation for performing homotopic thinning of a complex.
-It consists of removing two distinct elements $(f,g)$ from a complex $X$ under the condition that $g$ is contained in $f$ and is not contained in any other element of $X$.
-This operation may be repeated several times.
+The collapse operation is the basic operation for performing homotopic thinning of a complex, and consists of removing two distinct faces $(f,g)$ from a complex $X$ under the condition that they are free:
 
 \begin{definition}
-Let $\complex{n}{X}$, and let $f,g$ be two faces of $X$.
-The face $g$ is \myem{free for $X$}, and the pair $(f,g)$ is \myem{a free pair for $X$} if $f$ is the only face of $X$ such that $g$ is a proper face of $f$.
+Let $\complex{n}{X}$, and let $f,g$ be two faces of $X$. The face $g$ is \myem{free for $X$}, and the pair $(f,g)$ is \myem{a free pair for $X$} if $(\contstrict{g} \inter X) = \{f\}$. 
 \end{definition}
 
-In other terms, $(f,g)$ is a free pair for $X$ whenever $\contstrict{g} \inter X = \{f\}$.
-It can be easily seen that if $(f,g)$ is a free pair for a complex $X$ and $\dim(f)=m$, then $f$ is a facet and $\dim(g) = m-1$.
+It can be easily seen that if $(f,g)$ is a free pair for a complex $X$, then $f$ is a facet of $X$ and $\dim(g) = \dim(f)-1$.
 
@@
 
-Considering two distinct free pairs $(f,g)$ and $(i,j)$ for $\complex{n}{X}$ such that $\type{f,g}=\type{i,j}$ and $\orient{f,g} = \orient{i,j}$, we have $f \neq i$.
+Considering two distinct free pairs $(f,g)$ and $(i,j)$ for $\complex{n}{X}$ such that $\type{f,g} = \type{i,j}$ and $\orient{f,g} = \orient{i,j}$, we have $f \neq i$.
 From this observation and Cor.~\ref{cor:parallel_removal}, we deduce the following property.
@@
 
-Intuitively, we want the algorithm to remove free faces of a complex \quotes{layer by layer}: we don't want to have unequal thinning of the input complex.
+Intuitively, we want the algorithm to remove free faces of a complex \quotes{layer by layer} and avoid having unequal thinning of the input complex.
 Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex.
-We say that a $d$-face of $X$ is a \myem{border face} if it contains a free $(d-1)$-face.
-We define $\border{X}$ as the set of all border faces of $X$.
-We now  introduce a directional parallel thinning algorithm (Alg.~\ref{algo::pardircollapse_v0}).
+We define $\border{X}$ as the set all faces belonging to a free pair for $X$. We now introduce Alg.~\ref{algo::pardircollapse_v0}, a directional parallel thinning algorithm.
 
 On a single execution of the main loop of Alg.~\ref{algo::pardircollapse_v0}, only faces located on the border of the complex are removed (l.~\ref{pc:07}).
-Thanks to corollary \ref{cor:parallel_removal_v2}, we also remove faces with same direction and orientation in parallel (l.~\ref{pc::08}).
+Thanks to corollary \ref{cor:parallel_removal_v2}, we remove faces with same direction and orientation in parallel (l.~\ref{pc:08}).
 
@@
 Different definitions of the orientation and direction can be given, corresponding to different order of removal of free faces in the complex and leading to different results.
-Algorithm~\ref{algo::pardircollapse_v0} may be easily implemented to run in linear time complexity (proportionally to the number of faces of the complex).
+Algorithm~\ref{algo::pardircollapse_v0} can be implemented to run in linear time complexity (proportionally to the number of faces of the complex).
 Indeed, checking if a face is free or not may be easily done in constant time.
-Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the elements of $\cellstrict{f} \union \contstrict{g}$ in order to find new free faces, as other faces' status won't change.
-
-The following property is a direct consequence of of Prop.~\ref{prop:thinness} and the fact that Alg.~\ref{algo::pardircollapse_v0} stops when no more free faces can be found in the complex :
-
-\begin{proposition}
-\label{prop::alg_v0_thin}
-Let $\complex{n}{X}$, the result of $ParDirCollapse(X,\emptyset,+\infty)$ is a thin complex.
-\end{proposition}
+Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the elements of $(\cellstrict{f} \union \contstrict{g})$ in order to find new free faces.
 
@@
 However, when one wants to preserve "geometrical information" from the original shape in the skeleton, then one expects to obtain a "stickman" from a human shape.
-However, it is important to not retain "too much information" during thinning in order to avoid noisy branches in the skeleton.
+It is however important to not retain "too much information" during thinning in order to avoid noisy branches in the skeleton.
 
@@
 
-The three parameters previously defined are dependant on the order of direction and orientation chosen for algorithm $ParDirCollapse$.
+These three values are dependant on the order of direction and orientation chosen for Alg.~\ref{algo::pardircollapse_v0}.
 
 The lifespan of a face $f$ of $X$ indicates how many "rounds" this face "survives" as a facet in $X$, when removing free faces with Alg.~\ref{algo::pardircollapse_v0}. 
-
 The lifespan is a good indicator of how important a face can be in an object.
@@
 If $x \in \Compl{S}$, we set $\DD1{S}(x) = 0$.
-The map $\DD1$ is known as the opening function (based on the $1$-distance): it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.
+The map $\DD1{S}$ is known as the opening function of $S$ (based on the $1$-distance): it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.
 
@@
 %TODO ref vers image de decenterness
-On the other hand, if an edge has a high lifespan for $X$, then it means it was not removed quickly, after becoming a cell, by the thinning algorithm and might represent some precious visual information on the original object.
+On the other hand, if an edge has a high lifespan for $X$, then it means it was not removed quickly, after becoming a facet, by the thinning algorithm and might represent some precious visual information on the original object.
 An idea would be to keep, during thinning, all edges whose lifespan is superior to the decenterness value.
@@
 
+\bibliographystyle{splncs03}
+\bibliography{./bibli}
+
 \end{document}

### 2012-09-10 bidir_pathtrace_skeleton.tex

@@
 \title{A 3d curvilinear skeletonization algorithm with application to bidirectional path tracing}
-\author{John Chaussard \and Venceslas Biri \and Michel Couprie}
-\institute{Universit\'{e} Paris Est, LABINFO-IGM, A2SI-ESIEE \\ 2, boulevard Blaise Pascal, Cit\'{e} DESCARTES \\ BP 99  93162 Noisy le Grand CEDEX, France \\ \email{chaussaj@esiee.fr, bertrang@esiee.fr, coupriem@esiee.fr}}
+\author{John Chaussard\inst{1} \and Laurent No\"{e}l\inst{2} \and Venceslas Biri\inst{2} \and Michel Couprie\inst{2}}
+\institute{Universit\'{e} Paris 13, Sorbonne Paris Cit\'{e}, LAGA, CNRS(UMR 7539), \\ F-93430, Villetaneuse, France \\ \email{chaussard@math.univ-paris13.fr}
+\and
+Universit\'{e} Paris Est, LABINFO-IGM, A3SI-ESIEE \\ 2, boulevard Blaise Pascal, Cit\'{e} DESCARTES \\ BP 99  93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}
 

### 2012-09-10 bidir_pathtrace_skeleton.tex
Change title + add some notes in the intro

@@
 
-\title{A 3d curvilinear skeletonization algorithm with application to bidirectional path tracing}
+\title{A 3d curvilinear skeletonization algorithm with application to path tracing}
 \author{John Chaussard\inst{1} \and Laurent No\"{e}l\inst{2} \and Venceslas Biri\inst{2} \and Michel Couprie\inst{2}}
@@
 \label{sec::intro}
-Blah blah blah...
+Noise in photorealistic computer generated images.
+Need to accelerate convergence.
+Add knowledge of the geometry and topology of the scene thanks to discrete geometry.
+Use the skeleton.
+Need of a robust curvilinear skeleton that represent topology and approximate geometry.
+Use the cubical complex framework.
 

### 2012-09-10 pathtracint_application_skeleton.tex
Corrections partie application

--- 2012-09-08 pathtracint_application_skeleton.tex
@@
 
-\subsection{The path tracing algorithm}
-
-Path tracing is a global illumination algorithm that is able to render photo-realistic images of a scene viewed by a camera.
-The idea behind the algorithm is very simple: for each pixel of the image, throw rays of light through the pixel that will bounce on the scene to accumulate lighting exchanges.
-It is based on physical laws of radiometry.
+Path tracing is a global illumination algorithm that is able to render photo-realistic images of a virtual scene viewed by a camera.
+The idea behind the algorithm is very simple: for each pixel of the image, throw rays through the pixel that will bounce on the scene to accumulate lighting exchanges.
+In order to get fast convergence the rays must be efficiently distributed through the scene, toward the lights.
+Our idea is to use a curvilinear skeleton of the void of the scene to guide the rays in the correct direction.
+We start by defining some useful concepts related to path tracing and we present our skeleton based path tracing algorithm.
 
@@
 
-Let $S$ be a set of surfaces of $\Rset^3$ that don't overlap ($\forall x \in \Rset^3$ there exists at most one surface of $S$ that contains $x$) and $S_L \subset S$ a set of surface lights.
-$S$ is called the scene.
+Let $S$ be a set of surfaces of $\Rset^3$ that don't overlap and $S_L \subset S$ a set of surface lights. $S$ is called the scene.
 
-The basic radiometric quantity is the \myem{radiance} $L(x \rightarrow \Theta)$ that express the quantity of light energy produce by the surface point $x \in \Rset^3$ towards direction $\Theta \in \Omega_x$ per unit of time per unit area per unit solid angle (expressed in $W.m^{-2}.sr^{-1}$).
-This is the quantity that the eye actually "sees" and that must be computed.
+The basic radiometric quantity is the \myem{radiance} $L(x \rightarrow \Theta)$ that expresses the quantity of light energy produced by the surface point $x \in \Rset^3$ toward direction $\Theta \in \Omega_x$ per unit of time per unit area per unit solid angle (expressed in $W.m^{-2}.sr^{-1}$).
+This is the quantity that the eye actually "sees" and that must be computed for each pixel of the final image.
 
@@
 
-\begin{property}
-
-The radiance remains constant along straight lines in the void.
-
-\end{property}
-
-That property allows us to write the incoming radiance in terms of radiance.
+In the void, incoming radiance can be written in terms of radiance as follow:
 
@@
 It is not null if $x$ is a point of a light source ($x \in S_L$).
-In practice the emitted radiance is given with the set $S_L$ as an input of the algorithm (in can be constant across each of the lights or given by a texture).
+In practice the emitted radiance is given with the set $S_L$ as an input of the algorithm.
 
@@
 
-This equation express an intuitive idea: the radiance produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.
-
-The reflected radiance is computed by taking all the possible directions of the hemisphere around $x$ and integrate the incoming radiance along those directions scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$.
-
-The cosine is used to attenuate the radiance that arrive by a "direction rasante".
+This equation expresses an intuitive idea: the radiance produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.
+The reflected radiance is computed by integrating the incoming radiance scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$ over all the possible directions of the hemisphere around $x$.
+$N_x$ is the surface normal at point $x$.
 
 The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
-It has to respect three properties to be physically coherent:
-
-\begin{itemize}
-\item Helmotz reciprocity: $f_r(x, \Theta \leftrightarrow \Phi) = f_r(x, \Phi \leftrightarrow \Theta)$
-\item Positivity: $f_r(x, \Theta \leftrightarrow \Phi) \geq 0$
-\item Conservation of energy: $\int_{\Omega_x} f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi) d\omega_\Phi \leq 1$
-\end{itemize}
+Intuitively it represents the material properties of the surface at point $x$.
+A mirror doesn't reflect light the same manners a wall do, the BRDF encodes that.
 
@@
 \begin{align*}
-L(x \rightarrow \Theta) &= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi) d\omega_\Phi \\
-&= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi) d\omega_\Phi
+L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi) d\omega_\Phi
 \end{align*}
@@
 \begin{equation*}
-\langle L_r(x \rightarrow \Theta) \rangle = \frac{1}{N} \sum_{i = 1}^{N} \frac{f_r(x, \Theta \leftrightarrow \Phi_i) L(r(x, \Phi_i) \rightarrow -\Phi_i) cos(N_x, \Phi_i)}{p(\Phi_i)}
+\langle L_r(x \rightarrow \Theta) \rangle = \frac{f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi)}{p(\Phi)}
 \end{equation*}
 
-$p$ is a probability density function (pdf) that is used to sample $N$ samples $(\Phi_i)_{i = 1...N}$ of the hemisphere $\Omega_x$.
-
-This estimator is unbiased, it means that the esperance $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$.
-
-The variance of the estimator express its quality. The lower the variance is, the better is the estimator. We have:
+$p$ is a probability density function (pdf) that is used to sample $\Phi$, a direction of the hemisphere $\Omega_x$.
+This estimator is unbiased, it means that the expected value $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$.
+The variance of the estimator express its quality.
+The lower the variance is, the better is the estimator.
+We have:
 
 \begin{equation*}
-\sigma^2[\langle L_r(x \rightarrow \Theta) \rangle] = \frac{1}{N} \int_{\Omega_x} (\frac{f(\Phi)}{p(\Phi)} - L_r(x \rightarrow \Theta))^2 d\omega_\Phi
+\sigma^2[\langle L_r(x \rightarrow \Theta) \rangle] = \int_{\Omega_x} (\frac{f(\Phi)}{p(\Phi)} - L_r(x \rightarrow \Theta))^2 d\omega_\Phi
 \end{equation*}
 
-Then if we increase the number of samples we decrease the variance and the estimator gives better results.
-Another way to decrease the variance is to choose a good pdf.
+The variance clearly depends on the choosen pdf $p$.
 Actually the best is to choose a pdf that match the shape of the function to integrate (ie.
@@
 
-This str	ategy of choosing a good pdf is called \myem{importance sampling} and is over-used in global illumination to improve the convergence of the algorithms.
+This str	ategy of choosing an adapted pdf is called \myem{importance sampling} and is over-used in global illumination to improve the convergence of the algorithms.
 
@@
 
-Donner l'algorithme en pseudo-code
+\begin{algorithm}[h!]
+\label{algo::incoming_radiance}
+\caption[incoming_radiance]{Incoming radiance $L(x \leftarrow \Phi)$}
+\KwData{An incoming direction $\Phi$, a point $x$} 
+\KwResult{$L(x \leftarrow \Phi)$}
+\Return $L(r(x, \Phi) \rightarrow -\Phi)$\;
+\end{algorithm}
+
+\begin{algorithm}[h!]
+\label{algo::radiance}
+\caption[radiance]{Radiance $L(x \rightarrow \Theta)$}
+\KwData{An outcoming direction $\Theta$, a point $x$} 
+\KwResult{$L(x \rightarrow \Theta)$}
+$L \gets L_e(x \rightarrow \Theta)$\;
+\If{continue the recursion} {
+	$\Phi \gets sample(\Omega_x, p)$ \\
+	$L \gets L + \frac{f_r(x, \Phi, \Theta) L(x \leftarrow \Phi) cos(N_x, \Phi)}{p(\Phi)}$\;
+}
+\Return L\;
+\end{algorithm}
+
+This algorithm is called for each pixel of the final image to compute. To improve the quality of the result we can throw multiple rays accross each pixel and computing the mean of the results.
+
+In the radiance algorithm the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$. We will discuss what pdf we use in the next section.
 
@@
 
-Décrire la stratégie d'échantillonage basée sur le squelette
+The key idea behind our method is that light travels in the void of the scene.
+A curvilinear skeleton of that void gives us enough informations to compute some importance directions that tells us where the light comes from.
+Given that directions we can construct a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$
 
-\subsection{Multiple importance sampling}
+Here is the integrand of the integral estimated with Monte-Carlo:
 
-Décrire comment combiner les stratégies brdf + skeleton
+\begin{equation*}
+f_r(x, \Phi, \Theta) L(x \leftarrow \Phi) cos(N_x, \Phi)
+\end{equation*}
+
+It's a product of three functions that can be efficiently sample individually but that are hard to sample when they are combined.
+The most common strategy used to sample $\Omega_x$ is to use the BRDF (combined with the cosine) because we know it: it is given as an input of the algorithm, for exemple with textures.
+
+$L$ is rarelly used because we cannot evaluate it in constant time.
+Remember that it represents the distribution of light in the scene.
+It takes high values when $\Phi$ carries a lot of energy.
+Out method gives a way to sample according to $L$.
+
+\subsubsection{Construction of the importance points}
+
+The skeleton of the void is computed with our algorithm and is converted to a graph (the nodes are 3D positions and the edges represents the topology of the scene), given as an input of our path tracing.
+We then compute a set of points called \myem{importance points}.
+These points will be used to sample $\Omega_x$ in the path tracing algorithm.
+For each node of the skeleton $n$, one importance point $imp_n$ is computed.
+
+Let $L$ be a light of the scene ($L \in S_L$) and $n_L$ the shortest visible node by $L$.
+We first compute a tree of shortest paths along the skeleton with $n_L$ as root.
+We use the Dijkstra algorithm with a distance based on visibility from $n_L$: We put $d(n, m) = 1$ if $m$ is visible from $n_L$ and $d(n, m) = 10$ if not.
+It results that the paths that are enlighted by $L$ will have a shorter distance and will guide us to the light faster than the dark ones.
+
+Let $n$ be a node of the skeleton and $V_n$ the set of visible nodes from $n$ along the shortest path toward $n_L$.
+The importance point $imp_n$ associated to $n$ is computed by taking the barycenter of $V_n$.
+
+\subsubsection{Sampling according to $L$}
+
+Given a point $x$ and a direction $\Theta$ we want to compute $L(x \rightarrow \Theta)$ and then sample the hemisphere $\Omega_x$.
+We search for the nearest skeleton node $n$ and its importance point $imp_n$.
+We sample the hemisphere with a power-cosine pdf center on $\overrightarrow{ximp_n}$:
+
+\begin{equation*}
+p_{skel}(\Phi) = \frac{s + 1}{2\pi} cos^s \alpha
+\end{equation*}
+
+with $\alpha$ the angle between $\overrightarrow{ximp_n}$ and $\Phi$ and $s$ a parameter called skeleton strength. The more $s$ is high, the more we sample close $\overrightarrow{ximp_n}$.
+
+\subsubsection{Results}
+
+\subsection{Discussion}
+
+We presented an application of our skeletonization algorithm for global illumination.
+Our strategy work wells, especially in dark areas.
+It reduces noise by taking into account the distribution of energy in the scene.
+One of the problem is that we take into account only one importance direction for sampling, we totally ignore the BRDF and the cosine factor.
+A solution called Multiple Importance Sampling allow us to combine different strategies into one.
+We will not describe this strategy because it's not directly related to skeletonization, but it have allowed us to improve our results and remove some artifact of our method.

### 2012-09-11 pathtracint_application_skeleton.tex
Corrections partie application (2)
Removing some parts in path tracing

@@
 Path tracing is a global illumination algorithm that is able to render photo-realistic images of a virtual scene viewed by a camera.
-The idea behind the algorithm is very simple: for each pixel of the image, throw rays through the pixel that will bounce on the scene to accumulate lighting exchanges.
+The idea behind the algorithm is very elegant: for each pixel of the image, throw rays through the pixel that will bounce on the scene to accumulate lighting exchanges.
 In order to get fast convergence the rays must be efficiently distributed through the scene, toward the lights.
@@
 
-\subsubsection{Basic definitions of radiometry}
+%\subsection{Basic definitions of radiometry}
+%
+%Let $S$ be a set of surfaces of $\Rset^3$ that don't overlap and $S_L \subset S$ a set of surface lights. $S$ is called the scene.
+%
+%The basic radiometric quantity is the \myem{radiance} $L(x \rightarrow \Theta)$ that expresses the quantity of light energy produced by the surface point $x \in \Rset^3$ toward direction $\Theta \in \Omega_x$ per unit of time per unit area per unit solid angle (expressed in $W.m^{-2}.sr^{-1}$).
+%This is the quantity that the eye actually "sees" and that must be computed for each pixel of the final image.
+%
+%% image radiance
+%
+%We can define another quantity related to radiance called \myem{incoming radiance} $L(x \leftarrow \Phi)$ that represents the radiance that reach $x$ from direction $\Phi$.
+%
+%In the void, incoming radiance can be written in terms of radiance as follow:
+%
+%\begin{equation*}
+%L(x \leftarrow \Phi) = L(r(x, \Phi) \rightarrow -\Phi)
+%\end{equation*}
+%
+%$r(x, \Phi)$ is the surface point seen by $x$ in the direction $\Phi$. It can be defined formally by:
+%
+%\begin{align*}
+%r(x, \Phi) &= x + t_{min}\Phi \\
+%t_{min} &= min \lbrace t \in (0, +\infty]~ |~ x + t\Phi \in S \rbrace
+%\end{align*}
+%
+%If $t_{min} = \infty$ then $r(x, \Phi)$ is undefined and $L(x \leftarrow \Phi) = 0$.
+We call $r$ the \myem{raytrace} function.
+In practice it is implemented by throwing a ray in the scene and looking for the first intersection.
+%
+%A third quantity related to radiance is \myem{emitted radiance} $L_e(x \rightarrow \Theta)$.
+It is not null if $x$ is a point of a light source ($x \in S_L$).
+In practice the emitted radiance is given with the set $S_L$ as an input of the algorithm.
+%
+%The radiance is the solution of an equation called the Rendering Equation:
+%
+%\begin{align*}
+%L(x \rightarrow \Theta) &= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi) d\omega_\Phi \\
+%&= L_e(x \rightarrow \Theta) + L_r(x \rightarrow \Theta)
+%\end{align*}
+%
+%This equation expresses an intuitive idea: the radiance produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.
+The reflected radiance is computed by integrating the incoming radiance scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$ over all the possible directions of the hemisphere around $x$.
+$N_x$ is the surface normal at point $x$.
+%
+%The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
+Intuitively it represents the material properties of the surface at point $x$.
+A mirror doesn't reflect light the same manners a wall do, the BRDF encodes that.
+%
+%It's actually the rendering equation that a path tracer try to resolve.
+Let $O$ be the origin of the camera and $I$ the image we want to compute (a rectangle embedded in $\Rset^3$).
+Then the path tracing algorithm compute an approximation of the radiances $L(O \leftarrow \vec{OP})$, $\forall P \in I$.
 
-Let $S$ be a set of surfaces of $\Rset^3$ that don't overlap and $S_L \subset S$ a set of surface lights. $S$ is called the scene.
+\subsection{The path tracing}
 
-The basic radiometric quantity is the \myem{radiance} $L(x \rightarrow \Theta)$ that expresses the quantity of light energy produced by the surface point $x \in \Rset^3$ toward direction $\Theta \in \Omega_x$ per unit of time per unit area per unit solid angle (expressed in $W.m^{-2}.sr^{-1}$).
-This is the quantity that the eye actually "sees" and that must be computed for each pixel of the final image.
-
-% image radiance
-
-We can define another quantity related to radiance called \myem{incoming radiance} $L(x \leftarrow \Phi)$ that represents the radiance that reach $x$ from direction $\Phi$.
-
-In the void, incoming radiance can be written in terms of radiance as follow:
-
-\begin{equation*}
-L(x \leftarrow \Phi) = L(r(x, \Phi) \rightarrow -\Phi)
-\end{equation*}
-
-$r(x, \Phi)$ is the surface point seen by $x$ in the direction $\Phi$. It can be defined formally by:
+A CHANGER : INTRODUIRE L'EQUATION DU RENDU : du genre : for each pixel we seek the intersection point.
+At this point, to obtain the luminosity, we must solve the rendering equation.
+The rendering equation is a recursive integral equation.
+The integrand of the equation contains the radiance function that we must compute:
 
 \begin{align*}
-r(x, \Phi) &= x + t_{min}\Phi \\
-t_{min} &= min \lbrace t \in (0, +\infty]~ |~ x + t\Phi \in S \rbrace
+L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi) d\omega_\Phi
 \end{align*}
-
-If $t_{min} = \infty$ then $r(x, \Phi)$ is undefined and $L(x \leftarrow \Phi) = 0$.
-We call $r$ the \myem{raytrace} function.
-In practice it is implemented by throwing a ray in the scene and looking for the first intersection.
-
-A third quantity related to radiance is \myem{emitted radiance} $L_e(x \rightarrow \Theta)$.
-It is not null if $x$ is a point of a light source ($x \in S_L$).
-In practice the emitted radiance is given with the set $S_L$ as an input of the algorithm.
-
-The radiance is the solution of an equation called the Rendering Equation:
-
-\begin{align*}
-L(x \rightarrow \Theta) &= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi) d\omega_\Phi \\
-&= L_e(x \rightarrow \Theta) + L_r(x \rightarrow \Theta)
-\end{align*}
-
-This equation expresses an intuitive idea: the radiance produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.
+INTRODUIRE BRIEVEMENT la luminance du style : L being the light flux from point $x$ toward direction $\Theta$.
+This equation expresses an intuitive idea: the radiance $L$ produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.
 The reflected radiance is computed by integrating the incoming radiance scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$ over all the possible directions of the hemisphere around $x$.
 $N_x$ is the surface normal at point $x$.
-
 The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
@@
 
-It's actually the rendering equation that a path tracer try to resolve.
-Let $O$ be the origin of the camera and $I$ the image we want to compute (a rectangle embedded in $\Rset^3$).
-Then the path tracing algorithm compute an approximation of the radiances $L(O \leftarrow \vec{OP})$, $\forall P \in I$.
-
-\subsubsection{Solving the rendering equation}
-
-The rendering equation is a recursive integral equation. The integrand of the equation contains the radiance function that we must compute:
-
-\begin{align*}
-L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi) d\omega_\Phi
-\end{align*}
-
-In practive it is impossible to compute anatically the integral, it has to be estimated.
-The most commonly used method is the Monte-Carlo integration that provide the following estimator for the integral:
-
+In practice it is impossible to compute anatically the integral, it has to be estimated. RAJOUT : cannot send rays everywhere.
+The most commonly used method is the Monte-Carlo integration that provides the following estimator for the integral:
 \begin{equation*}
@@
 The variance of the estimator express its quality.
-The lower the variance is, the better is the estimator.
-We have:
-
-\begin{equation*}
-\sigma^2[\langle L_r(x \rightarrow \Theta) \rangle] = \int_{\Omega_x} (\frac{f(\Phi)}{p(\Phi)} - L_r(x \rightarrow \Theta))^2 d\omega_\Phi
-\end{equation*}
-
 The variance clearly depends on the choosen pdf $p$.
-Actually the best is to choose a pdf that match the shape of the function to integrate (ie.
+Actually the best is to choose a pdf that matches the shape of the function to integrate (ie.
 give high density to samples that have high values for the function and low density to samples that have low values).
 
-This str	ategy of choosing an adapted pdf is called \myem{importance sampling} and is over-used in global illumination to improve the convergence of the algorithms.
+This strategy of choosing an adapted pdf is called \myem{importance sampling} and is over-used in global illumination to improve the convergence of the algorithms.
 
-\subsubsection{The algorithm}
-
-\begin{algorithm}[h!]
-\label{algo::incoming_radiance}
-\caption[incoming_radiance]{Incoming radiance $L(x \leftarrow \Phi)$}
-\KwData{An incoming direction $\Phi$, a point $x$} 
-\KwResult{$L(x \leftarrow \Phi)$}
-\Return $L(r(x, \Phi) \rightarrow -\Phi)$\;
-\end{algorithm}
-
-\begin{algorithm}[h!]
-\label{algo::radiance}
-\caption[radiance]{Radiance $L(x \rightarrow \Theta)$}
-\KwData{An outcoming direction $\Theta$, a point $x$} 
-\KwResult{$L(x \rightarrow \Theta)$}
-$L \gets L_e(x \rightarrow \Theta)$\;
-\If{continue the recursion} {
-	$\Phi \gets sample(\Omega_x, p)$ \\
-	$L \gets L + \frac{f_r(x, \Phi, \Theta) L(x \leftarrow \Phi) cos(N_x, \Phi)}{p(\Phi)}$\;
-}
-\Return L\;
-\end{algorithm}
-
+INTRODUIRE BRIEVEMENT L'ALGORITHME...
 This algorithm is called for each pixel of the final image to compute. To improve the quality of the result we can throw multiple rays accross each pixel and computing the mean of the results.
@@
 In the radiance algorithm the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$. We will discuss what pdf we use in the next section.
+
+PARLER DE RAYON QUI TOUCHE PAS LUMIERE + BRUIT
 
@@
 
-It's a product of three functions that can be efficiently sample individually but that are hard to sample when they are combined.
+It is a product of three functions that can be efficiently sample individually but that are hard to sample when they are combined.
 The most common strategy used to sample $\Omega_x$ is to use the BRDF (combined with the cosine) because we know it: it is given as an input of the algorithm, for exemple with textures.

### 2012-09-11 pathtracint_application_skeleton.tex
Ajout images

@@
 
-A CHANGER : INTRODUIRE L'EQUATION DU RENDU : du genre : for each pixel we seek the intersection point.
+Let $O$ be the origin of the camera in the scene.
+For each pixel $P$ of the final image we search for the nearest intersection point $x$ of the ray $(O, \overrightarrow{OP} = -\Theta)$ with the scene.
 At this point, to obtain the luminosity, we must solve the rendering equation.
@@
 \end{align*}
-INTRODUIRE BRIEVEMENT la luminance du style : L being the light flux from point $x$ toward direction $\Theta$.
+
+$L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$.
+$L_e$ is the emitted radiance that is not zero only on light sources.
+$r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$.
+The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
+Intuitively it represents the material properties of the surface at point $x$.
+A mirror doesn't reflect light the same manners a wall do, the BRDF encodes that.
+
 This equation expresses an intuitive idea: the radiance $L$ produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.
@@
 $N_x$ is the surface normal at point $x$.
-The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
-Intuitively it represents the material properties of the surface at point $x$.
-A mirror doesn't reflect light the same manners a wall do, the BRDF encodes that.
 
@@
 
+Here is a result of our algorithm:
+
+\begin{center}
+\includegraphics[scale=0.2]{images/skel_based_rt.png}
+\end{center}
+
+We observe a net decreasing of the noise: we get a faster convergence to the true image.
+
 \subsection{Discussion}

### 2012-09-11 pathtracint_application_skeleton.tex
Corrections app

@@
 The variance of the estimator express its quality.
-The variance clearly depends on the choosen pdf $p$.
+The variance of the estimator depends on the choosen pdf $p$.
 Actually the best is to choose a pdf that matches the shape of the function to integrate (ie.
@@
 
-INTRODUIRE BRIEVEMENT L'ALGORITHME...
-This algorithm is called for each pixel of the final image to compute. To improve the quality of the result we can throw multiple rays accross each pixel and computing the mean of the results.
+The implantation of $L$ is as follow: first we sample a random direction $\Phi$ based on the pdf $p$.
+Then we apply the estimator that call $L$ recursively to compute $L_r$.
+Then we return the sum of the emitted radiance $L_e$ and the result for $L_r$.
+The recursion stops when a maximal number of bounces have been reach.
 
-In the radiance algorithm the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$. We will discuss what pdf we use in the next section.
+This algorithm is called for each pixel of the final image to compute. To improve the quality of the final image we can throw multiple rays accross each pixel and compute the mean of the results.
 
-PARLER DE RAYON QUI TOUCHE PAS LUMIERE + BRUIT
+In the radiance algorithm the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$.
+A bad pdf picks directions that don't reach the light before the end of the recursion and results in a lot of noise in the final image.
+We will discuss the choice of $p$ in the next section.
 

### 2012-09-11 pathtracint_application_skeleton.tex
Ajout ref et changement images

@@
 For each pixel $P$ of the final image we search for the nearest intersection point $x$ of the ray $(O, \overrightarrow{OP} = -\Theta)$ with the scene.
-At this point, to obtain the luminosity, we must solve the rendering equation.
+At this point, to obtain the luminosity, we must solve the rendering equation \cite{Ka86}.
 The rendering equation is a recursive integral equation.
@@
 
-In practice it is impossible to compute anatically the integral, it has to be estimated. RAJOUT : cannot send rays everywhere.
+In practice it is impossible to compute anatically the integral, it has to be estimated.
 The most commonly used method is the Monte-Carlo integration that provides the following estimator for the integral:
@@
 
-\subsubsection{Results}
+\clearpage
+\subsection{Results and discussion}
 
-Here is a result of our algorithm:
+We present the results of our algorithm applied to four scenes.
+the first row shows our result and the second row shows a standard path tracing with the same input parameters.
+On each picture a MSE (mean square error) value with a reference image is shown.
 
 \begin{center}
-\includegraphics[scale=0.2]{images/skel_based_rt.png}
+\hspace*{-40px}
+\includegraphics[scale=0.2]{images/four_scenes.png}
 \end{center}
 
-We observe a net decreasing of the noise: we get a faster convergence to the true image.
-
-\subsection{Discussion}
+We observe a net decreasing of the noise for our algorithm, especially in dark areas.
+The MSE values are lower for the top pictures, meaning that we converge faster to the reference image than standard path tracing does.
 
 We presented an application of our skeletonization algorithm for global illumination.
-Our strategy work wells, especially in dark areas.
-It reduces noise by taking into account the distribution of energy in the scene.
+Our strategy works well, it reduces noise by taking into account the distribution of energy in the scene.
 One of the problem is that we take into account only one importance direction for sampling, we totally ignore the BRDF and the cosine factor.
 A solution called Multiple Importance Sampling allow us to combine different strategies into one.
-We will not describe this strategy because it's not directly related to skeletonization, but it have allowed us to improve our results and remove some artifact of our method.
+We will not describe this strategy because it's not directly related to skeletonization, but it has allowed us to improve our results and remove some artifacts of our method.

### 2012-09-11 make.sh

@@ -1,0 +1,2 @@
+./clean.sh ; pdflatex $1.tex ; bibtex $1 ; pdflatex $1.tex ; pdflatex $1.tex
+

### 2012-09-11 clean.sh
Ajout clean.sh

@@ -1,0 +1,1 @@
+rm *~ *.backup *.aux *.toc *.blg *.log *.out *.bbl *.pdf *.nav *.pyg *.snm *.lof 2> /dev/null

### 2012-09-11 application_to_path_tracing.tex
Renommage fichier tex application

--- 2012-09-11 pathtracint_application_skeleton.tex
@@
 For each pixel $P$ of the final image we search for the nearest intersection point $x$ of the ray $(O, \overrightarrow{OP} = -\Theta)$ with the scene.
-At this point, to obtain the luminosity, we must solve the rendering equation \cite{Ka86}.
-The rendering equation is a recursive integral equation.
+At this point, to obtain the luminosity, we must solve the rendering equation \cite{Ka86} which is a recursive integral equation.
 The integrand of the equation contains the radiance function that we must compute:
@@
 $L_e$ is the emitted radiance that is not zero only on light sources.
-$r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$.
+$r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$ of the hemisphere $\Omega_x$.
+$N_x$ is the surface normal at point $x$.
 The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
 Intuitively it represents the material properties of the surface at point $x$.
-A mirror doesn't reflect light the same manners a wall do, the BRDF encodes that.
+A mirror does not reflect light the same manners a wall do and the BRDF describes that kind of behavior.
 
@@
 The reflected radiance is computed by integrating the incoming radiance scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$ over all the possible directions of the hemisphere around $x$.
-$N_x$ is the surface normal at point $x$.
 
-In practice it is impossible to compute anatically the integral, it has to be estimated.
-The most commonly used method is the Monte-Carlo integration that provides the following estimator for the integral:
+The most commonly used method to compute the integral is the Monte-Carlo integration that provides the following estimator:
 \begin{equation*}
@@
 
-$p$ is a probability density function (pdf) that is used to sample $\Phi$, a direction of the hemisphere $\Omega_x$.
-This estimator is unbiased, it means that the expected value $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$.
-The variance of the estimator express its quality.
-The variance of the estimator depends on the choosen pdf $p$.
+where $p$ is a probability density function (pdf) that is used to sample $\Phi$.
+This estimator is unbiased, meaning that the expected value $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$.
+The variance of the estimator expresses its quality and depends on the choosen pdf $p$.
 Actually the best is to choose a pdf that matches the shape of the function to integrate (ie.
@@
 
-The implantation of $L$ is as follow: first we sample a random direction $\Phi$ based on the pdf $p$.
+The algorithm to compute $L$ is as follow: first we sample a random direction $\Phi$ based on the pdf $p$.
 Then we apply the estimator that call $L$ recursively to compute $L_r$.
-Then we return the sum of the emitted radiance $L_e$ and the result for $L_r$.
-The recursion stops when a maximal number of bounces have been reach.
+Finally we return the sum of the emitted radiance $L_e$ and the result for $L_r$.
+The recursion stops when a maximal number of bounces has been reach.
 
-This algorithm is called for each pixel of the final image to compute. To improve the quality of the final image we can throw multiple rays accross each pixel and compute the mean of the results.
+This computation is done multiple time for each pixel (at different positions over the pixel).
+We average the results and get an estimation of the mean radiance passing through the pixel in direction of the camera's origin.
 
-In the radiance algorithm the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$.
+In the radiance computation the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$.
 A bad pdf picks directions that don't reach the light before the end of the recursion and results in a lot of noise in the final image.
@@
 The key idea behind our method is that light travels in the void of the scene.
-A curvilinear skeleton of that void gives us enough informations to compute some importance directions that tells us where the light comes from.
-Given that directions we can construct a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$
+As stated in the introduction, a curvilinear skeleton of the void gives us enough informations to compute some importance directions where light comes from.
+Given these directions we can construct a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$
 
-Here is the integrand of the integral estimated with Monte-Carlo:
+The integrand of $L_r$ is:
 
 \begin{equation*}
-f_r(x, \Phi, \Theta) L(x \leftarrow \Phi) cos(N_x, \Phi)
+f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi)
 \end{equation*}
@@
 
-$L$ is rarelly used because we cannot evaluate it in constant time.
+$L$ is rarelly used because its evaluation cannot be done in constant time.
 Remember that it represents the distribution of light in the scene.
 It takes high values when $\Phi$ carries a lot of energy.
-Out method gives a way to sample according to $L$.
+Our method gives a way to sample according to $L$.
 
@@
 
-The skeleton of the void is computed with our algorithm and is converted to a graph (the nodes are 3D positions and the edges represents the topology of the scene), given as an input of our path tracing.
+The skeleton of the void is computed with our algorithm \ref{sec::skeletonization} and is converted to a graph (the nodes are 3D positions and the edges represents the topology of the scene), given as an input of our path tracing.
 We then compute a set of points called \myem{importance points}.
@@
 
-Let $n$ be a node of the skeleton and $V_n$ the set of visible nodes from $n$ along the shortest path toward $n_L$.
-The importance point $imp_n$ associated to $n$ is computed by taking the barycenter of $V_n$.
+Let $n$ be a node of the skeleton and $V_n$ the set of visible nodes from $n$ along the shortest path toward $n_L$. The importance point $imp_n$ associated to $n$ is the barycenter of $V_n$.
 
@@
 We search for the nearest skeleton node $n$ and its importance point $imp_n$.
-We sample the hemisphere with a power-cosine pdf center on $\overrightarrow{ximp_n}$:
+We sample the hemisphere with a power-cosine pdf centered on $\overrightarrow{ximp_n}$:
 
@@
 
-with $\alpha$ the angle between $\overrightarrow{ximp_n}$ and $\Phi$ and $s$ a parameter called skeleton strength. The more $s$ is high, the more we sample close $\overrightarrow{ximp_n}$.
+with $\alpha$ the angle between $\overrightarrow{ximp_n}$ and $\Phi$, $s$ being a parameter called skeleton strength. The higher $s$ is, the more we sample close $\overrightarrow{ximp_n}$.
 
@@
 
-We present the results of our algorithm applied to four scenes.
+We present some results of our algorithm applied to four scenes.
 the first row shows our result and the second row shows a standard path tracing with the same input parameters.
@@
 \begin{center}
-\hspace*{-40px}
-\includegraphics[scale=0.2]{images/four_scenes.png}
+\includegraphics[scale=0.17]{images/four_scenes.png}
 \end{center}
 
-We observe a net decreasing of the noise for our algorithm, especially in dark areas.
+We observe a decrease of the noise for our algorithm, especially in dark areas.
 The MSE values are lower for the top pictures, meaning that we converge faster to the reference image than standard path tracing does.
@@
 Our strategy works well, it reduces noise by taking into account the distribution of energy in the scene.
-One of the problem is that we take into account only one importance direction for sampling, we totally ignore the BRDF and the cosine factor.
-A solution called Multiple Importance Sampling allow us to combine different strategies into one.
-We will not describe this strategy because it's not directly related to skeletonization, but it has allowed us to improve our results and remove some artifacts of our method.
+One of the problem is that we take into account only one importance direction for sampling, totally ignoring the BRDF and the cosine factor.
+A solution called Multiple Importance Sampling will allow us to combine different strategies into one.

### 2012-09-11 application_to_path_tracing.tex
Petite correction de reference

@@
 
-The skeleton of the void is computed with our algorithm \ref{sec::skeletonization} and is converted to a graph (the nodes are 3D positions and the edges represents the topology of the scene), given as an input of our path tracing.
+The skeleton of the void is computed with our algorithm (section \ref{sec::skeletonization}) and is converted to a graph (the nodes are 3D positions and the edges represents the topology of the scene), given as an input of our path tracing.
 We then compute a set of points called \myem{importance points}.

### 2012-09-12 application_to_path_tracing.tex
Petites corrections suite à la relecture de Michel

--- 2012-09-11 application_to_path_tracing.tex
@@
 $N_x$ is the surface normal at point $x$.
-The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and express the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
+The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and expresses the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
 Intuitively it represents the material properties of the surface at point $x$.
-A mirror does not reflect light the same manners a wall do and the BRDF describes that kind of behavior.
+A mirror does not reflect light in the same manner as a wall does and the BRDF describes that kind of behavior.
 
@@
 
-This strategy of choosing an adapted pdf is called \myem{importance sampling} and is over-used in global illumination to improve the convergence of the algorithms.
+This strategy of choosing an adapted pdf is called \myem{importance sampling} and is used in global illumination to improve the convergence speed of the algorithms.
 
 The algorithm to compute $L$ is as follow: first we sample a random direction $\Phi$ based on the pdf $p$.
-Then we apply the estimator that call $L$ recursively to compute $L_r$.
+Then we apply the estimator that calls $L$ recursively to compute $L_r$.
 Finally we return the sum of the emitted radiance $L_e$ and the result for $L_r$.
-The recursion stops when a maximal number of bounces has been reach.
+The recursion stops when a maximal number of bounces has been reached.
 
-This computation is done multiple time for each pixel (at different positions over the pixel).
+This computation is done multiple times for each pixel (at different positions over the pixel).
 We average the results and get an estimation of the mean radiance passing through the pixel in direction of the camera's origin.
@@
 The key idea behind our method is that light travels in the void of the scene.
-As stated in the introduction, a curvilinear skeleton of the void gives us enough informations to compute some importance directions where light comes from.
-Given these directions we can construct a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$
+As stated in the introduction, a curvilinear skeleton of the void gives us enough information to compute some importance directions where light comes from.
+Given these directions we can construct a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$.
 
@@
 \begin{equation*}
-f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi)
+f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi).
 \end{equation*}
 
-It is a product of three functions that can be efficiently sample individually but that are hard to sample when they are combined.
+It is a product of three functions that can be efficiently sampled individually but that are hard to sample when they are combined.
 The most common strategy used to sample $\Omega_x$ is to use the BRDF (combined with the cosine) because we know it: it is given as an input of the algorithm, for exemple with textures.
 
-$L$ is rarelly used because its evaluation cannot be done in constant time.
+The term $L$ is rarely used because its evaluation cannot be done in constant time.
 Remember that it represents the distribution of light in the scene.
@@
 
-The skeleton of the void is computed with our algorithm \ref{sec::skeletonization} and is converted to a graph (the nodes are 3D positions and the edges represents the topology of the scene), given as an input of our path tracing.
+The skeleton of the void is computed with our algorithm (section \ref{sec::skeletonization}) and is converted to a graph (the nodes are 3D positions and the edges represent the topology of the scene), given as an input of our path tracing.
 We then compute a set of points called \myem{importance points}.
@@
 
-Let $L$ be a light of the scene ($L \in S_L$) and $n_L$ the shortest visible node by $L$.
+Let $L$ be a light source of the scene ($L \in S_L$) and $n_L$ the nearest node that is visible by $L$.
 We first compute a tree of shortest paths along the skeleton with $n_L$ as root.
@@
 
-\clearpage
 \subsection{Results and discussion}
@@
 We present some results of our algorithm applied to four scenes.
-the first row shows our result and the second row shows a standard path tracing with the same input parameters.
+The first row shows our result and the second row shows a standard path tracing with the same input parameters.
 On each picture a MSE (mean square error) value with a reference image is shown.
@@
 Our strategy works well, it reduces noise by taking into account the distribution of energy in the scene.
-One of the problem is that we take into account only one importance direction for sampling, totally ignoring the BRDF and the cosine factor.
+One of the problems is that we take into account only one importance direction for sampling, totally ignoring the BRDF and the cosine factor.
 A solution called Multiple Importance Sampling will allow us to combine different strategies into one.

### 2012-09-13 application_to_path_tracing.tex
Nouvelles corrections
Correction + modif partie path tracing + images

@@
 \section{Application to path tracing}
+\label{sec::appli_pathtrace}
 
-Path tracing is a global illumination algorithm that is able to render photo-realistic images of a virtual scene viewed by a camera.
-The idea behind the algorithm is very elegant: for each pixel of the image, throw rays through the pixel that will bounce on the scene to accumulate lighting exchanges.
-In order to get fast convergence the rays must be efficiently distributed through the scene, toward the lights.
-Our idea is to use a curvilinear skeleton of the void of the scene to guide the rays in the correct direction.
-We start by defining some useful concepts related to path tracing and we present our skeleton based path tracing algorithm.
+In this section, we explain how to use the curvilinear skeleton of the voids of the scene in order to enhance the path tracing algorithm.
+We start by defining some basic elements related to path tracing and we present our skeleton based path tracing algorithm.
 
@@
 Let $O$ be the origin of the camera in the scene.
-For each pixel $P$ of the final image we search for the nearest intersection point $x$ of the ray $(O, \overrightarrow{OP} = -\Theta)$ with the scene.
-At this point, to obtain the luminosity, we must solve the rendering equation \cite{Ka86} which is a recursive integral equation.
+For each pixel $P$ of the final image, let $x$ be the nearest intersection point of the ray $(O, \overrightarrow{OP} = -\Theta)$ with an element of the scene.
+To obtain the luminosity, we must solve the rendering equation \cite{Ka86} which is a recursive integral equation.
 The integrand of the equation contains the radiance function that we must compute:
@@
 \begin{align*}
-L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi) d\omega_\Phi
+L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) d\omega_\Phi
 \end{align*}
@@
 $L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$.
-$L_e$ is the emitted radiance that is not zero only on light sources.
-$r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$ of the hemisphere $\Omega_x$.
-$N_x$ is the surface normal at point $x$.
-The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and expresses the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$.
+$L_e$ is the emitted radiance (non null only on light sources).
+The point $r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$ of the hemisphere $\Omega_x$.
+The $f_s$ function expresses how much luminosity is exchanged, on point $x$, between an incoming ray $\Phi$ and the outgoing ray $\Theta$.
 Intuitively it represents the material properties of the surface at point $x$.
-A mirror does not reflect light in the same manner as a wall does and the BRDF describes that kind of behavior.
 
-This equation expresses an intuitive idea: the radiance $L$ produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.
-The reflected radiance is computed by integrating the incoming radiance scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$ over all the possible directions of the hemisphere around $x$.
+This integrand expresses an intuitive idea: the reflected radiance from $x$ towards the camera is the result of computing all incoming luminosity on $x$ scaled by a factor which depends on the material on $x$ and the angle of the ray going from the camera to $x$.
+The most common method used to estimate the integral is the Monte-Carlo integration which provides the following:
 
-The most commonly used method to compute the integral is the Monte-Carlo integration that provides the following estimator:
 \begin{equation*}
-\langle L_r(x \rightarrow \Theta) \rangle = \frac{f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi)}{p(\Phi)}
+\langle L_r(x \rightarrow \Theta) \rangle = \frac{f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi)}{p(\Phi)}
 \end{equation*}
@@
 The variance of the estimator expresses its quality and depends on the choosen pdf $p$.
-Actually the best is to choose a pdf that matches the shape of the function to integrate (ie.
+The best choice is a pdf that matches the shape of the function to integrate (ie.
 give high density to samples that have high values for the function and low density to samples that have low values).
+The strategy of choosing an adapted pdf is called \myem{importance sampling} and is used in global illumination to improve the convergence speed of the algorithms.
 
-This strategy of choosing an adapted pdf is called \myem{importance sampling} and is used in global illumination to improve the convergence speed of the algorithms.
-
-The algorithm to compute $L$ is as follow: first we sample a random direction $\Phi$ based on the pdf $p$.
-Then we apply the estimator that calls $L$ recursively to compute $L_r$.
-Finally we return the sum of the emitted radiance $L_e$ and the result for $L_r$.
-The recursion stops when a maximal number of bounces has been reached.
+The algorithm computes $L$ as follow: first, it chooses one random direction $\Phi$ based on the pdf $p$.
+Then, it applies the estimator by calling $L$ recursively to compute $L_r$ (shoots a new ray in the direction $\Phi$, computes the new hit point with the scene, and compute the new radiance at this point).
+Finally, it returns the sum of the emitted radiance $L_e$ and the result for $L_r$.
+The recursion stops either when a light or a maximal number of bounces has been reached.
+%ASK: euh, c'est L ou L_r%
+%c'est bon de dire que ça s'arrete au light%
 
 This computation is done multiple times for each pixel (at different positions over the pixel).
-We average the results and get an estimation of the mean radiance passing through the pixel in direction of the camera's origin.
-
-In the radiance computation the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$.
-A bad pdf picks directions that don't reach the light before the end of the recursion and results in a lot of noise in the final image.
-We will discuss the choice of $p$ in the next section.
+We average the results and get an estimation of the mean radiance passing through the pixel and heading towards the camera.
+%at different positions over the pixel  ???%
+A bad pdf would lead in picking directions that don't reach the light before the end of the recursion, and produce results with a lot of noise in the final image.
+In the next part, we explain how to produce a pdf based on the curvilinear skeleton.
 
@@
 
-The key idea behind our method is that light travels in the void of the scene.
-As stated in the introduction, a curvilinear skeleton of the void gives us enough information to compute some importance directions where light comes from.
-Given these directions we can construct a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$.
+As stated in the introduction, a curvilinear skeleton of the void (with some preprocessing achieved on it) gives us information on which directions the light comes from.
+Given these directions, we can build a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$.
 
 The integrand of $L_r$ is:
+
+%a revoir, je ne comprends pas tout
 
@@
 
-The skeleton of the void is computed with our algorithm (section \ref{sec::skeletonization}) and is converted to a graph (the nodes are 3D positions and the edges represent the topology of the scene), given as an input of our path tracing.
-We then compute a set of points called \myem{importance points}.
-These points will be used to sample $\Omega_x$ in the path tracing algorithm.
+The skeleton of the voids of the scene is computed using Alg.~\ref{algo::3d_curvskel} and is converted to a graph (the nodes are the 0-faces and the edges are the 1-faces).
+We then compute a set of \myem{importance points}: these points will be used to sample $\Omega_x$ in the path tracing algorithm.
 For each node of the skeleton $n$, one importance point $imp_n$ is computed.
+Intuitively, the importance point associated to $n$ gives us a direction to follow in order to find a light source.
+%TODO a verifier quand meme
 
+%La, pas clair !!!
 Let $L$ be a light source of the scene ($L \in S_L$) and $n_L$ the nearest node that is visible by $L$.
@@
 
-We observe a decrease of the noise for our algorithm, especially in dark areas.
+We observe that images produced by our algorithm have less noise than the ones produced by the regular path tracing (for a same number of rays shot per pixel).
 The MSE values are lower for the top pictures, meaning that we converge faster to the reference image than standard path tracing does.

### 2012-09-14 application_to_path_tracing.tex
Changing to the new results image
Modifs

@@
 \begin{align*}
-L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) d\omega_\Phi
+L(x \rightarrow \Theta) &= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) d\omega_\Phi \\
+&= L_e(x \rightarrow \Theta) + L_r(x \rightarrow \Theta)
 \end{align*}
@@
 Finally, it returns the sum of the emitted radiance $L_e$ and the result for $L_r$.
-The recursion stops either when a light or a maximal number of bounces has been reached.
+The recursion stops when a maximal number of bounces has been reached.
 %ASK: euh, c'est L ou L_r%
@@
 \begin{equation*}
-f_r(x, \Theta \leftrightarrow \Phi) L(x \leftarrow \Phi) cos(N_x, \Phi).
+f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi).
 \end{equation*}
 
-It is a product of three functions that can be efficiently sampled individually but that are hard to sample when they are combined.
-The most common strategy used to sample $\Omega_x$ is to use the BRDF (combined with the cosine) because we know it: it is given as an input of the algorithm, for exemple with textures.
+It is a product of two functions that can be efficiently sampled individually but that are hard to sample when they are combined.
+The most common strategy used to sample $\Omega_x$ is to use the function $f_s$ because we know it: it is given as an input of the algorithm, for exemple with textures that describe materials.
 
@@
 Let $n$ be a node of the skeleton and $V_n$ the set of visible nodes from $n$ along the shortest path toward $n_L$. The importance point $imp_n$ associated to $n$ is the barycenter of $V_n$.
+
+\begin{center}
+\includegraphics[scale=0.5]{images/importance_points.png}
+\end{center}
 
@@
 
+\begin{center}
+\includegraphics[scale=0.17]{images/skel_based_rt.png}
+\end{center}
+
+This image is a zoom that allows us to see more precisely the the reduction of noise for our method.
+
+\section{Conclusion}
+
 We presented an application of our skeletonization algorithm for global illumination.

### 2012-09-14 application_to_path_tracing.tex
Some minor modifications

@@
 
+\begin{figure}[htp]
 \begin{center}
-\includegraphics[scale=0.17]{images/four_scenes.png}
+\includegraphics[scale=0.17]{images/four_scenes_for_dgci.png}
+\caption{Four images comparing the classic path tracing (image $a$ and $c$) and our method (image $b$ and $d$).
+Note the noise reduction of our method in the left wall of images $a$ and $b$, and on the floor on image $c$ and $d$.}
 \end{center}
+\end{figure}
 
@@
 The MSE values are lower for the top pictures, meaning that we converge faster to the reference image than standard path tracing does.
-
-\begin{center}
-\includegraphics[scale=0.17]{images/skel_based_rt.png}
-\end{center}
 

### 2012-09-14 skeleton_and_pathtracing.tex
Renommage fichier tex application
Small change

--- 2012-09-10 bidir_pathtrace_skeleton.tex
@@
 
-%\usepackage{graphicx} %Pouvoir insérer des images
+%\usepackage{graphicx} %Pouvoir inserer des images
 \usepackage{epsfig}
-\usepackage{amsmath} %Permet d'utiliser une jolie panoplie de symboles mathématiques
+\usepackage{amsmath} %Permet d'utiliser une jolie panoplie de symboles mathematiques
 \usepackage{amssymb}
@@
 
-\def\axisloc#1{\mathcal{LOC}_{#1}}
-\def\axislocbis#1#2{\mathcal{LOC}_{#1}^{>#2}}
-\def\axislc#1{\mathcal{LC}_{#1}}
 \newcommand{\centerness}{Decenter}
@@
 
-\def\mydist1{D_1}
-\def\mydd1{\Omega_1}
-\def\mydecent{\mathcal{DC}_1}
-\def\dist1{d_1}
+\def\mydist1{D}
+\def\mydd1{\Omega}
+\def\mydecent{\mathcal{DC}}
+\def\dist1{d}
 \def\Dist1#1{\mydist1(#1)}
@@
 
-\def\birth#1#2{Birth_{#2}(#1)}
-\def\death#1#2{Death_{#2}(#1)}
-\def\lifespan#1#2{Lifespan_{#2}(#1)}
+\def\birth#1#2{\mbox{\it{}Birth\/}_{#2}(#1)}
+\def\death#1#2{\mbox{\it{}Death\/}_{#2}(#1)}
+\def\lifespan#1#2{\mbox{\it{}Lifespan\/}_{#2}(#1)}
 
+\def\V2C{\Psi}
 \def\Nset{\mathbb{N}}
@@
 \def\cellstrict#1{\hat{#1}^*}
-\def\cont#1{\check{#1}}
-\def\contstrict#1{\check{#1}^*}
 \def\closure#1{#1^-}
 \def\princ#1{#1^+}
-\def\detach#1#2{#2 \oslash #1}
-\def\attach#1#2{\mbox{\it{}Attach\/}(\cell{#1},#2)}
 \def\mydim#1{\dim(#1)}
@@
 
-\def\figpath#1{../../Images/EPS/#1}
+\def\figpath{./images}
+
+\def\figpath{images}
+\def\myincludegraphics#1#2{\includegraphics[width=#1]{\figpath/#2}}
 
@@
 
-\title{A 3d curvilinear skeletonization algorithm with application to bidirectional path tracing}
-\author{John Chaussard\inst{1} \and Laurent No\"{e}l\inst{2} \and Venceslas Biri\inst{2} \and Michel Couprie\inst{2}}
+\title{A 3d curvilinear skeletonization algorithm with application to path tracing}
+\author{John Chaussard\inst{1} \and Laurent No\"{e}l\inst{2} \and Venceslas Biri\inst{2} \and Michel Couprie\inst{2}\thanks{This work has been partially supported by the \quotes{ANR-2010-BLAN-0205 KIDICO} project}}
 \institute{Universit\'{e} Paris 13, Sorbonne Paris Cit\'{e}, LAGA, CNRS(UMR 7539), \\ F-93430, Villetaneuse, France \\ \email{chaussard@math.univ-paris13.fr}
 \and
-Universit\'{e} Paris Est, LABINFO-IGM, A3SI-ESIEE \\ 2, boulevard Blaise Pascal, Cit\'{e} DESCARTES \\ BP 99  93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}
+Universit\'{e} Paris Est, LIGM, A3SI-ESIEE \\ 2, boulevard Blaise Pascal, Cit\'{e} DESCARTES \\ BP 99  93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}
 
@@
 \begin{abstract}
-What we propose is really great...
+We present a novel 3d curvilinear skeletonization algorithm which produces filtered skeletons without needing any user input, thanks to a new parallel algorithm based on the cubical complex framework.
+We then propose a new path tracing algorithm which uses such curvilinear skeleton in order to produce less noisy images in less time than the classical approach.
 \end{abstract}
@@
 \label{sec::intro}
-Blah blah blah...
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{cc}
+	\myincludegraphics{0.60\textwidth}{dgci_ray_corridor.png} &
+	\myincludegraphics{0.40\textwidth}{dgci_compare_door.png}
+\end{tabular}
+\caption{Images $a$ and $b$ shows some rays of the path tracing guided by the material of the objects (image $a$) or our skeleton based method (image $b$).
+Images $c$ and $d$ point out the noise reduction between the classic path tracing (image $c$) and our method (image $d$)}
+\end{center}
+\end{figure}
+Path tracing algorithms are able to render photorealistic images from a scene description by simulating light propagation.
+The photorealistic aspect is achieved by simulating diffusion of light: elements of the scene which are illuminated by a light source can cast part of the light they receive.
+Consequently, parts of the scene which are not directly illuminated by any light source actually receive light, thus giving a realistic effect of light diffusion to the resulting image.
+
+A path tracing algorithm sends rays (of light) from the camera (the viewpoint of the scene), and let them bounce around the scene until they encounter a directly illuminated zone.
+When a ray hits a point of an object (like a wall), it gathers information on the object's material.
+The ray is then reemitted in a random direction.
+Once a ray reaches a light source, the information it gathered through its journey in the scene is averaged in order to decide of the color of the first impact point.
+This color is then allocated to one of the pixels of the image, and the process is repeated for each pixel of the output image.
+To avoid \quotes{unrealistic} results due to an unusual reemission direction of a ray, the algorithm actually sends, for each pixel of the image, many rays: the more rays, the longer the algorithm will take to produce an image, but the more realistic the result will look.
+A path tracing can be seen as a random walk process through the scene.
+
+In order to avoid infinite bouncing of rays in the scene, a limit is set for the number of bounces a ray can do.
+If a ray does not reach any light source after a given number of bounces, it is eliminated from the scene.
+A lost ray is as a waste of computation time, and minimizing this loss would speed up the algorithm.
+
+Our proposition is to reduce both the time taken by a path tracing algorithms and the noise in the output images by guiding rays towards the scene's light sources, thus reducing the number of rays \quotes{lost} during the computation.
+To do this, we perform some precomputation on the void space of the scene (the medium where the rays of light propagate) in order to have, later in the algorithm, a clue on towards which direction a ray should bounce in order to meet a light source.
+Such precomputation can take time and become more expensive than the original algorithm.
+
+To keep the precomputation fast, we perform it only on a subset of points of the voids of the scene.
+The precomputation we want to perform relies on paths and visibility tests, therefore want this set of points to be representative of both the topology and the geometry of the original scene.
+For example, it should possess branches propagating in the elongated parts of the scene.
+A filtered curvilinear skeleton of (the voids of) the scene meets all the required properties for our subset of points: it is a thin subset of points of the scene, with the same topology and retaining the main geometrical information of the scene.
+Note that in this application, the filtering of the skeleton (elimination of spurious branches) need to be both efficient and completely automatic, for one cannot ask a user to set parameters for each scene.
+
+The purpose of this paper is twofold: we first introduce a new curvilinear skeletonization algorithm (Sec.~\ref{sec::skeletonization}) in the cubical complex framework (Sec.~\ref{sec::binary_to_cubic}), producing filtered skeletons without needing any user input (Sec.~\ref{sec::aspect_no_param}).
+Then, we expose a new method, based on such skeletons, to enhance the results and the performance of a path tracing algorithm (see Sec.~\ref{sec::appli_pathtrace}).
 
 \section{The cubical complex framework}
-
-\subsection{Basic definitions}
+\label{sec::binary_to_cubic}
 
@@
 We denote by $\allfaces{n}$ the set composed of all faces in $\Z^n$. Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.
-
 Let $f \in \allfaces{n}$.
 We set $\cell{f} = \{g \in \allfaces{n}\vert g \subseteq f\}$, and $\cellstrict{f} = \cell{f} \setminus \{f\}$.
-Any element of $\cell{f}$ is \myem{a face of $f$}, and any element of $\cellstrict{f}$ is \myem{a proper face of $f$}.
-We set $\cont{f} = \{g \in \allfaces{n} \vert f \subseteq g\}$, and $\contstrict{f} = \cont{f} \setminus \{f\}$.
-It is plain that $g \in \cell{f}$ iff $f \in \cont{g}$.
-
-A set $X$ of faces in $\allfaces{n}$ is a \myem{cell}, or \myem{$m$-cell}, if there exists an $m$-face $f \in X$ such that $X=\cell{f}$.
+Any element of $\cell{f}$ (resp.
+$\cellstrict{f}$) is \myem{a face of $f$} (resp.
+\myem{a proper face of $f$}.
 The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$.
-The set $\Compl{X}$ is $\allfaces{n} \setminus X$.
 
@@
 
-Informally, in 3d, a set of faces $X$ is a complex if each side (square) of a cube of $X$ also belong to $X$, each edge of a square of $X$ also belong to $X$, and each vertex of an edge of $X$ also belong to $X$.
+%Informally, in 3d, a set of faces $X$ is a complex if each side (square) of a cube of $X$ also belong to $X$, each edge of a square of $X$ also belong to $X$, and each vertex of an edge of $X$ also belong to $X$.
 
 A face $f \in X$ is \myem{a facet of $X$} if $f$ is not a proper face of any face of $X$.
-We denote by $\princ{X}$ the set composed of all facets of $X$ .
-A complex $X$ is \myem{pure} if all its facets have the same dimension.
 The \myem{dimension of $X$} is $\mydim{X} = \max\{\mydim{f} \suchthat f \in X\}$.
@@
 
-\subsection{From binary images to cubical complex}
-\label{sec::binary_to_cubic}
+Traditionally, a binary image is a finite subset of $\Zset^n$ (called voxel image when $n=3$).
+To transpose such image to the cubical complex framework, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$.
+Let $x=(x_1,...,x_n) \in S$, we define the n-face $\V2C(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$.
+We can extend the map $\V2C$ to sets: $\V2C(S)=\{\V2C(x) | x \in S\}$.
+Given a set $S \subset \Zset^n$, we associate to it the cubical complex $\closure{\V2C(S)}$.
 
-Traditionally, a binary image is a finite subset of $\Zset^n$ (called voxel image when $n=3$). We give now a simple way to transpose such image to the cubical complex framework.
-
-To do so, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$.
-More precisely, let $x=(x_1,...,x_n) \in S$, we define the n-face $\Phi(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$.
-We can extend the map $\Phi$ to sets: $\Phi(S)=\{\Phi(x) | x \in S\}$.
-Given a set $S \subset \Zset^n$, we associate to it the cubical complex $\closure{\Phi(S)}$.
-
-\subsection{Thinning: the collapse operation}
-The collapse operation is the basic operation for performing homotopic thinning of a complex, and consists of removing two distinct faces $(f,g)$ from a complex $X$ under the condition that they are free:
+The collapse operation is the basic operation for performing homotopic thinning of a complex, and consists of removing two distinct faces $(f,g)$ from a complex $X$ under the condition that they form a free pair:
 
 \begin{definition}
-Let $\complex{n}{X}$, and let $f,g$ be two faces of $X$. The face $g$ is \myem{free for $X$}, and the pair $(f,g)$ is \myem{a free pair for $X$} if $(\contstrict{g} \inter X) = \{f\}$. 
+Let $\complex{n}{X}$, and let $f,g$ be two faces of $X$.
+The face $g$ is \myem{free for $X$}, and the pair $(f,g)$ is \myem{a free pair for $X$} if $f$ is the only face of $X$ which strictly contains $g$.
 \end{definition}
@@
 
+Recent works in the cubical complex framework brought new parallel thinning algorithms in the voxel framework (\cite{BertrandCouprie06}, \cite{Ber07}).
+
 \section{A parallel directional thinning based on cubical complex}
+\label{sec::skeletonization}
 
-\subsection{Removing free pairs in parallel}
-In the cubical complex framework, parallel removal of simple pairs can be easily achieved when following simple rules that we give now.
+In the cubical complex framework, parallel removal of free pairs can be easily achieved when following simple rules that we give now.
 First, we need to define the \myem{direction} and the \myem{orientation} of a free face.
+Let $(f,g)$ be a free pair for $\complex{n}{X}$ : we have $\mydim{g} = \mydim{f}-1$, and it can be seen that $g = f \inter f'$, where $f'$ is the translate of $f$ by one of the $2n$ vectors of $\Zset^n$ which have all their coordinates equal to 0 except one, which is either equal to +1 or -1.
+Let $v$ be this vector, and $c$ its non-null coordinate.
+We define $\type{f,g}$, called the \myem{direction} of the free pair $(f,g)$, as the index of $c$ in $v$.
+The \myem{orientation} of the free pair $(f,g)$ is defined as $\orient{f,g}=1$ if $c=1$, and $\orient{f,g}=0$ else.
 
-Let  $f \in \allfaces{n}$, the \myem{center of $f$} is the center of mass of the points in $f$, that is, $c_f= \frac{1}{\Card{f}}\sum_{a \in f}{a}$.
-The center of $f$ is an element of $[\Zhalf]^n$, where $\Zhalf$ denotes the set of half integers.
-Let $\complex{n}{X}$, let $(f,g)$ be a free pair for $X$, and let $c_f$ and $c_g$ be the respective centers of the faces $f$ and $g$. We denote by $V(f,g)$ the vector $(c_f -c_g)$ of $[\Zhalf]^n$.
+%On Fig.~\ref{fig:parallel_collapse}, the free pair $(a,c)$ and the free pair $(d,e)$ have different directions; the free pairs $(a,b)$ and $(d,e)$ have the same direction, but different orientations.
 
-We define a surjective function $\type{}: \allfaces{n} \times \allfaces{n} \rightarrow \dinterval{0}{n-1}$ such that, for all free pairs $(f,g)$ and $(i,j)$ for $X$, $\type{f,g} = \type{i,j}$ if and only if $V(f,g)$ and $V(i,j)$ are collinear.
-The number $\type{f,g}$ is called the \myem{direction} of the free pair $(f,g)$.
-Let $(f,g)$ be a free pair, the vector $V(f,g)$ has only one non-null coordinate: the pair $(f,g)$ has a \myem{positive orientation}, and we write $\orient{f,g}=1$, if the non-null coordinate of $V(f,g)$ is positive; otherwise $(f,g)$ has a \myem{negative orientation}, and we write $\orient{f,g}=0$.
-%On Fig.~\ref{fig:parallel_collapse}, the free pair $(a,c)$ and the free pair $(d,e)$ have different directions; the free pairs $(a,b)$ and $(d,e)$ have the same direction, but opposite orientations.
-
-Now, we give a property of collapse which brings a necessary and sufficient condition for removing two free pairs of faces in parallel from a complex, while preserving topology.
+Now, we give a property of collapse (previously proven in \cite{ChauCou2009}) which brings a necessary and sufficient condition for removing two free pairs of faces in parallel from a complex, while preserving topology.
 
@@
 
-\begin{proof}
-If $f=k$, then it is plain that $(k,\ell)$ is not a free pair for $Y=X \setminus \{f,g\}$ as $k=f \notin Y$.
+%\begin{proof}
+%If $f=k$, then it is plain that $(k,\ell)$ is not a free pair for $Y=X \setminus \{f,g\}$ as $k=f \notin Y$.
 Also, $(f,g)$ is not free for $X \setminus \{k,\ell\}$.
@@
 $\finproof$
-\end{proof}
+%\end{proof}
 
@@
 
-\subsection{A directional parallel thinning algorithm}
-
-Intuitively, we want the algorithm to remove free faces of a complex \quotes{layer by layer} and avoid having unequal thinning of the input complex.
+Intuitively, we want our thinning algorithm to remove free faces of a complex \quotes{layer by layer} and to avoid having unequal thinning of the input complex.
 Therefore, we want each execution of the algorithm to remove free faces located on the border of the input complex.
@@
 On a single execution of the main loop of Alg.~\ref{algo::pardircollapse_v0}, only faces located on the border of the complex are removed (l.~\ref{pc:07}).
-Thanks to corollary \ref{cor:parallel_removal_v2}, we remove faces with same direction and orientation in parallel (l.~\ref{pc:08}).
+Thanks to corollary \ref{cor:parallel_removal_v2}, we can remove faces with same direction and orientation in parallel (l.~\ref{pc:08}), while guaranteeing topology preservation.
+Figure \ref{fig::example_algo} depicts the first steps of the algorithm.
 
-\begin{algorithm}[h!]
+\begin{algorithm}[ht!]
 \label{algo::pardircollapse_v0}
@@
 
-Different definitions of the orientation and direction can be given, corresponding to different order of removal of free faces in the complex and leading to different results.
-Algorithm~\ref{algo::pardircollapse_v0} can be implemented to run in linear time complexity (proportionally to the number of faces of the complex).
+\begin{figure}[htp]
+\begin{center}
+\begin{tabular}{ccccc}
+	\myincludegraphics{0.19\textwidth}{invader.pdf} &
+	\myincludegraphics{0.19\textwidth}{invader_step1_2.pdf} &
+	\myincludegraphics{0.19\textwidth}{invader_step2_2.pdf} &
+	\myincludegraphics{0.19\textwidth}{invader_step3.pdf} &
+	\myincludegraphics{0.19\textwidth}{invader_step4.pdf} \\
+\end{tabular}
+\caption[Four first iterations of Alg.~\ref{algo::pardircollapse_v0}]{\label{fig::example_algo} \textbf{Four first iterations of Alg.~\ref{algo::pardircollapse_v0}} running on the left-most shape.}
+\end{center}
+\end{figure}
+
+Different definitions of orientation and direction can be given, each corresponding to a different order of free faces removal in the complex and leading to different results.
+Algorithm~\ref{algo::pardircollapse_v0} can be implemented to run in linear time complexity (proportionally to the number of faces in the complex).
 Indeed, checking if a face is free or not may be easily done in constant time.
-Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the elements of $(\cellstrict{f} \union \contstrict{g})$ in order to find new free faces.
+Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the faces contained in $f$ in order to find new free faces.
 
@@
 \label{sec::aspect_no_param}
-Algorithm \ref{algo::pardircollapse_v0} does not necessarily preserves "geometrical information" from the original object in the resulting skeleton.
-For example, if the input was a filled human shape (one connected component, no tunnel nor cavity) and that $W=\emptyset$ and $l=+\inf$, then the result of Alg.~\ref{algo::pardircollapse_v0} would be a simple vertex.
-However, when one wants to preserve "geometrical information" from the original shape in the skeleton, then one expects to obtain a "stickman" from a human shape.
-It is however important to not retain "too much information" during thinning in order to avoid noisy branches in the skeleton.
+As previously said, our goal is to use the skeleton of the voids of a scene in order to guide light rays from the camera of the scene towards the light.
+Moreover, this skeleton needs to capture the main geometrical features of the original scene.
+For example, if the input is a corridor (the void of the corridor), then the skeleton should be a line following the main direction of the corridor.
 
-In the following, we will see a new method in the cubical complex, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory visual properties.
-This will be achieved by adding, in the set $W$ of Alg.~\ref{algo::pardircollapse_v0}, some faces from the original object during the thinning process.
+Generally, two strategies are possible to achieve this goal: find, during the skeletonization process, points whose neighbourhood configuration seems interesting and keep them in the result \cite{CouCoeurZrour07} \cite{HR08} \cite{CCT09ext}), or choose, before skeletonization, interesting points of the object which should remain untouched, based on a function on these points and a filtering parameter \cite{BertrandCouprie06} \cite{Pal07}.
+
+Algorithm ~\ref{algo::pardircollapse_v0} does not necessarily preserve geometrical features of the input object in the resulting skeleton, which end up being a single vertex.
+
+In the following, we introduce a new method in the cubical complex framework, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory geometrical properties.
+Our method is based on the two previously listed strategies: its finds, during thinning, elements whose neighbourhood configuration is interesting, and uses a function on these elements to decide to preserve them in the result.
 
 \subsection{The lifespan of a face}
-In the following, we define more notions in the cubical complex. The first one we present is the \myem{death date} of a face.
+In the following, we define additional functions in the cubical complex, related to the thinning process (Alg.~\ref{algo::pardircollapse_v0}), which are essential for the filtering step of the skeletonization.
+The first one we present is the \myem{death date} of a face.
 
@@
 
-The death date of a face indicates how many layers of free faces should be removed from a complex $X$, using Alg.~\ref{algo::pardircollapse_v0}, before removing completely the face from $X$.
+Intuitively, the death date of a face indicates how many layers of free faces should be removed from a complex $X$, using Alg.~\ref{algo::pardircollapse_v0}, before removing completely the face from $X$.
 We now define the \myem{birth date} of a face:
@@
 \begin{definition}
-Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or either $f \notin ParDirCollapse(X, \emptyset, b)$.
+Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or $f \notin ParDirCollapse(X, \emptyset, b)$.
 \end{definition}
 
-The birth date indicates how many layers of free faces must be removed from $X$ with Alg.\ref{algo::pardircollapse_v0} before transforming $f$ into a facet of $X$ (we consider a face "lives" when it is a facet).
-Finally, we can define the \myem{lifespan} of a face :
+The birth date indicates how many layers of free faces must be removed from $X$ with Alg.\ref{algo::pardircollapse_v0} before transforming $f$ into a facet of $X$ (we consider that a face \quotes{lives} when it is a facet).
+Finally, we define the \myem{lifespan} of a face :
 
@@
 
-These three values are dependant on the order of direction and orientation chosen for Alg.~\ref{algo::pardircollapse_v0}.
+These three values depend on the order of direction and orientation chosen for Alg.~\ref{algo::pardircollapse_v0}.
 
-The lifespan of a face $f$ of $X$ indicates how many "rounds" this face "survives" as a facet in $X$, when removing free faces with Alg.~\ref{algo::pardircollapse_v0}. 
-The lifespan is a good indicator of how important a face can be in an object.
-Typically, higher the lifespan is, and more representative of an object's visual feature the face is.
-The lifespan, sometimes called \myem{saliency}, was used in \cite{Liu2009} (with the name "medial persistence") in order to propose a thinning algorithm in cubical complexes based on two parameters.
+The lifespan of a face $f$ of $X$ indicates how many \quotes{rounds} this face \quotes{survives} as a facet in $X$, when removing free pairs with Alg.~\ref{algo::pardircollapse_v0}, and is a good indicator of how important a face can be in an object.
+Typically, the higher the lifespan is, and the more representative of an object's geometrical feature the face is.
+The lifespan, sometimes called \myem{saliency}, was used in \cite{Liu2009} (with the name \quotes{medial persistence}) in order to propose a thinning algorithm in cubical complexes based on two parameters.
 
@@
 In addition to the lifespan of a face, the proposed homotopic thinning method uses information on distance between faces in order to decide if a face should be kept safe from deletion.
-We define hereafter the various notions needed for this, based on distance in the voxel framework.
+We define hereafter various notions based on distances in the voxel framework.
 
@@
 The length of $\mathcal{C}$ is $k$.
-Remark that $d_1$ is indeed the $6$-distance in the 3d voxel framework.
 
-We set $\dist1(x,y)$ as the length of the shortest $1$-path from $x$ to $y$.
-Let $S \subset \Zset^n$, we set, for all $x \in Zset^n$, the map $\Dist1{S}: \Zset^n \rightarrow \Nset$ such that $\Dist1{S}(x) = \displaystyle \min_{y \in \Compl{S}} \dist1(x,y)$.
+We set $\dist1(x,y)$ as the length of a shortest $1$-path from $x$ to $y$; remark that $\dist1$ is indeed the $6$-distance in the 3d voxel framework.
+Let $S \subset \Zset^n$, we set $\Compl{S} = \Zset^n \setminus S$, and for all $x \in \Zset^n$, the map $\Dist1{S}: \Zset^n \rightarrow \Nset$ is such that $\Dist1{S}(x) = \displaystyle \min_{y \in \Compl{S}} \dist1(x,y)$.
 
@@
 If $x \in \Compl{S}$, we set $\DD1{S}(x) = 0$.
-The map $\DD1{S}$ is known as the opening function of $S$ (based on the $1$-distance): it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.
+The map $\DD1{S}$ is known as the opening function of $S$ based on the $1$-distance (also called the granulometry function) \cite{Matheron67}: it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.
 
 Given $S \subset \Zset^n$, the value of $\DD1{S}(x)$ of every $x \in S$ can be naively computed by performing successive morphological dilations of values of the map $\Dist1{S}$.
-%TODO Parler du fait que l'on a une version dans la thèse, et une version plus amélioré dans thèse et public avenir
+A linear algorithm for computing the map $\DD1{S}$ (with regards to the size of the input image) was proposed in \cite{Chaussard_phd}, and will be explored further in details in a future paper.
 
-Finally, we define the \myem{decenterness map} :
+Finally, we define the \myem{decenterness map}:
 
@@
 
-In order to extend all these previous maps defined in $\Zset^n$ to the cubical complex framework, let us introduce the map $\Phi^{-1}$, inverse of the bijective map $\Phi : \subfaces{n}{n} \rightarrow \Zset^{n}$ defined in Sec.~\ref{sec::binary_to_cubic}.
+An exemple of these maps is shown on Fig.~\ref{fig::globalopening}.
+
+In order to extend all these previous maps defined in $\Zset^n$ to the cubical complex framework, we use the map $\V2C^{-1}$, inverse of the bijective map $\V2C : \Zset^{n} \rightarrow \subfaces{n}{n}$ defined in Sec.~\ref{sec::binary_to_cubic}.
 It is used to project any $n$-face of $\allfaces{n}$ into $\Zset^n$.
-We indifferently use $\Phi^{-1}$ as a map from $\subfaces{n}{n}$ to $\Zset^n$, and as a map from $\mathcal{P}(\subfaces{n}{n})$ to $\mathcal{P}(\Zset^n)$.
+This map induces a map from $\mathcal{P}(\subfaces{n}{n})$ to $\mathcal{P}(\Zset^n)$, that we also denote by $\V2C^{-1}$.
 
-Given $Y \subset \allfaces{n}$, we define the map $\cplDist1{Y}: \allfaces{n} \rightarrow \Nset$ as follows: for all $f \in \allfaces{n}$,
+Given $Y \subset \allfaces{n}$, we set $S = \V2C^{-1}(Y \inter \subfaces{n}{n})$. We define the map $\cplDist1{Y}: \allfaces{n} \rightarrow \Nset$ as follows: for all $f \in \allfaces{n}$,
 
@@
 \begin{array}{ll}
-\Dist1{\Phi^{-1}(Y \inter \subfaces{n}{n})}(Phi^{-1}(f)) & \text{if $f$ is an $n$-face}\\
-\displaystyle \max_{g \in \contstrict{f} \inter \subfaces{n}{n}} \cplDist1{Y}(g) & \text{else}
+\Dist1{S}(\V2C^{-1}(f)) & \text{if $f$ is an $n$-face}\\
+\displaystyle \max_{f \in (\cellstrict{g} \inter \subfaces{n}{n})} \cplDist1{Y}(g) & \text{otherwise}
 \end{array}
@@
 \end{math}
+\bigskip
 
@@
 
+\begin{figure}[htp]
+\begin{center}
+\begin{tabular}{cccc}
+	\myincludegraphics{0.22\textwidth}{fish.pdf} &
+	\myincludegraphics{0.22\textwidth}{fishd.pdf} &
+	\myincludegraphics{0.22\textwidth}{fishdd.pdf} &
+	\myincludegraphics{0.22\textwidth}{fishcenter.pdf} \\
+\end{tabular}
+\caption[Example of opening and decenterness map]{\label{fig::globalopening} \textbf{Example of opening and decenterness map - } From left to right: a shape $S \subset \Zset^2$ (in gray), $\Dist1{S}$, $\DD1{S}$ and $\decent{S}$ (low values have dark colour).}
+\end{center}
+\end{figure}
+
 \subsection{Parameter-free thinning based on the lifespan, opening function and decenterness}
 As previously said, we add edges to the set $W$ of Alg.~\ref{algo::pardircollapse_v0} in order to retain, in the resulting curvilinear skeleton, important edges from the original object.
-Given a cubical complex $X$, if an edge of $X$ has a high decenterness value for $X$, then it is probably located too close to the border of $X$ and does not represent an interesting visual feature to preserve.
+Given a cubical complex $X$, if an edge of $X$ has a high decenterness value for $X$, then it is probably located too close to the border of $X$ and does not represent an interesting geometrical feature to preserve.
 %TODO ref vers image de decenterness
-On the other hand, if an edge has a high lifespan for $X$, then it means it was not removed quickly, after becoming a facet, by the thinning algorithm and might represent some precious visual information on the original object.
+On the other hand, if an edge has a high lifespan for $X$, then it means it was not removed quickly, after becoming a facet, by the thinning algorithm and might represent some precious geometrical information on the original object.
 An idea would be to keep, during thinning, all edges whose lifespan is superior to the decenterness value.
@@
 We can identify surfacic areas of a complex as zones where squares have a high lifespan.
-Therefore, in order to avoid spurious branches in surfacic areas, we need to make it harder for edges to appear in these zones.
+Therefore, in order to avoid spurious branches in surfacic areas, we need to make it harder for edges to be preserved in these zones.
 It can be achieved by deciding that an edge will be kept safe from deletion by the thinning algorithm if its lifespan is superior to the decenterness value plus the lifespan of squares \quotes{around} this edge.
@@
 
-\input{pathtracint_application_skeleton}
+\subsection{Results}
+
+Algorithm \ref{algo::3d_curvskel} allows to obtain a filtered curvilinear skeleton from a three dimensional complex.
+The results presented in Fig.~\ref{fig::result_skel} show that the skeletons contain the main geometrical information from the input shapes, and do not contain spurious branches.
+
+\begin{figure}[htp]
+\begin{center}
+\begin{tabular}{cc|cc}
+	\myincludegraphics{0.18\textwidth}{hand2.jpg} &
+	\myincludegraphics{0.18\textwidth}{hand_curvskel.jpg} &
+	\myincludegraphics{0.24\textwidth}{neptune.jpg} &
+	\myincludegraphics{0.24\textwidth}{neptune_curvskel.jpg} \\
+\end{tabular}
+\caption[Results of algorithm \ref{algo::3d_curvskel}]{\label{fig::result_skel} \textbf{Results of algorithm \ref{algo::3d_curvskel}} for two shapes: a hand (left) and a statue (right).
+In each pair, the rightmost image represents the skeleton.}
+\end{center}
+\end{figure}
+
+\input{application_to_path_tracing}
 

### 2012-09-14 application_to_path_tracing.tex
Small change
presque 12 pages
Essai

@@
 with $\alpha$ the angle between $\overrightarrow{ximp_n}$ and $\Phi$, $s$ being a parameter called skeleton strength. The higher $s$ is, the more we sample close $\overrightarrow{ximp_n}$.
-
-\subsection{Results and discussion}
-
-We present some results of our algorithm applied to four scenes.
-The first row shows our result and the second row shows a standard path tracing with the same input parameters.
-On each picture a MSE (mean square error) value with a reference image is shown.
-
-\begin{figure}[htp]
+\begin{figure}[thp]
 \begin{center}
 \includegraphics[scale=0.17]{images/four_scenes_for_dgci.png}
-\caption{Four images comparing the classic path tracing (image $a$ and $c$) and our method (image $b$ and $d$).
+\caption{\label{fig:results}Four images comparing the classic path tracing (image $a$ and $c$) and our method (image $b$ and $d$).
 Note the noise reduction of our method in the left wall of images $a$ and $b$, and on the floor on image $c$ and $d$.}
@@
 
-We observe that images produced by our algorithm have less noise than the ones produced by the regular path tracing (for a same number of rays shot per pixel).
-The MSE values are lower for the top pictures, meaning that we converge faster to the reference image than standard path tracing does.
-
-This image is a zoom that allows us to see more precisely the the reduction of noise for our method.
+\subsection{Results and discussion}
 

### 2012-09-14 skeleton_and_pathtracing.tex

@@
 \and
-Universit\'{e} Paris Est, LIGM, A3SI-ESIEE \\ 2, boulevard Blaise Pascal, Cit\'{e} DESCARTES \\ BP 99  93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}
+Universit\'{e} Paris Est, LIGM, A3SI-ESIEE \\ 2, boulevard Blaise Pascal, 93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}
 
@@
 \label{sec::intro}
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{cc}
-	\myincludegraphics{0.60\textwidth}{dgci_ray_corridor.png} &
-	\myincludegraphics{0.40\textwidth}{dgci_compare_door.png}
-\end{tabular}
-\caption{Images $a$ and $b$ shows some rays of the path tracing guided by the material of the objects (image $a$) or our skeleton based method (image $b$).
-Images $c$ and $d$ point out the noise reduction between the classic path tracing (image $c$) and our method (image $d$)}
-\end{center}
-\end{figure}
+
 Path tracing algorithms are able to render photorealistic images from a scene description by simulating light propagation.
@@
 Note that in this application, the filtering of the skeleton (elimination of spurious branches) need to be both efficient and completely automatic, for one cannot ask a user to set parameters for each scene.
+
+\begin{figure}[tb]
+\begin{center}
+\begin{tabular}{cc}
+	\myincludegraphics{0.60\textwidth}{dgci_ray_corridor.png} &
+	\myincludegraphics{0.40\textwidth}{dgci_compare_door.png}
+\end{tabular}
+\caption{Images $a$ and $b$ shows some rays of the path tracing guided by the material of the objects (image $a$) or our skeleton based method (image $b$).
+Images $c$ and $d$ point out the noise reduction between the classic path tracing (image $c$) and our method (image $d$)}
+\end{center}
+\end{figure}
 
@@
 
-\begin{algorithm}[ht!]
+\begin{algorithm}[tb]
 \label{algo::pardircollapse_v0}
@@
 
-\begin{figure}[htp]
+\begin{figure}[tb]
 \begin{center}
@@
 Algorithm~\ref{algo::pardircollapse_v0} can be implemented to run in linear time complexity (proportionally to the number of faces in the complex).
-Indeed, checking if a face is free or not may be easily done in constant time.
-Moreover, when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the faces contained in $f$ in order to find new free faces.
+Indeed, checking if a face is free or not may be easily done in constant time and when a free pair $(f,g)$ is removed from the input complex, it is sufficient to scan the faces contained in $f$ in order to find new free faces.
 
-\section{Aspect preservation during thinning: a parameter-free method}
+\section{Aspect preservation during thinning}
 \label{sec::aspect_no_param}
@@
 
-\subsection{Parameter-free thinning based on the lifespan, opening function and decenterness}
+\subsection{Parameter-free filtered thinning}
 As previously said, we add edges to the set $W$ of Alg.~\ref{algo::pardircollapse_v0} in order to retain, in the resulting curvilinear skeleton, important edges from the original object.
 Given a cubical complex $X$, if an edge of $X$ has a high decenterness value for $X$, then it is probably located too close to the border of $X$ and does not represent an interesting geometrical feature to preserve.
-%TODO ref vers image de decenterness
 On the other hand, if an edge has a high lifespan for $X$, then it means it was not removed quickly, after becoming a facet, by the thinning algorithm and might represent some precious geometrical information on the original object.
@@
 
-\begin{algorithm}[ht]
+\begin{algorithm}[tb]
 \label{algo::3d_curvskel}
@@
 Algorithm \ref{algo::3d_curvskel} allows to obtain a filtered curvilinear skeleton from a three dimensional complex.
-The results presented in Fig.~\ref{fig::result_skel} show that the skeletons contain the main geometrical information from the input shapes, and do not contain spurious branches.
+The results presented in Fig.~\ref{fig::result_skel} show that the skeletons contain the main geometrical information from the input shapes, and no spurious branch.
 
-\begin{figure}[htp]
+\begin{figure}[tb]
 \begin{center}

### 2012-09-14 application_to_path_tracing.tex
Inserting new biblio

@@
 \begin{align*}
-L(x \rightarrow \Theta) &= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) d\omega_\Phi \\
-&= L_e(x \rightarrow \Theta) + L_r(x \rightarrow \Theta)
+L(x \rightarrow \Theta) &= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) d\omega_\Phi
 \end{align*}
@@
 This integrand expresses an intuitive idea: the reflected radiance from $x$ towards the camera is the result of computing all incoming luminosity on $x$ scaled by a factor which depends on the material on $x$ and the angle of the ray going from the camera to $x$.
-The most common method used to estimate the integral is the Monte-Carlo integration which provides the following:
+The most common method used to estimate the integral, denoted by $L_r(x \rightarrow \Theta)$, is the Monte-Carlo integration which provides the following:
 
@@
 We sample the hemisphere with a power-cosine pdf centered on $\overrightarrow{ximp_n}$:
-
 \begin{equation*}
@@
 \end{equation*}
+with $\alpha$ the angle between $\overrightarrow{ximp_n}$ and $\Phi$, $s$ being a parameter called skeleton strength. The higher $s$ is, the more we sample close $\overrightarrow{ximp_n}$.
 
-with $\alpha$ the angle between $\overrightarrow{ximp_n}$ and $\Phi$, $s$ being a parameter called skeleton strength. The higher $s$ is, the more we sample close $\overrightarrow{ximp_n}$.
-\begin{figure}[thp]
+\subsection{Results and discussion}
+
+Some results are presented in figure~\ref{fig:results} where we can see the decrease in noise compare to the regular path tracing.
+We computed the MSE value for different scenes that are presented in table~\ref{tab:timing}
+
+\begin{figure}[tbhp]
 \begin{center}
@@
 
-\subsection{Results and discussion}
+\begin{table}[t!]
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+Scene & fig.\ref{fig:redcorridor} & fig.\ref{fig:examples} far left & fig.\ref{fig:examples} far right \\
+\hline
+MSE 100& 145 / 57 & 301 / 156 & 881 / 826 \\
+\hline
+MSE 40& 434 / 260 & 924 / 628 & 2678 / 2570 \\
+\hline
+\end{tabular}
+\caption{Timing(s) to reach an MSE of 100 and 40 against reference images. In each cell, the left number is for the standard path tracing, the right one  our method}
+\label{tab:timing}
+\end{table}
 
@@
 
-We presented an application of our skeletonization algorithm for global illumination.
-Our strategy works well, it reduces noise by taking into account the distribution of energy in the scene.
-One of the problems is that we take into account only one importance direction for sampling, totally ignoring the BRDF and the cosine factor.
-A solution called Multiple Importance Sampling will allow us to combine different strategies into one.
+We presented in this article a new skeletonization algorithm that both preserves geometrical features of objects and produces a pure curvilinear skeleton.
+These two properties allow us to improve the path tracing algorithm in guiding the rays towards the main illuminated area of the scene.
+Our algorithm is faster and produce less noise than the classical path tracing method.

### 2012-09-14 application_to_path_tracing.tex
Modifying image

@@
 \section{Conclusion}
-
 We presented in this article a new skeletonization algorithm that both preserves geometrical features of objects and produces a pure curvilinear skeleton.

### 2012-09-14 application_to_path_tracing.tex
on va voir

--- 2012-09-14 application_to_path_tracing.tex
@@
 The $f_s$ function expresses how much luminosity is exchanged, on point $x$, between an incoming ray $\Phi$ and the outgoing ray $\Theta$.
-Intuitively it represents the material properties of the surface at point $x$.
+It represents the material properties of the surface at point $x$.
 
@@
 The recursion stops when a maximal number of bounces has been reached.
-%ASK: euh, c'est L ou L_r%
-%c'est bon de dire que ça s'arrete au light%
 
@@
 We average the results and get an estimation of the mean radiance passing through the pixel and heading towards the camera.
-%at different positions over the pixel  ???%
 A bad pdf would lead in picking directions that don't reach the light before the end of the recursion, and produce results with a lot of noise in the final image.
@@
 
-%a revoir, je ne comprends pas tout
-
 \begin{equation*}
@@
 
-It is a product of two functions that can be efficiently sampled individually but that are hard to sample when they are combined.
-The most common strategy used to sample $\Omega_x$ is to use the function $f_s$ because we know it: it is given as an input of the algorithm, for exemple with textures that describe materials.
-
-The term $L$ is rarely used because its evaluation cannot be done in constant time.
-Remember that it represents the distribution of light in the scene.
-It takes high values when $\Phi$ carries a lot of energy.
-Our method gives a way to sample according to $L$.
+The most common strategy used to sample $\Omega_x$ is to use the function $f_s$ because it is an input of the algorithm.
+The term $L$, which represents the distribution of light in the scene, is unknown. Our method gives a way to sample according to $L$.
 
@@
 The skeleton of the voids of the scene is computed using Alg.~\ref{algo::3d_curvskel} and is converted to a graph (the nodes are the 0-faces and the edges are the 1-faces).
-We then compute a set of \myem{importance points}: these points will be used to sample $\Omega_x$ in the path tracing algorithm.
-For each node of the skeleton $n$, one importance point $imp_n$ is computed.
-Intuitively, the importance point associated to $n$ gives us a direction to follow in order to find a light source.
-%TODO a verifier quand meme
+We then compute a set of \myem{importance points}, which will be used to sample $\Omega_x$ in the path tracing algorithm.
+To each node $n$ of the skeleton, one importance point $imp_n$ is computed.
+Intuitively, the importance point associated to $n$ is the direction to follow in order to find a light source.
 
 %La, pas clair !!!
-Let $L$ be a light source of the scene ($L \in S_L$) and $n_L$ the nearest node that is visible by $L$.
-We first compute a tree of shortest paths along the skeleton with $n_L$ as root.
-We use the Dijkstra algorithm with a distance based on visibility from $n_L$: We put $d(n, m) = 1$ if $m$ is visible from $n_L$ and $d(n, m) = 10$ if not.
-It results that the paths that are enlighted by $L$ will have a shorter distance and will guide us to the light faster than the dark ones.
+Let $L$ be a light source of the scene and $n_L$ the nearest node of the skeleton that is visible by $L$.
+For each node, we compute the shortest path to $n_L$ along the skeleton.
+To do so, we use the Dijkstra algorithm and weight the edges of the skeleton depending on a visibility criteria: the weight of an edge $e$ is $1$ if $e$ is visible from $n_L$ and $10$ else.
+It results that the paths that are illuminated by a light will be shorter than paths located in dark areas.
 
-Let $n$ be a node of the skeleton and $V_n$ the set of visible nodes from $n$ along the shortest path toward $n_L$. The importance point $imp_n$ associated to $n$ is the barycenter of $V_n$.
+Let $n$ be a node of the skeleton and $V_n$ the set of visible nodes from $n$ along the shortest path toward $n_L$.
+The importance point $imp_n$ associated to $n$ is the barycenter of $V_n$.
+An example is shown on Fig.~\ref{fig::example_imp}.
 
+\begin{figure}[tb]
 \begin{center}
-\includegraphics[scale=0.5]{images/importance_points.png}
+\begin{tabular}{c|c}
+	\myincludegraphics{0.3\textwidth}{graph_heuristic.pdf} &
+	\myincludegraphics{0.3\textwidth}{graph_heuristic.pdf} \\
+\end{tabular}
+\caption[Example of importance point construction]{\label{fig::example_imp} \textbf{On the left}, we illustrate the importance point $imp_n$ associated to a node $n$.
+The dotted black line shows the closest node to $n$ not included in $V_n$.
+\textbf{On the right}, an example of the curvilinear skeleton (in red) obtained in the Sponza scene.}
 \end{center}
+\end{figure}
 
@@
 
-Given a point $x$ and a direction $\Theta$ we want to compute $L(x \rightarrow \Theta)$ and then sample the hemisphere $\Omega_x$.
-We search for the nearest skeleton node $n$ and its importance point $imp_n$.
-We sample the hemisphere with a power-cosine pdf centered on $\overrightarrow{ximp_n}$:
+Given a point $x$ on the scene and a direction $\Theta$, we want to compute $L(x \rightarrow \Theta)$ and then sample the hemisphere $\Omega_x$.
+We search for the nearest skeleton node $n$ to $x$ and its importance point $imp_n$.
+We sample the hemisphere with a power-cosine pdf centered on $\overrightarrow{x\,imp_n}$:
 \begin{equation*}
@@
 \end{equation*}
-with $\alpha$ the angle between $\overrightarrow{ximp_n}$ and $\Phi$, $s$ being a parameter called skeleton strength. The higher $s$ is, the more we sample close $\overrightarrow{ximp_n}$.
+with $\alpha$ the angle between $\overrightarrow{x\,imp_n}$ and $\Phi$, $s$ being a parameter called skeleton strength. The higher $s$ is, the closer to $\overrightarrow{x\,imp_n}$ we sample.
 
@@
 
-Some results are presented in figure~\ref{fig:results} where we can see the decrease in noise compare to the regular path tracing.
-We computed the MSE value for different scenes that are presented in table~\ref{tab:timing}
+Some results are presented on Fig.~\ref{fig:results} and Fig.~\ref{fig::intro_image} where we can see that our method produce less noisy images compared to the regular path tracing.
+We present, in table~\ref{tab:timing}, the time taken by each method to produce an image of same quality (the quality is measured by the mean square error (MSE) with a reference image) in different scenes.
 
@@
 
-\begin{table}[t!]
+\begin{figure}[tb]
 \centering
@@
 \end{tabular}
-\caption{Timing(s) to reach an MSE of 100 and 40 against reference images. In each cell, the left number is for the standard path tracing, the right one  our method}
+\caption{Time to reach an MSE of 100 and 40 against reference images. In each cell, the left number is for the standard path tracing, the right is for our method}
 \label{tab:timing}
-\end{table}
+\end{figure}
 

### 2012-09-14 skeleton_and_pathtracing.tex
on va voir

@@
 We present a novel 3d curvilinear skeletonization algorithm which produces filtered skeletons without needing any user input, thanks to a new parallel algorithm based on the cubical complex framework.
-We then propose a new path tracing algorithm which uses such curvilinear skeleton in order to produce less noisy images in less time than the classical approach.
+These skeletons are used a new path tracing algorithm in order to produce less noisy images in less time than the classical approach.
 \end{abstract}
@@
 
-Path tracing algorithms are able to render photorealistic images from a scene description by simulating light propagation.
+Path tracing algorithms~\cite{Ka86,LW93} are able to render photorealistic images from a scene description by simulating light propagation.
 The photorealistic aspect is achieved by simulating diffusion of light: elements of the scene which are illuminated by a light source can cast part of the light they receive.
@@
 
-A path tracing algorithm sends rays (of light) from the camera (the viewpoint of the scene), and let them bounce around the scene until they encounter a directly illuminated zone.
-When a ray hits a point of an object (like a wall), it gathers information on the object's material.
-The ray is then reemitted in a random direction.
-Once a ray reaches a light source, the information it gathered through its journey in the scene is averaged in order to decide of the color of the first impact point.
-This color is then allocated to one of the pixels of the image, and the process is repeated for each pixel of the output image.
-To avoid \quotes{unrealistic} results due to an unusual reemission direction of a ray, the algorithm actually sends, for each pixel of the image, many rays: the more rays, the longer the algorithm will take to produce an image, but the more realistic the result will look.
+A path tracing algorithm sends rays from the camera, and let them bounce around the scene until they encounter a directly illuminated zone.
+When a ray hits a point of any surface, it is reemitted in a random direction.
+Once a ray reaches a directly illuminated area, all luminance gathered through its journey in the scene is averaged in order to set the color one of the pixels of the image, and the process is repeated for each pixel of the output image.
+To avoid the noise resulting from this stochastic process, the algorithm actually sends, for each pixel of the image, many rays: the more rays, the longer the algorithm will take to produce an image, but the more realistic the result will look.
 A path tracing can be seen as a random walk process through the scene.
@@
 In order to avoid infinite bouncing of rays in the scene, a limit is set for the number of bounces a ray can do.
-If a ray does not reach any light source after a given number of bounces, it is eliminated from the scene.
-A lost ray is as a waste of computation time, and minimizing this loss would speed up the algorithm.
+If a ray does not reach any directly illuminated area after a given number of bounces, its contribution is null.
+Such lost ray is as a waste of computation time, and minimizing this loss would speed up the algorithm.
 
@@
 To do this, we perform some precomputation on the void space of the scene (the medium where the rays of light propagate) in order to have, later in the algorithm, a clue on towards which direction a ray should bounce in order to meet a light source.
-Such precomputation can take time and become more expensive than the original algorithm.
+Unfortunately, such precomputation may take time and become more expensive than the original algorithm.
 
@@
 \end{tabular}
-\caption{Images $a$ and $b$ shows some rays of the path tracing guided by the material of the objects (image $a$) or our skeleton based method (image $b$).
-Images $c$ and $d$ point out the noise reduction between the classic path tracing (image $c$) and our method (image $d$)}
+
+\caption{\label{fig::intro_image} Images $a$ and $b$ shows some rays of the path tracing guided by the material of the objects (image $a$) or our skeleton based method (image $b$).
+Images $c$ and $d$ point out the noise reduction between the classic path tracing (image $c$) and our method (image $d$).}
 \end{center}
@@
 The purpose of this paper is twofold: we first introduce a new curvilinear skeletonization algorithm (Sec.~\ref{sec::skeletonization}) in the cubical complex framework (Sec.~\ref{sec::binary_to_cubic}), producing filtered skeletons without needing any user input (Sec.~\ref{sec::aspect_no_param}).
-Then, we expose a new method, based on such skeletons, to enhance the results and the performance of a path tracing algorithm (see Sec.~\ref{sec::appli_pathtrace}).
+Then, we expose a new method, based on such skeletons, to enhance the results and the performance of the path tracing algorithm (see Sec.~\ref{sec::appli_pathtrace}).
 

### 2012-09-14 application_to_path_tracing.tex
Modifying caption

--- 2012-09-14 application_to_path_tracing.tex
@@
 \includegraphics[scale=0.17]{images/four_scenes_for_dgci.png}
-\caption{\label{fig:results}Four images comparing the classic path tracing (image $a$ and $c$) and our method (image $b$ and $d$).
-Note the noise reduction of our method in the left wall of images $a$ and $b$, and on the floor on image $c$ and $d$.}
+\caption{\label{fig:results}Same scene using the classical path tracing (image $a$) or our method (image $b$). Images $c$ and $d$ are both details of respectivelly images $a$ and $b$ }
 \end{center}

### 2012-09-14 application_to_path_tracing.tex
Plop
Correction

@@
 \includegraphics[scale=0.17]{images/four_scenes_for_dgci.png}
-\caption{\label{fig:results}Same scene using the classical path tracing (image $a$) or our method (image $b$). Images $c$ and $d$ are both details of respectivelly images $a$ and $b$ }
+\caption{\label{fig:results}Same scene rendered using the classical path tracing (image $a$) or our method (image $b$). Images $c$ and $d$ are both details of respectively images $a$ and $b$ }
 \end{center}

### 2012-09-14 application_to_path_tracing.tex
Remis des modifs qui etaient ecrasees

--- 2012-09-14 application_to_path_tracing.tex
@@
 \includegraphics[scale=0.17]{images/four_scenes_for_dgci.png}
-\caption{\label{fig:results}Four images comparing the classic path tracing (image $a$ and $c$) and our method (image $b$ and $d$).
-Note the noise reduction of our method in the left wall of images $a$ and $b$, and on the floor on image $c$ and $d$.}
+\caption{\label{fig:results}Same scene rendered using the classical path tracing (image $a$) or our method (image $b$). Images $c$ and $d$ are both details of respectively images $a$ and $b$ }
 \end{center}

### 2012-09-14 application_to_path_tracing.tex
Test

@@
 \subsection{Results and discussion}
-
 Some results are presented on Fig.~\ref{fig:results} and Fig.~\ref{fig::intro_image} where we can see that our method produce less noisy images compared to the regular path tracing.

### 2012-09-14 application_to_path_tracing.tex
final ??

--- 2012-09-14 application_to_path_tracing.tex
@@
 
-The skeleton of the voids of the scene is computed using Alg.~\ref{algo::3d_curvskel} and is converted to a graph (the nodes are the 0-faces and the edges are the 1-faces).
+The skeleton of the voids of the scene is computed using Alg.~\ref{algo::3d_curvskel} (as shown on Fig.~\ref{fig::example_imp} on the right) and is converted to a graph (the nodes are the 0-faces and the edges are the 1-faces).
 We then compute a set of \myem{importance points}, which will be used to sample $\Omega_x$ in the path tracing algorithm.
@@
 The importance point $imp_n$ associated to $n$ is the barycenter of $V_n$.
-An example is shown on Fig.~\ref{fig::example_imp}.
+An example is shown on Fig.~\ref{fig::example_imp} on the left.
 
@@
 \begin{center}
-\begin{tabular}{c|c}
-	\myincludegraphics{0.3\textwidth}{graph_heuristic.pdf} &
-	\myincludegraphics{0.3\textwidth}{graph_heuristic.pdf} \\
+\begin{tabular}{l|r}
+	\myincludegraphics{0.3\textwidth}{graph_heuristic.pdf} ~~~~~~&
+	~~~~~~ \myincludegraphics{0.6\textwidth}{skel.jpg} \\
 \end{tabular}
@@
 \hline
-Scene & fig.\ref{fig:redcorridor} & fig.\ref{fig:examples} far left & fig.\ref{fig:examples} far right \\
+Scene & Corridor & Sponza 1 & Sponza 2 (Fig. \ref{fig:results}) \\
 \hline

### 2012-09-14 application_to_path_tracing.tex
Final changes

--- 2012-09-14 application_to_path_tracing.tex
@@
 For each pixel $P$ of the final image, let $x$ be the nearest intersection point of the ray $(O, \overrightarrow{OP} = -\Theta)$ with an element of the scene.
-To obtain the luminosity, we must solve the rendering equation \cite{Ka86} which is a recursive integral equation.
-The integrand of the equation contains the radiance function that we must compute:
+To obtain the luminosity, we must solve the rendering equation \cite{Ka86} which is a recursive integral equation, the integrand containing the radiance function that we must compute:
 
@@
 
-$L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$.
-$L_e$ is the emitted radiance (non null only on light sources).
+$L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$ and $L_e$ is the emitted radiance (non null only on light sources).
 The point $r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$ of the hemisphere $\Omega_x$.
@@
 
-This integrand expresses an intuitive idea: the reflected radiance from $x$ towards the camera is the result of computing all incoming luminosity on $x$ scaled by a factor which depends on the material on $x$ and the angle of the ray going from the camera to $x$.
+The previous equation expresses an intuitive idea: the reflected radiance from $x$ towards the camera is the result of computing all incoming luminosity on $x$ scaled by a factor which depends on the material on $x$ and the angle of the ray going from the camera to $x$.
 The most common method used to estimate the integral, denoted by $L_r(x \rightarrow \Theta)$, is the Monte-Carlo integration which provides the following:
@@
 
-where $p$ is a probability density function (pdf) that is used to sample $\Phi$.
+The function $p$ is a probability density function (pdf) that is used to sample $\Phi$.
 This estimator is unbiased, meaning that the expected value $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$.
@@
 give high density to samples that have high values for the function and low density to samples that have low values).
-The strategy of choosing an adapted pdf is called \myem{importance sampling} and is used in global illumination to improve the convergence speed of the algorithms.
+The strategy of choosing an adapted pdf is called \myem{importance sampling}~\cite{Ka86} and is used in global illumination to improve the convergence speed of the algorithms.
 
@@
 Finally, it returns the sum of the emitted radiance $L_e$ and the result for $L_r$.
-The recursion stops when a maximal number of bounces has been reached.
+The recursion stops when either a maximal number of bounces or a directly illuminated area have been reached.
 
-This computation is done multiple times for each pixel (at different positions over the pixel).
-We average the results and get an estimation of the mean radiance passing through the pixel and heading towards the camera.
+This computation is done multiple times for each pixel. We average the results and get an estimation of the mean radiance passing through the pixel and heading towards the camera.
 A bad pdf would lead in picking directions that don't reach the light before the end of the recursion, and produce results with a lot of noise in the final image.
@@
 The most common strategy used to sample $\Omega_x$ is to use the function $f_s$ because it is an input of the algorithm.
-The term $L$, which represents the distribution of light in the scene, is unknown. Our method gives a way to sample according to $L$.
+The term $L$, which represents the distribution of light in the scene, is unknown. Our method gives a way to sample $L$.
 
@@
 To do so, we use the Dijkstra algorithm and weight the edges of the skeleton depending on a visibility criteria: the weight of an edge $e$ is $1$ if $e$ is visible from $n_L$ and $10$ else.
-It results that the paths that are illuminated by a light will be shorter than paths located in dark areas.
+It results that illuminated paths will be shorter than paths located in dark areas.
 
@@
 \subsection{Results and discussion}
-
-Some results are presented on Fig.~\ref{fig:results} and Fig.~\ref{fig::intro_image} where we can see that our method produce less noisy images compared to the regular path tracing.
-We present, in table~\ref{tab:timing}, the time taken by each method to produce an image of same quality (the quality is measured by the mean square error (MSE) with a reference image) in different scenes.
+Some results are presented on Fig.~\ref{fig:results} and Fig.~\ref{fig::intro_image} where we can see that our method produces less noisy images compared to the regular path tracing.
+We present, in Fig~\ref{tab:timing}, the time taken by each method to produce an image of same quality (the quality is measured by the mean square error (MSE) with a reference image) in different scenes: our method is the fastest.
 
@@
 \end{tabular}
-\caption{Time to reach an MSE of 100 and 40 against reference images. In each cell, the left number is for the standard path tracing, the right is for our method}
+\caption{Time (s) to reach an MSE of 100 and 40 against reference images. In each cell, the left number is for the standard path tracing, the right is for our method}
 \label{tab:timing}

### 2012-09-14 skeleton_and_pathtracing.tex
Change a image
Final changes

@@
 When a ray hits a point of any surface, it is reemitted in a random direction.
-Once a ray reaches a directly illuminated area, all luminance gathered through its journey in the scene is averaged in order to set the color one of the pixels of the image, and the process is repeated for each pixel of the output image.
-To avoid the noise resulting from this stochastic process, the algorithm actually sends, for each pixel of the image, many rays: the more rays, the longer the algorithm will take to produce an image, but the more realistic the result will look.
+Once it reaches a directly illuminated area, all luminance gathered through its journey in the scene is averaged in order to set one of the image pixels, a process repeated for each pixel of the output image.
+To avoid the noise resulting from this stochastic process, the algorithm actually sends many rays for each pixel of the image: the more rays, the longer the algorithm will take to produce an image, but the more realistic the result will look.
 A path tracing can be seen as a random walk process through the scene.
@@
 If a ray does not reach any directly illuminated area after a given number of bounces, its contribution is null.
-Such lost ray is as a waste of computation time, and minimizing this loss would speed up the algorithm.
+Such lost ray is a waste of computation time, and minimizing this loss would speed up the algorithm (see Fig~\ref{fig::intro_image}.a).
 
-Our proposition is to reduce both the time taken by a path tracing algorithms and the noise in the output images by guiding rays towards the scene's light sources, thus reducing the number of rays \quotes{lost} during the computation.
-To do this, we perform some precomputation on the void space of the scene (the medium where the rays of light propagate) in order to have, later in the algorithm, a clue on towards which direction a ray should bounce in order to meet a light source.
+Our proposition is to reduce both the time taken by a path tracing algorithms and the noise in the output images by guiding rays towards the scene's light sources (see Fig~\ref{fig::intro_image}.b), thus reducing the number of rays \quotes{lost} during the computation.
+To do this, we perform some precomputation on the void space of the scene (the medium where the rays of light propagate) in order to have, later in the algorithm, a clue on the direction a ray should bounce in order to meet an illuminated area.
 Unfortunately, such precomputation may take time and become more expensive than the original algorithm.
@@
 To keep the precomputation fast, we perform it only on a subset of points of the voids of the scene.
-The precomputation we want to perform relies on paths and visibility tests, therefore want this set of points to be representative of both the topology and the geometry of the original scene.
+Since this precomputation relies on paths and visibility tests, we want this set of points to be representative of both the topology and the geometry of the original scene.
 For example, it should possess branches propagating in the elongated parts of the scene.
-A filtered curvilinear skeleton of (the voids of) the scene meets all the required properties for our subset of points: it is a thin subset of points of the scene, with the same topology and retaining the main geometrical information of the scene.
+A filtered curvilinear skeleton of the scene's void meets all the required properties for our subset of points: it is a thin subset of points, with the same topology and retaining the main geometrical information of the scene.
 Note that in this application, the filtering of the skeleton (elimination of spurious branches) need to be both efficient and completely automatic, for one cannot ask a user to set parameters for each scene.
@@
 
-\caption{\label{fig::intro_image} Images $a$ and $b$ shows some rays of the path tracing guided by the material of the objects (image $a$) or our skeleton based method (image $b$).
-Images $c$ and $d$ point out the noise reduction between the classic path tracing (image $c$) and our method (image $d$).}
+\caption{\label{fig::intro_image} Note how rays are guided toward the illuminated area thanks to our skeleton based algorithm (image $b$), compared to classical path tracing (image $a$).
+Images $c$ and $d$ point out the noise reduction between the classic path tracing ($c$) and our method ($d$).}
 \end{center}
@@
 The purpose of this paper is twofold: we first introduce a new curvilinear skeletonization algorithm (Sec.~\ref{sec::skeletonization}) in the cubical complex framework (Sec.~\ref{sec::binary_to_cubic}), producing filtered skeletons without needing any user input (Sec.~\ref{sec::aspect_no_param}).
-Then, we expose a new method, based on such skeletons, to enhance the results and the performance of the path tracing algorithm (see Sec.~\ref{sec::appli_pathtrace}).
+Then, we expose a new method, based on such skeletons, to enhance the results (see Fig~\ref{fig::intro_image}.c and d) and the performance of the path tracing algorithm (Sec.~\ref{sec::appli_pathtrace}).
 
@@
 
-\begin{figure}[tb]
+\begin{figure}[bth]
 \begin{center}
@@
 
-Generally, two strategies are possible to achieve this goal: find, during the skeletonization process, points whose neighbourhood configuration seems interesting and keep them in the result \cite{CouCoeurZrour07} \cite{HR08} \cite{CCT09ext}), or choose, before skeletonization, interesting points of the object which should remain untouched, based on a function on these points and a filtering parameter \cite{BertrandCouprie06} \cite{Pal07}.
+Generally, two strategies are possible to achieve this goal: find, during the skeletonization process, points whose neighbourhood configuration seems interesting and keep them in the result \cite{CouCoeurZrour07} \cite{HR08} \cite{CCT09ext}, or choose, before skeletonization, interesting points of the object which should remain untouched, based on a function on these points and a filtering parameter \cite{BertrandCouprie06} \cite{Pal07}.
 
-Algorithm ~\ref{algo::pardircollapse_v0} does not necessarily preserve geometrical features of the input object in the resulting skeleton, which end up being a single vertex.
+Algorithm~\ref{algo::pardircollapse_v0} does not necessarily preserve geometrical features of the input object in the resulting skeleton, which end up being a single vertex.
 
 In the following, we introduce a new method in the cubical complex framework, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory geometrical properties.
-Our method is based on the two previously listed strategies: its finds, during thinning, elements whose neighbourhood configuration is interesting, and uses a function on these elements to decide to preserve them in the result.
+Our method is based on the two previously listed strategies: its finds, during thinning, elements whose neighbourhood configuration is interesting, and uses a function on these elements to decide to preserve them, or not, in the result.
 

### 2012-11-29 skeleton_and_pathtracing.tex
Premières corrections rewiever 1

@@
 
-\title{A 3d curvilinear skeletonization algorithm with application to path tracing}
+\title{A 3D curvilinear skeletonization algorithm with application to path tracing}
 \author{John Chaussard\inst{1} \and Laurent No\"{e}l\inst{2} \and Venceslas Biri\inst{2} \and Michel Couprie\inst{2}\thanks{This work has been partially supported by the \quotes{ANR-2010-BLAN-0205 KIDICO} project}}
@@
 \begin{abstract}
-We present a novel 3d curvilinear skeletonization algorithm which produces filtered skeletons without needing any user input, thanks to a new parallel algorithm based on the cubical complex framework.
-These skeletons are used a new path tracing algorithm in order to produce less noisy images in less time than the classical approach.
+We present a novel 3D curvilinear skeletonization algorithm which produces filtered skeletons without needing any user input, thanks to a new parallel algorithm based on the cubical complex framework.
+These skeletons are used in a new path tracing algorithm in order to produce less noisy images in less time than the classical approach.
 \end{abstract}
@@
 Once it reaches a directly illuminated area, all luminance gathered through its journey in the scene is averaged in order to set one of the image pixels, a process repeated for each pixel of the output image.
-To avoid the noise resulting from this stochastic process, the algorithm actually sends many rays for each pixel of the image: the more rays, the longer the algorithm will take to produce an image, but the more realistic the result will look.
+To reduce the noise resulting from this stochastic process, the algorithm actually sends many rays for each pixel of the image: the more rays, the longer the algorithm will take to produce an image, but the more realistic the result will look.
 A path tracing can be seen as a random walk process through the scene.
@@
 In order to avoid infinite bouncing of rays in the scene, a limit is set for the number of bounces a ray can do.
-If a ray does not reach any directly illuminated area after a given number of bounces, its contribution is null.
+If a ray does not reach any directly illuminated area after a given number of bounces, its contribution is considered null.
 Such lost ray is a waste of computation time, and minimizing this loss would speed up the algorithm (see Fig~\ref{fig::intro_image}.a).
@@
 
-In the 3d voxel framework, objects are made of voxels. In the 3d cubical complex framework, objects are made of cubes, squares, lines and vertices. 
+In the 3D voxel framework, objects are made of voxels. In the 3D cubical complex framework, objects are made of cubes, squares, lines and vertices. 
 Let $\Z$ be the set of integers, we consider the family of sets $\mathbb{F}^1_0$ and $\mathbb{F}^1_1$, such that $\mathbb{F}^1_0 = \{\{a\} \suchthat a \in \Z\}$ and $\mathbb{F}^1_1 = \{ \{a, a+1\} \suchthat a \in \Z\}$.
@@
 
-%Informally, in 3d, a set of faces $X$ is a complex if each side (square) of a cube of $X$ also belong to $X$, each edge of a square of $X$ also belong to $X$, and each vertex of an edge of $X$ also belong to $X$.
+%Informally, in 3D, a set of faces $X$ is a complex if each side (square) of a cube of $X$ also belong to $X$, each edge of a square of $X$ also belong to $X$, and each vertex of an edge of $X$ also belong to $X$.
 
@@
 Traditionally, a binary image is a finite subset of $\Zset^n$ (called voxel image when $n=3$).
-To transpose such image to the cubical complex framework, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$.
+To transpose such an image $S$ to the cubical complex framework, we associate to each element of $S\subseteq \Zset^n$ an n-face of $\allfaces{n}$.
 Let $x=(x_1,...,x_n) \in S$, we define the n-face $\V2C(x) = \{x_1, x_1 + 1\} \times \ldots \times \{x_n, x_n + 1\}$.
@@
 
-We set $\dist1(x,y)$ as the length of a shortest $1$-path from $x$ to $y$; remark that $\dist1$ is indeed the $6$-distance in the 3d voxel framework.
+We set $\dist1(x,y)$ as the length of a shortest $1$-path from $x$ to $y$; remark that $\dist1$ is indeed the $6$-distance in the 3D voxel framework.
 Let $S \subset \Zset^n$, we set $\Compl{S} = \Zset^n \setminus S$, and for all $x \in \Zset^n$, the map $\Dist1{S}: \Zset^n \rightarrow \Nset$ is such that $\Dist1{S}(x) = \displaystyle \min_{y \in \Compl{S}} \dist1(x,y)$.

### 2012-11-29 skeleton_and_pathtracing.tex
Autres corrections reviewer 1, a John de jouer maintenant

@@
 In the following, we introduce a new method in the cubical complex framework, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory geometrical properties.
-Our method is based on the two previously listed strategies: its finds, during thinning, elements whose neighbourhood configuration is interesting, and uses a function on these elements to decide to preserve them, or not, in the result.
+Our method is based on the two previously listed strategies: it finds, during thinning, elements whose neighbourhood configuration is interesting, and uses a function on these elements to decide whether to preserve them, or not, in the result.
 
@@
 The \myem{maximal $1$-ball of $S$ centered on $x$} is the set $\MaxOneBall{S}(x)=\{y \in \Zset^n | \dist1(x,y) < \Dist1{S}(x)\}$.
-We set, for all $x \in S$, the map $\DD1{S}: \Zset^n \rightarrow \Nset$ such that $\DD1{S}(x) = \displaystyle \max_{x \in \MaxOneBall{S}(y)} \Dist1{S}(y)$: this value indicates the radius of the largest maximal $1$-ball contained in $S$ and containing $x$.
+We set, for all $x \in S$, the map $\DD1{S}: \Zset^n \rightarrow \Nset$ such that $\DD1{S}(x) = \displaystyle \max_{x \in \MaxOneBall{S}(y)} \Dist1{S}(y)$: this value indicates the radius of a largest maximal $1$-ball contained in $S$ and containing $x$.
 If $x \in \Compl{S}$, we set $\DD1{S}(x) = 0$.

### 2012-11-29 application_to_path_tracing.tex
Premières corrections reviewer 2 et 3

@@
 The $f_s$ function expresses how much luminosity is exchanged, on point $x$, between an incoming ray $\Phi$ and the outgoing ray $\Theta$.
-It represents the material properties of the surface at point $x$.
+It represents the material properties of the surface at $x$.
 
@@
 The best choice is a pdf that matches the shape of the function to integrate (ie.
-give high density to samples that have high values for the function and low density to samples that have low values).
+gives high density to samples that have high values for the function and low density to samples that have low values).
 The strategy of choosing an adapted pdf is called \myem{importance sampling}~\cite{Ka86} and is used in global illumination to improve the convergence speed of the algorithms.
@@
 This computation is done multiple times for each pixel. We average the results and get an estimation of the mean radiance passing through the pixel and heading towards the camera.
-A bad pdf would lead in picking directions that don't reach the light before the end of the recursion, and produce results with a lot of noise in the final image.
+A bad pdf would lead in picking directions that do not reach the light before the end of the recursion, and produce results with a lot of noise in the final image.
 In the next part, we explain how to produce a pdf based on the curvilinear skeleton.
@@
 
-As stated in the introduction, a curvilinear skeleton of the void (with some preprocessing achieved on it) gives us information on which directions the light comes from.
+As stated in the introduction, a curvilinear skeleton of the void (with some preprocessing performed on it) gives us information on which directions the light comes from.
 Given these directions, we can build a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$.
@@
 %La, pas clair !!!
-Let $L$ be a light source of the scene and $n_L$ the nearest node of the skeleton that is visible by $L$.
+Let $L$ be the light source of the scene and $n_L$ the nearest node of the skeleton that is visible by $L$.
 For each node, we compute the shortest path to $n_L$ along the skeleton.
@@
 An example is shown on Fig.~\ref{fig::example_imp} on the left.
+
+The algorithm can be extended to multiple light sources by taking into account, for each node, only the closest light source when following the curvilinear skeleton path.
 

### 2012-11-29 skeleton_and_pathtracing.tex
Premières corrections reviewer 2 et 3

@@
 Let $\Z$ be the set of integers, we consider the family of sets $\mathbb{F}^1_0$ and $\mathbb{F}^1_1$, such that $\mathbb{F}^1_0 = \{\{a\} \suchthat a \in \Z\}$ and $\mathbb{F}^1_1 = \{ \{a, a+1\} \suchthat a \in \Z\}$.
-Any subset $f$ of $\Z^n$ such that $f$ is the cartesian product of $m$ elements of $\mathbb{F}^1_1$ and $(n-m)$ elements of $\mathbb{F}^1_0$ is called a face or an \myem{$m$-face} of $\Z^n$, $m$ is the dimension of $f$, we write $dim(f)=m$.
+Any subset $f$ of $\Z^n$ such that $f$ is the Cartesian product of $m$ elements of $\mathbb{F}^1_1$ and $(n-m)$ elements of $\mathbb{F}^1_0$ is called a face or an \myem{$m$-face} of $\Z^n$, $m$ is the dimension of $f$, we write $dim(f)=m$.
 A $0$-face is called a \myem{vertex}, a $1$-face is an \myem{edge}, a $2$-face is a \myem{square}, and a $3$-face is a \myem{cube}.
@@
 $\cellstrict{f}$) is \myem{a face of $f$} (resp.
-\myem{a proper face of $f$}.
+\myem{a proper face of $f$}).
 The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$.
@@
 
-Let $\complex{n}{Y}$, the complex $X$ \myem{collapses onto} $Y$ if there exists a sequence of complexes $(X_0,...,X_\ell)$ of $\allfaces{n}$ such that $X=X_0$, $Y=X_\ell$ and for all $i \in \dinterval{1}{\ell}, X_{i}$ is an elementary collapse of $X_{i-1}$.
+Let $\complex{n}{Y}$.
+The complex $X$ \myem{collapses onto} $Y$ if there exists a sequence of complexes $(X_0,...,X_\ell)$ of $\allfaces{n}$ such that $X=X_0$, $Y=X_\ell$ and for all $i \in \dinterval{1}{\ell}, X_{i}$ is an elementary collapse of $X_{i-1}$.
 We also say, in this case, that $Y$ \myem{is a collapse of} $X$.
@@
 Given $S \subset \Zset^n$, the value of $\DD1{S}(x)$ of every $x \in S$ can be naively computed by performing successive morphological dilations of values of the map $\Dist1{S}$.
-A linear algorithm for computing the map $\DD1{S}$ (with regards to the size of the input image) was proposed in \cite{Chaussard_phd}, and will be explored further in details in a future paper.
+A linear algorithm for computing the map $\DD1{S}$ (with regard to the size of the input image) was proposed in \cite{Chaussard_phd}, and will be explored further in details in a future paper.
 
@@
 
-An exemple of these maps is shown on Fig.~\ref{fig::globalopening}.
+An example of these maps is shown on Fig.~\ref{fig::globalopening}.
 
@@
 Informally, if $f$ is a 3-face, then $\cplDist1{Y}(f)$ is the length of the shortest 1-path between the voxel \quotes{corresponding} to $f$ and the set of voxels corresponding to $Y$.
-The same way, we define $\cplDD1{Y}$ and $\cpldecent{Y}$.
+In the same way, we define $\cplDD1{Y}$ and $\cpldecent{Y}$.
 

### 2012-12-03 skeleton_and_pathtracing.tex
Corrections integrees

@@
 
-\def\mydist1{D}
-\def\mydd1{\Omega}
-\def\mydecent{\mathcal{DC}}
-\def\dist1{d}
+\def\mydist1{D_1}
+\def\mydd1{\Omega_1}
+\def\mydecent{\mathcal{DC}_1}
+\def\dist1{d_1}
 \def\Dist1#1{\mydist1(#1)}
@@
 \def\cellstrict#1{\hat{#1}^*}
-\def\closure#1{#1^-}
+\def\closure#1{\widehat{#1}}
 \def\princ#1{#1^+}
@@
 \def\dinterval#1#2{\{#1,\ldots,#2\}}
-\newcommand{\Compl}[1]{\overline{#1}}
+\newcommand{\Compl}[1]{#1^c}
 \def\finproof{\square}
@@
 
-We denote by $\allfaces{n}$ the set composed of all faces in $\Z^n$. Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.
+Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.
+We denote by $\allfaces{n}$ the set composed of all $m$-faces in $\Z^n$ : $\allfaces{n} = \bigcup \limits_{m \in [0;n]} \subfaces{n}{m}$.
 Let $f \in \allfaces{n}$.
@@
 \KwResult{A cubical complex}
-\lnl{pc:01}\While{there exists free faces in $X \setminus W$ and $\ell > 0$}{
-\lnl{pc:02}  $L=\border{X}^-$\;
-\lnl{pc:03}  \For{$t=0 \rightarrow n-1$}{
-\lnl{pc:04}    \For{$s=0 \rightarrow 1$}{
-\lnl{pc:05}      \For{$d=n \rightarrow 1$}{
-\lnl{pc:06}        $E=\{(f,g)$ free for $X \suchthat g \notin W,$\\
-                      $\;$~~~~~~~$\type{f,g}=t,$ $\orient{f,g}=s,$ $\mydim{f}=d\}$\;
-\lnl{pc:07}        $G=\{(f,g) \in E \suchthat f \in L$ and $g \in L\}$\;
-\lnl{pc:08}        $X=X \setminus G$\;
+\While{there exists free faces in $X \setminus W$ and $\ell > 0$}{  \label{pc:01}
+$L=\closure{\border{X}}$\; \label{pc:02}  
+\For{$t=1 \rightarrow n$}{ \label{pc:03}  
+\For{$s=0 \rightarrow 1$}{ \label{pc:04}  
+\For{$d=n \rightarrow 1$}{ \label{pc:05}
+$E=\{(f,g)$ free for $X \suchthat g \notin W,$
+                      $\type{f,g}=t,$ $\orient{f,g}=s,$ $\mydim{f}=d\}$\;  \label{pc:06}       
+$G=\{(f,g) \in E \suchthat f \in L$ and $g \in L\}$\; \label{pc:07}  
+$X=X \setminus G$\;  \label{pc:08}
 }
@@
 }
-\lnl{pc:09}  $l = l-1$\;
+$l = l-1$\; \label{pc:10}
 }
-\lnl{pc:10} \Return $X$\;
+\Return $X$\; \label{pc:11}
 \end{algorithm}
@@
 In the following, we introduce a new method in the cubical complex framework, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory geometrical properties.
-Our method is based on the two previously listed strategies: it finds, during thinning, elements whose neighbourhood configuration is interesting, and uses a function on these elements to decide whether to preserve them, or not, in the result.
+Our method is based on the two previously listed strategies: it finds, during thinning, elements with a specific neighbourhood configuration, and uses a function on these elements to decide whether to preserve them, or not, in the result.
 
@@
 \begin{definition}
-Let $f \in \complex{n}{X}$, the \myem{death date of $f$ in $X$}, denoted by $\death{f}{X}$, is the smallest integer $d$ such that $f \notin ParDirCollapse(X, \emptyset, d)$. 
+Let $f \in \complex{n}{X}$. The \myem{death date of $f$ in $X$}, denoted by $\death{f}{X}$, is the smallest integer $\delta$ such that $f \notin ParDirCollapse(X, \emptyset, \delta)$. 
 \end{definition}
@@
 \begin{definition}
-Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or $f \notin ParDirCollapse(X, \emptyset, b)$.
+Let $f \in \complex{n}{X}$.
+The \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or $f \notin ParDirCollapse(X, \emptyset, b)$.
 \end{definition}
@@
 \begin{definition}
-Let $f \in \complex{n}{X}$, the lifespan of $f$ in $X$ is the integer
+Let $f \in \complex{n}{X}$. The lifespan of $f$ in $X$ is the integer
 
@@
 
-Two points $x,y \in \Zset^n$ are \myem{$1$-neighbours} if the Euclidean distance between $x$ and $y$ is equal or inferior to $1$ (also called direct neighbours).
+%Two points $x,y \in \Zset^n$ are \myem{$1$-neighbours} if the Euclidean distance between $x$ and $y$ is equal to $1$ (also called direct neighbours).
 A \myem{$1$-path from $x$ to $y$} is a sequence $\mathcal{C}=(z_0,...,z_k)$ of points of $\Zset^n$ such that $z_0=x$, $z_k=y$, and for all $j \in [1;k]$, $z_j$ and $z_{j-1}$ are $1$-neighbours.
@@
 
-We set $\dist1(x,y)$ as the length of a shortest $1$-path from $x$ to $y$; remark that $\dist1$ is indeed the $6$-distance in the 3D voxel framework.
+%We set $\dist1(x,y)$ as the length of a shortest $1$-path from $x$ to $y$; remark that $\dist1$ is indeed the $6$-distance in the 3D voxel framework.
+
+We set $\dist1(x,y)$ as the L1 distance between $x$ and $y$ (Manhattan distance).
 Let $S \subset \Zset^n$, we set $\Compl{S} = \Zset^n \setminus S$, and for all $x \in \Zset^n$, the map $\Dist1{S}: \Zset^n \rightarrow \Nset$ is such that $\Dist1{S}(x) = \displaystyle \min_{y \in \Compl{S}} \dist1(x,y)$.
 
-The \myem{maximal $1$-ball of $S$ centered on $x$} is the set $\MaxOneBall{S}(x)=\{y \in \Zset^n | \dist1(x,y) < \Dist1{S}(x)\}$.
+The \myem{maximal $1$-ball in $S$ centered on $x$} is the set $\MaxOneBall{S}(x)=\{y \in \Zset^n | \dist1(x,y) < \Dist1{S}(x)\}$.
 We set, for all $x \in S$, the map $\DD1{S}: \Zset^n \rightarrow \Nset$ such that $\DD1{S}(x) = \displaystyle \max_{x \in \MaxOneBall{S}(y)} \Dist1{S}(y)$: this value indicates the radius of a largest maximal $1$-ball contained in $S$ and containing $x$.
@@
 \end{tabular}
-\caption[Example of opening and decenterness map]{\label{fig::globalopening} \textbf{Example of opening and decenterness map - } From left to right: a shape $S \subset \Zset^2$ (in gray), $\Dist1{S}$, $\DD1{S}$ and $\decent{S}$ (low values have dark colour).}
+\caption[Examples of opening and decenterness map]{\label{fig::globalopening} \textbf{Examples of opening and decenterness map - } From left to right: a shape $S \subset \Zset^2$ (in gray), $\Dist1{S}$, $\DD1{S}$ and $\decent{S}$ (low values have dark colour).}
 \end{center}
@@
 \KwResult{A cubical complex $\complex{3}{Y}$}
-\lnl{curvskel:01} $W=\{f \in X | \lifespan{f}{X} > \cpldecent{X}(f) + \birth{f}{X} - \cplDist1{X}(f)$ and $\mydim{f}=1\}$\;
+$W=\{f \in X | \lifespan{f}{X} > \cpldecent{X}(f) + \birth{f}{X} - \cplDist1{X}(f)$ and $\mydim{f}=1\}$  \label{curvskel:01}  \; 
 \Return $ParDirCollapse(X,W,+\infty)$\;
@@
 In conclusion, $\birth{f}{X} - \cplDist1{X}(f)$ is an approximation of the lifespan of the squares containing $f$.
+
+Although the output of Alg.\ref{algo::3d_curvskel} may contain $2$-faces (depending on the topological properties of the input), the algorithm is said to be a \myem{curvilinear skeletonization} algorithm because it only adds $1$-faces (edges) in the inhibitor set $W$ (used for the visual aspect preservation step) (on line \ref{curvskel:01} of Alg.~\ref{algo::3d_curvskel}).
+For the same reason, we say that the output of Alg.~\ref{algo::3d_curvskel} is a \myem{curvilinear skeleton}.
 

### 2012-12-03 skeleton_and_pathtracing.tex
Petit ajout

@@
 
-Although the output of Alg.\ref{algo::3d_curvskel} may contain $2$-faces (depending on the topological properties of the input), the algorithm is said to be a \myem{curvilinear skeletonization} algorithm because it only adds $1$-faces (edges) in the inhibitor set $W$ (used for the visual aspect preservation step) (on line \ref{curvskel:01} of Alg.~\ref{algo::3d_curvskel}).
+Although the output of Alg.\ref{algo::3d_curvskel} may contain $2$-faces, the algorithm is said to be a \myem{curvilinear skeletonization} algorithm because it only adds $1$-faces (edges) in the inhibitor set $W$ (used for the visual aspect preservation step) (on line \ref{curvskel:01} of Alg.~\ref{algo::3d_curvskel}).
 For the same reason, we say that the output of Alg.~\ref{algo::3d_curvskel} is a \myem{curvilinear skeleton}.
+In Sec.~\ref{sec::appli_pathtrace}, the outputs of Alg.~\ref{algo::3d_curvskel} were always one dimensional complexes.
 

### 2012-12-04 reviews/answer_to_reviewers.tex
Version finale

\documentclass[final,envcountsame]{article}

\usepackage[margin=1.2in]{geometry}

\begin{document}

\title{Answers to the reviewers}
\author{John Chaussard \and Laurent No\"{e}l \and Venceslas Biri \and Michel Couprie}
\maketitle

We would like to thank the reviewers for their time and for all the constructive feedback they provided us.

\section{Answer to reviewer 1}

\begin{itemize}

\item p. 3, §2: I do not know if the notation $X^{­}$ is often used but I feel this may be a bad choice: suggesting that maybe a $X^{+}$ also exists.. Why not use $\widehat{X}$ which will easily recall the corresponding $\hat{f}$ notation?

\textit{Done. We also changed the sign for the complement, which was $\overline{X}$, to $X^c$, in order to make sure the reader will not mistake the overline symbol with the widehat.} \\

\item p. 5 and 8: why did you numbered the first line of each algorithm "2" instead of "1"?!

\textit{This was not made on purpose, and was due to the deprecated \textbackslash lnl tag used to label lines in the algorithm in order to refer to them later. Using \textbackslash label tag instead solved the problem.} \\

\item p. 6, section 4.2, §2: I was not able to figure how two *different* points in $Z^n$ may have an Euclidean distance strictly lower than 1. Did you meant "is equal to 1" instead of "is equal or inferior to 1"? (I assume this was only because the common sentence is "is equal of inferior to a distance $n>1$")

\textit{This was made in order to include the point itself in the set of its direct neighbours. As this is unnecessary, we changed the definition of 1-neighbours to points having an euclidean distance equal to 1. However, we finally decided to remove this part and directly refer to $L_1$ distance, as reviewer 2 suggested.} \\

\item p. 7, end of the same sentence: ".. the radius of the largest maximal 1­ball contained in $S$ and containing $x$.": I see no reason for this maximal ball to be unique. The maximal radius is obviously unique, but many equivalent maximal balls of this radius may cover $x$, isn't it?

\textit{This is right, and it has been corrected in the new version.} \\

\item \begin{itemize}
\item p. 1, in the abstract, "These skeletons are used * a new path tracing": I suppose "in" is missing here. 
\item p. 1, §2, "To avoid the noise resulting..". I suggest to replace "avoid" by "reduce" since, as far as I know, nobody knows a magical method enabling to avoid any error in such a context! 
\item p. 1, §3, "If a ray.. its contribution is * null.": I suggest "is considered null" or "is set to null" since this is not physically correct but practically mandatory. 
\item p. 2, section 2 and further: I would spell "3d" as "3D". 
\item p. 3, §2: I would have defined $F^n_m$ first and $F^n$ just after as the union of $F^n_m$ for all $m$. Defining $F^n$ as "the set of all faces" (without specifically spelling it as "m­faces") may confuse the reader and let him believe that only 2­faces (what is most of the time named "faces") are concerned. 
\item p. 3, §6, "To transpose such *an* image *$S$* to the cubical complex.." 
\item p. 5, last §, "two previously listed strategies: *it* finds": "it" instead of "its". 
\item p. 5, same sentence: "..to decide *whether* to preserve them, or not, in the result." 
\item p. 7, §1: The maximal 1­ball *of* $S$..": I suggest to replace "of" by "in". 
\end{itemize}

\textit{Done.} \\

\end{itemize}

\section{Answer to reviewer 2}

\begin{itemize}

\item The discussion in Section 5.2 assumes some things. For example, if there are different sources of light and the image is quite symmetric, it may easily happen that $imp_n=n$. In this case the discussion does not go through.
Also, I get the feeling that the presentation is tailored to ONE light source. It would be better to make clear in the discussion that the method works for several light­sources.

\textit{If there are multiple light sources in the scene, the one closest to $n$ is chosen in order to computer $imp_n$, therefore, we cannot have both points equal. If $n$ is equidistant to two light sources, then (in the current algorithm's version), we choose randomly one light source.
We added the strategy about multiple light sources at the end of section 5.2, just before the "sampling according to L" paragraph.}

\item In the experiments, there is a significant speed up in most cases, but in Sponza 2. Is there any explanation for this difference in the behavior of the speed up?

\textit{For Sponza 2, some speckles appear in the resulting image. Because of these "extreme pixels", our algorithm takes more time to get close enough to the required MSE. We are currently investigating the source of these speckles, and how to reduce them. Unfortunately, the page restriction set for the DGCI article does not allow us to discuss this matter in the paper.}\\

\item Why do not define d(x,y) as the $L_1$ ­distance instead of using a graph­based distance?

\textit{We changed the way we introduce, in section 4.2, the distance d, and now rely on the well-known definition of the L1 distance}.\\

\item Why do you call the skeleton curvilinear? What is curvilinear about it?

\textit{The skeleton is called curvilinear because the skeletonization algorithm preserves, in the visual aspect preservation step, only edges. Surfaces may appear if the topology of the object requires it : in this case, these surfaces were not preserved for geometrical reasons, but for topological reasons. Surfacic skeletons also exist : in this case, surfaces may be preserved during the visual aspect preservation step. Our algorithm is a curvilinear skeletonization algorithm.
We explained why this terminology choice in the article, at the end of section 4.3.}\\

\item Do not use $d$ in definition 7 because $d$ is already used in the Algorithm. (It is confusing.)

\textit{Done. We use delta instead.}\\

\item I do not understand the 3rd paragraph of Section 4.

\textit{The skeleton obtained using Alg. 1 could end up being too much reduced, as Alg. 1 does not preserve the visual aspect of the object during thinning. For example, the skeleton of a corridor could end up being reduced to a single vertex. We changed this part in order to make this point clearer.}\\

\item The path­tracing algorithm is quite standard and I assume that it appears in textbooks of Comp Graphics. I would suggest referring to one such book (perhaps instead of [8], but both references would also be ok.).

\textit{We keep reference [8] as it is widely used articles about path tracing. Although another reference, such as the one suggested, would be very interesting, the page limitation do not allow us to add a single reference.}

\item \begin{itemize}

\item page 3, line 8: ")" is missing.
\item Def 3: "Let Y..., the complex X collapses....". These are two different sentences and there should be a period between them, not a comma.
\item "elements [whose ...] is interesting"

\end{itemize}

\textit{Done.}

\end{itemize}

\section{Answer to reviewer 3}

\begin{itemize}

\item \begin{itemize}
\item Abstract, l. 4: used in a new path tracing algorithm
\item p. 3, l.1: Cartesian
\item Algorithm 1: for $t=1­­>n$ do
\item p. 5 l. ­3: it finds
\item p. 7 l. 11: with regard. l.17: example; 
\item in Fig. 3. l. ­1: In the same way
\item p. 9 : at point x; at x
\item p. 10: l. 5 gives. l. 19: don't­­ $>$ do not. l. ­18: preprocessing performed on it. 
\end{itemize}

\textit{Done.}\\

\end{itemize}

\end{document}

### 2012-12-04 application_to_path_tracing.tex
Re commit

@@
 
-In this section, we explain how to use the curvilinear skeleton of the voids of the scene in order to enhance the path tracing algorithm.
-We start by defining some basic elements related to path tracing and we present our skeleton based path tracing algorithm.
+In this section, we explain how the curvilinear skeleton of the voids of the scene can enhance the path tracing algorithm.
+We start with some basic elements related to path tracing and then present our modified path tracing algorithm.
 
@@
 To obtain the luminosity, we must solve the rendering equation \cite{Ka86} which is a recursive integral equation, the integrand containing the radiance function that we must compute:
-
-\begin{align*}
-L(x \rightarrow \Theta) &= L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) d\omega_\Phi
-\end{align*}
-
-$L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$ and $L_e$ is the emitted radiance (non null only on light sources).
+\begin{center}
+$L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) d\omega_\Phi$
+\end{center}
+where $L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$ and $L_e$ is the emitted radiance (non null only on light sources).
 The point $r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$ of the hemisphere $\Omega_x$.
 The $f_s$ function expresses how much luminosity is exchanged, on point $x$, between an incoming ray $\Phi$ and the outgoing ray $\Theta$.
-It represents the material properties of the surface at $x$.
 
@@
 This estimator is unbiased, meaning that the expected value $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$.
-The variance of the estimator expresses its quality and depends on the choosen pdf $p$.
+The variance of the estimator expresses its quality and depends on the chosen pdf $p$.
 The best choice is a pdf that matches the shape of the function to integrate (ie.
@@
 
-The algorithm computes $L$ as follow: first, it chooses one random direction $\Phi$ based on the pdf $p$.
-Then, it applies the estimator by calling $L$ recursively to compute $L_r$ (shoots a new ray in the direction $\Phi$, computes the new hit point with the scene, and compute the new radiance at this point).
+The algorithm computes $L$ as follow: it chooses one random direction $\Phi$ based on the pdf $p$ and applies the estimator by calling $L$ recursively to compute $L_r$ (shoots a new ray in the direction $\Phi$, computes the new hit point with the scene, and compute the new radiance at this point).
 Finally, it returns the sum of the emitted radiance $L_e$ and the result for $L_r$.
@@
 Given these directions, we can build a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$.
-
 The integrand of $L_r$ is:
@@
 The most common strategy used to sample $\Omega_x$ is to use the function $f_s$ because it is an input of the algorithm.
-The term $L$, which represents the distribution of light in the scene, is unknown. Our method gives a way to sample $L$.
+The term $L$, representing the distribution of light in the scene, is unknown. Our method gives a way to sample $L$.
 
@@
 An example is shown on Fig.~\ref{fig::example_imp} on the left.
+The algorithm can be extended to multiple light sources by taking into account, when computing the importance point of a skeleton node, only the closest light source.
 
-The algorithm can be extended to multiple light sources by taking into account, for each node, only the closest light source when following the curvilinear skeleton path.
+\subsubsection{Sampling according to $L$}
 
-\begin{figure}[tb]
+Given a point $x$ on the scene and a direction $\Theta$, we want to compute $L(x \rightarrow \Theta)$ and then sample the hemisphere $\Omega_x$.
+We search for the nearest skeleton node $n$ to $x$ and its importance point $imp_n$.
+We sample the hemisphere with a power-cosine pdf centered on $\overrightarrow{x\,imp_n}$:
+\begin{center}
+$p_{skel}(\Phi) = \frac{s + 1}{2\pi}  * cos^s \alpha$
+\end{center}
+with $\alpha$ the angle between $\overrightarrow{x\,imp_n}$ and $\Phi$, $s$ being a parameter called skeleton strength. The higher $s$ is, the closer to $\overrightarrow{x\,imp_n}$ we sample.
+
+\subsection{Results and discussion}
+Some results are presented on Fig.~\ref{fig:results} and Fig.~\ref{fig::intro_image} where we can see that our method produces less noisy images compared to the regular path tracing.
+We present, in Fig~\ref{tab:timing}, the time taken by each method to produce an image of same quality (the quality is measured by the mean square error (MSE) with a reference image) in different scenes: our method is the fastest.
+
+\section{Conclusion}
+We presented in this article a new skeletonization algorithm that both preserves geometrical features of objects and produces a pure curvilinear skeleton.
+These two properties allow us to improve the path tracing algorithm in guiding the rays towards the main illuminated area of the scene.
+Our algorithm is faster and produce less noise than the classical path tracing method.
+
+\begin{figure}[t]
 \begin{center}
@@
 
-\subsubsection{Sampling according to $L$}
-
-Given a point $x$ on the scene and a direction $\Theta$, we want to compute $L(x \rightarrow \Theta)$ and then sample the hemisphere $\Omega_x$.
-We search for the nearest skeleton node $n$ to $x$ and its importance point $imp_n$.
-We sample the hemisphere with a power-cosine pdf centered on $\overrightarrow{x\,imp_n}$:
-\begin{equation*}
-p_{skel}(\Phi) = \frac{s + 1}{2\pi} cos^s \alpha
-\end{equation*}
-with $\alpha$ the angle between $\overrightarrow{x\,imp_n}$ and $\Phi$, $s$ being a parameter called skeleton strength. The higher $s$ is, the closer to $\overrightarrow{x\,imp_n}$ we sample.
-
-\subsection{Results and discussion}
-Some results are presented on Fig.~\ref{fig:results} and Fig.~\ref{fig::intro_image} where we can see that our method produces less noisy images compared to the regular path tracing.
-We present, in Fig~\ref{tab:timing}, the time taken by each method to produce an image of same quality (the quality is measured by the mean square error (MSE) with a reference image) in different scenes: our method is the fastest.
-
-\begin{figure}[tbhp]
+\begin{figure}[t]
 \begin{center}
@@
 \end{figure}
-
-\section{Conclusion}
-We presented in this article a new skeletonization algorithm that both preserves geometrical features of objects and produces a pure curvilinear skeleton.
-These two properties allow us to improve the path tracing algorithm in guiding the rays towards the main illuminated area of the scene.
-Our algorithm is faster and produce less noise than the classical path tracing method.

### 2012-12-04 skeleton_and_pathtracing.tex
Re commit

@@
 \def\cell#1{\hat{#1}}
-\def\cellstrict#1{\hat{#1}^*}
+\def\cellstrict#1{\cell{#1}^*}
 \def\closure#1{\widehat{#1}}
-\def\princ#1{#1^+}
+\def\princ#1{\check{#1}}
 \def\mydim#1{\dim(#1)}
@@
 We present a novel 3D curvilinear skeletonization algorithm which produces filtered skeletons without needing any user input, thanks to a new parallel algorithm based on the cubical complex framework.
-These skeletons are used in a new path tracing algorithm in order to produce less noisy images in less time than the classical approach.
+These skeletons are used in a modified path tracing algorithm in order to produce less noisy images in less time than the classical approach.
 \end{abstract}
@@
 
-Algorithm~\ref{algo::pardircollapse_v0} does not necessarily preserve geometrical features of the input object in the resulting skeleton, which end up being a single vertex.
-
+Algorithm~\ref{algo::pardircollapse_v0} does not necessarily preserve geometrical features of the input object in the resulting skeleton (for example, the skeleton of a corridor could be reduced to a single vertex).
 In the following, we introduce a new method in the cubical complex framework, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory geometrical properties.
@@
 For the same reason, we say that the output of Alg.~\ref{algo::3d_curvskel} is a \myem{curvilinear skeleton}.
-In Sec.~\ref{sec::appli_pathtrace}, the outputs of Alg.~\ref{algo::3d_curvskel} were always one dimensional complexes.
+In the scenes studied in Sec.~\ref{sec::appli_pathtrace}, the outputs of Alg.~\ref{algo::3d_curvskel} were one dimensional complexes.
 
@@
 \begin{tabular}{cc|cc}
-	\myincludegraphics{0.18\textwidth}{hand2.jpg} &
-	\myincludegraphics{0.18\textwidth}{hand_curvskel.jpg} &
-	\myincludegraphics{0.24\textwidth}{neptune.jpg} &
-	\myincludegraphics{0.24\textwidth}{neptune_curvskel.jpg} \\
+	\myincludegraphics{0.16\textwidth}{hand2.jpg} &
+	\myincludegraphics{0.16\textwidth}{hand_curvskel.jpg} &
+	\myincludegraphics{0.21\textwidth}{neptune.jpg} &
+	\myincludegraphics{0.21\textwidth}{neptune_curvskel.jpg} \\
 \end{tabular}

### 2012-12-05 reviews/answer_to_reviewers.tex
commit or not commit ?

@@
 
-\textit{If there are multiple light sources in the scene, the one closest to $n$ is chosen in order to computer $imp_n$, therefore, we cannot have both points equal.
-If $n$ is equidistant to two light sources, then (in the current algorithm's version), we choose randomly one light source.
+\textit{If there are multiple light sources in the scene, the one closest to $n$ is chosen in order to compute $imp_n$, therefore, we cannot have both points equal.
+If $n$ is equidistant from two light sources, then (in the current algorithm's version), we choose randomly one light source.
 We added the strategy about multiple light sources at the end of section 5.2, just before the "sampling according to L" paragraph.}
@@
 Surfaces may appear if the topology of the object requires it : in this case, these surfaces were not preserved for geometrical reasons, but for topological reasons.
-Surfacic skeletons also exist : in this case, surfaces may be preserved during the visual aspect preservation step.
+Surface skeletons also exist : in this case, surfaces may be preserved during the visual aspect preservation step.
 Our algorithm is a curvilinear skeletonization algorithm.
-We explained why this terminology choice in the article, at the end of section 4.3.}\\
+We explained this terminology choice in the article, at the end of section 4.3.}\\
 

### 2012-12-05 application_to_path_tracing.tex
commit or not commit ?

@@
 
-The algorithm computes $L$ as follow: it chooses one random direction $\Phi$ based on the pdf $p$ and applies the estimator by calling $L$ recursively to compute $L_r$ (shoots a new ray in the direction $\Phi$, computes the new hit point with the scene, and compute the new radiance at this point).
+The algorithm computes $L$ as follow: it chooses one random direction $\Phi$ based on the pdf $p$ and applies the estimator by calling $L$ recursively to compute $L_r$ (shoots a new ray in the direction $\Phi$, computes the new hit point with the scene, and computes the new radiance at this point).
 Finally, it returns the sum of the emitted radiance $L_e$ and the result for $L_r$.
@@
 
-The skeleton of the voids of the scene is computed using Alg.~\ref{algo::3d_curvskel} (as shown on Fig.~\ref{fig::example_imp} on the right) and is converted to a graph (the nodes are the 0-faces and the edges are the 1-faces).
+The skeleton of the voids of the scene is computed using Alg.~\ref{algo::3d_curvskel} (as shown on Fig.~\ref{fig::example_imp} on the right) and is converted into a graph (the nodes are the 0-faces and the edges are the 1-faces).
 We then compute a set of \myem{importance points}, which will be used to sample $\Omega_x$ in the path tracing algorithm.
@@
 Given a point $x$ on the scene and a direction $\Theta$, we want to compute $L(x \rightarrow \Theta)$ and then sample the hemisphere $\Omega_x$.
-We search for the nearest skeleton node $n$ to $x$ and its importance point $imp_n$.
+We look for the nearest skeleton node $n$ to $x$ and its importance point $imp_n$.
 We sample the hemisphere with a power-cosine pdf centered on $\overrightarrow{x\,imp_n}$:
@@
 Some results are presented on Fig.~\ref{fig:results} and Fig.~\ref{fig::intro_image} where we can see that our method produces less noisy images compared to the regular path tracing.
-We present, in Fig~\ref{tab:timing}, the time taken by each method to produce an image of same quality (the quality is measured by the mean square error (MSE) with a reference image) in different scenes: our method is the fastest.
+We present, in Fig~\ref{tab:timing}, the time taken by each method to produce an image of same quality (the quality is measured by the mean square error (MSE) with a reference image) in different scenes: our method is the fastest one.
 

### 2012-12-05 skeleton_and_pathtracing.tex
commit or not commit ?

@@
 \and
-Universit\'{e} Paris Est, LIGM, A3SI-ESIEE \\ 2, boulevard Blaise Pascal, 93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}
+Universit\'{e} Paris Est, LIGM, A3SI-ESIEE, CNRS(UMR 8049) \\ 2, boulevard Blaise Pascal, 93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}
 

### 2012-12-05 skeleton_and_pathtracing.tex
Ajout de l'institut pour vence et laurent

--- 2012-12-04 skeleton_and_pathtracing.tex
@@
 \title{A 3D curvilinear skeletonization algorithm with application to path tracing}
-\author{John Chaussard\inst{1} \and Laurent No\"{e}l\inst{2} \and Venceslas Biri\inst{2} \and Michel Couprie\inst{2}\thanks{This work has been partially supported by the \quotes{ANR-2010-BLAN-0205 KIDICO} project}}
+\author{John Chaussard\inst{1} \and Laurent No\"{e}l\inst{2} \and Venceslas Biri\inst{2} \and Michel Couprie\inst{3}\thanks{This work has been partially supported by the \quotes{ANR-2010-BLAN-0205 KIDICO} project}}
 \institute{Universit\'{e} Paris 13, Sorbonne Paris Cit\'{e}, LAGA, CNRS(UMR 7539), \\ F-93430, Villetaneuse, France \\ \email{chaussard@math.univ-paris13.fr}
 \and
-Universit\'{e} Paris Est, LIGM, A3SI-ESIEE \\ 2, boulevard Blaise Pascal, 93162 Noisy le Grand CEDEX, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr, michel.couprie@esiee.fr}}
+Universit\'{e} Paris Est, LIGM, CNRS(UMR 8049) \\ Cit\'{e} Descartes, 5 boulevard Descartes, 77454 Marne-la-Vallée CEDEX 22, France \\ \email{laurent.noel@esiee.fr, v.biri@esiee.fr}
+\and
+Universit\'{e} Paris Est, LIGM, A3SI-ESIEE, CNRS(UMR 8049) \\ 2, boulevard Blaise Pascal, 93162 Noisy le Grand CEDEX, France \\ \email{michel.couprie@esiee.fr}
+}
 
@@
 
-The collapse operation is the basic operation for performing homotopic thinning of a complex, and consists of removing two distinct faces $(f,g)$ from a complex $X$ under the condition that they form a free pair:
+The collapse operation is the basic operation for performing homotopic thinning of a complex, and consists of removing free pairs of faces:
 
@@
 \begin{tabular}{cc|cc}
-	\myincludegraphics{0.16\textwidth}{hand2.jpg} &
-	\myincludegraphics{0.16\textwidth}{hand_curvskel.jpg} &
+	\myincludegraphics{0.20\textwidth}{hand2.jpg} &
+	\myincludegraphics{0.20\textwidth}{hand_curvskel.jpg} &
 	\myincludegraphics{0.21\textwidth}{neptune.jpg} &
